### 2018-09-08 Project1/report/ms.tex
Updates from ShareLaTeX

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}

\begin{document}

\title{A spectacular title}

\author{Stig-Nicolai Foyn, Christer Dreierstad}

\email{stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}

\altaffiltext{1}{Institute of physics, University of
  Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}

%\date{Received - / Accepted -}

\begin{abstract}
Comparing general and tailored methods for solving second order differential equations this study will present that producing a specialized method reduces the run time of a numerical solver without losing numerical accuracy. Reducing the floating point operations (FLOPS) will be proven efficient for large scale calculations.
  %State problem. Briefly describe method and data. Summarize main results.
\end{abstract}
\keywords{computational science --- numerical methods: error estimation, run time --- methods: numerical, Triadiagonal marices}

\section{Introduction}
\label{sec:introduction}
This study on numerical algorithms, generalized and tailored for specific problems,  will solve a differential equation numerically. Analyzing the results, comparing the run time and observing the numerical error of the different methods.

Discuss background, physical importance and possibly some history of
the problem that is being studied in this paper.

\section{Method}
\label{sec:method}

\subsection{Generalized tridiagonal matrix}

\subsection{Specific algorithm}

\subsection{Armadillo}

Describe method. Define data model and likelihood. Outline how the
likelihood was computed (grid or MCMC).

Define the power law model in terms of $Q$ and $n$.

Summarize properties of data. Which data are used (experiment,
frequencies etc.)? Pixel resolution ($N_{\textrm{side}}$),
$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis
for other researchers.

Show a sky map of the smoothed data. Use the Healpix routine
``smoothing'' to do this; it works just like anafast. Smooth with a
$7^{\circ}$ beam, and plot with ``map2gif''. Show the RMS pattern as
well.

\section{Results}
\label{sec:results}
\subsection{}

Show the 2D likelihood contours. Summarize constraints on $Q$ and
$n$.

\section{Conclusions}
\label{sec:conclusions}

Summarize results. Discuss their importance, referring to the
discovery to the initial seeds for structure formation. Mention that
these results are in good agreement with expectations from
inflationary theory.

%\begin{figure}[t]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

\begin{deluxetable}{lccc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Summary of main results.}
\tablecolumns{4}
\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
\startdata
Item 1 & Item 2 & Item 3 & Item 4
\enddata
\end{deluxetable}

\begin{acknowledgements}
  Who do you want to thank for helping out with this project?
\end{acknowledgements}

\begin{thebibliography}{}

\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89

\end{thebibliography}

\end{document}

### 2018-09-08 Project1/report/ms.tex
First draft, not done.

@@
 
-\title{A spectacular title}
+\title{Project 1}
 
@@
 Reducing the floating point operations (FLOPS) will be proven efficient for large scale calculations.
+
   %State problem. Briefly describe method and data. Summarize main results.
 \end{abstract}
-\keywords{computational science --- numerical methods: error estimation, run time --- methods: numerical, Triadiagonal marices}
+\keywords{computational science --- numerical methods: error estimation, run time --- methods: numerical, tridiagonal matrices}
 
@@
 \label{sec:introduction}
-This study on numerical algorithms, generalized and tailored for specific problems,  will solve a differential equation numerically.
-Analyzing the results, comparing the run time and observing the numerical error of the different methods.
+This study on generalized and tailored numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations.
+By comparing the run time and observing the numerical error of the calculations .
+To specify, the equation in question is the one dimensional Poisson equation:
+%
+\begin{equation}
+    {\nabla^{2}} \Phi = -4\pi \rho(r),
+\end{equation}
+%
+where $\Phi$ is the electrostatic potential generated by a localized charge distribution and $\rho$ is the charge density.
+This equation is used in electromagnetism, but in this case it was only necessary to analyze a normalized equation on the same form.
+Our normalized one dimensional Poisson equation can then be written as:
+%
+\begin{equation}\label{eq:-u''}
+    -u''(x) = f(x).
+\end{equation}
+%
+Further rewriting this as a set of linear equation will allow us to solve it using our numerical algorithms.
 
-Discuss background, physical importance and possibly some history of
-the problem that is being studied in this paper.
+%Discuss background, physical importance and possibly some history of
+%the problem that is being studied in this paper.
 
@@
 \label{sec:method}
+For the study we consider a solution for the source term $f(x)$ to be
+%
+\begin{gather}\label{eq:f(x)}
+    f(x) = 100e^{-10x}.
+\end{gather}
+%
+This equation has a closed form solution 
+%
+\begin{gather}\label{eq:u(x)}
+u(x) = 1-(1-e^{10})x-e^{-10x},
+\end{gather}
+%
+which yields $-f(x)$ when taking the second derivative. We consider the Poisson equation to uphold Dirichlet boundary conditions; that is $u(0)=u(1)=0$, so $x \in [0,1]$.
+\[
+\begin{array}{c}
+\begin{bmatrix}\label{eq:Au=f}
+b_1     & c_1           & 0         & \dots     & \dots     & 0 \\
+a_1     & b_2           & c_2       & 0         & \dots     & \dots \\
+0       & a_2           & b_3       & c_3       & 0     & \dots\\
+\dots  &  0            & a_3       & b_4       & \dots    & 0 \\
+\dots  & \dots        & 0 & \dots    & \dots    & c_{n-1} \\
+0       & \dots         & \dots         & 0         & a_{n-1}   & b_n   \\
+\end{bmatrix}
+\begin{bmatrix}
+v_1 \\
+v_2 \\
+\dots \\
+\dots \\
+\dots \\
+v_n
+\end{bmatrix}
+=
+\begin{bmatrix}
+\Tilde{f}_1 \\
+\Tilde{f}_2 \\
+\dots \\
+\dots \\
+\dots \\
+\Tilde{f}_n
+\end{bmatrix}
+\end{array}
+
+\]
 
@@
 
-Describe method. Define data model and likelihood. Outline how the
-likelihood was computed (grid or MCMC).
-
-Define the power law model in terms of $Q$ and $n$.
-
-Summarize properties of data. Which data are used (experiment,
-frequencies etc.)? Pixel resolution ($N_{\textrm{side}}$),
-$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis
-for other researchers.
-
-Show a sky map of the smoothed data. Use the Healpix routine
-``smoothing'' to do this; it works just like anafast. Smooth with a
-$7^{\circ}$ beam, and plot with ``map2gif''. Show the RMS pattern as
-well.
+%Summarize properties of data.
+Which data are used (experiment,frequencies etc.)?
+Pixel resolution ($N_{\textrm{side}}$),$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysisfor other researchers.
 
@@
 \label{sec:results}
-\subsection{}
+\subsection{Numerical error}
 
-Show the 2D likelihood contours. Summarize constraints on $Q$ and
-$n$.
+\begin{figure}[H]
+    \centering
+    \includegraphics[size=0.25\textwidth]{p1bplot.png}
+    \caption{The solution of the differential equation using a generalized algorithm, the three different plots show the $nxn$ matrix with $n=10$, $n=100$, $n=1000$ }
+    \label{fig1: Generalized algorithm}
+\end{figure}
+
+\subsection{Program CPU time}
+\begin{deluxetable}{lccc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{This table contains average CPU times for all the algorithms}
+\tablecolumns{4}
+\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
+\startdata
+Item 1 & Item 2 & Item 3 & Item 4
+\enddata
+\end{deluxetable}
 
@@
 
-Summarize results. Discuss their importance, referring to the
-discovery to the initial seeds for structure formation. Mention that
-these results are in good agreement with expectations from
-inflationary theory.
+%Summarize results.
+Discuss their importance, referring to the discovery to the initial seeds for structure formation.
+Mention that these results are in good agreement with expectations from inflationary theory.
 
@@
 \begin{acknowledgements}
-  Who do you want to thank for helping out with this project?
+
 \end{acknowledgements}

### 2018-09-10 Project1/report/ms.tex
Updates from ShareLaTeX

@@
 \usepackage{epsfig,floatflt}
+\usepackage{fourier}
+
+\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
+\SetMathAlphabet{\mathcal}{bold}{OMS}{cmsy}{b}{n}
 
@@
 %
-\begin{equation}
+\begin{equation*}
     {\nabla^{2}} \Phi = -4\pi \rho(r),
-\end{equation}
+\end{equation*}
 %
@@
 %
-\begin{equation}\label{eq:-u''}
-    -u''(x) = f(x).
-\end{equation}
-%
+\begin{equation*}\label{eq:-u''}
+    -\frac{du^2}{dx^2} = f(x).
+\end{equation*}
+
 Further rewriting this as a set of linear equation will allow us to solve it using our numerical algorithms.
@@
 \begin{gather}\label{eq:f(x)}
-    f(x) = 100e^{-10x}.
+    f(x) = 100e^{-10x},
 \end{gather}
 %
-This equation has a closed form solution 
+which has a closed form solution 
 %
@@
 %
-which yields $-f(x)$ when taking the second derivative. We consider the Poisson equation to uphold Dirichlet boundary conditions; that is $u(0)=u(1)=0$, so $x \in [0,1]$.
-\[
+which yields $-f(x)$ when taking the second derivative.
+
+We consider the one dimensional Poisson equation to uphold Dirichlet boundary conditions; $u(0)=u(1)=0$, so $x \in [0,1]$. The second order derivative $u(x)$ can be represented as
+%
+\begin{gather*}
+ -\frac{du^2}{dx^2} = -\left(\frac{u(x+h) + u(x-h) - 2u(x)}{h^2} + \mathcal{O}(h^2)\right),
+\end{gather*}
+%
+where h is the step size. By approximating the above equation and discretizing over x we get
+%
+\begin{gather*}
+    -\frac{du^2}{dx^2} = -\frac{u_{i+1} + u_{i-1} - 2u_i}{h^2} = f(x_i) = f_i.
+\end{gather*}
+%
+To simplify the calculations and reduce the FLOPS we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
+%
+\begin{gather}\label{eq:u_discretized}
+    u_{i+1} + u_{i-1} - 2u_i = f'_i.
+\end{gather}
+%
+The left hand side of the above equation can be represented as a matrix vector multiplication.
+
+\[ \boldsymbol{Au} =
 \begin{array}{c}
@@
 \begin{bmatrix}
-v_1 \\
-v_2 \\
+u_1 \\
+u_2 \\
 \dots \\
@@
 \dots \\
-v_n
+u_n
 \end{bmatrix}
@@
 \begin{bmatrix}
-\Tilde{f}_1 \\
-\Tilde{f}_2 \\
+f'_1 \\
+f_2 \\
 \dots \\
@@
 \dots \\
-\Tilde{f}_n
+f'_n
 \end{bmatrix}
 \end{array}
+= \boldsymbol{f'_i}
+\].
 
-\]
+%\begin{bmatrix}
+%\Tilde{f}_1 \\
+%\Tilde{f}_2 \\
+%\dots \\
+%\dots \\
+%\dots \\
+%\Tilde{f}_n
+%\end{bmatrix}
+%\end{array}
 

### 2018-09-10 Project1/pplot1d.py
Plot for 1d

@@ -12,7 +12,8 @@ for lines in open(infile, "r"):
     log10h.append(float(col[1]))
 
 plt.figure()
+plt.title('Logarithmic plot of error as a function of step size h',size=18)
 plt.plot(log10h,eps)
-plt.xlabel('$\\log_{10}(h)$');plt.ylabel('$\\epsilon_i$')
-plt.legend(['$\\epsilon_i$'])
+plt.xlabel('$\\log_{10}(h)$', size=18);plt.ylabel('$\\epsilon_i$',size=18)
+plt.legend(['$\\epsilon_i$'],prop={'size': 18})
 plt.show()

### 2018-09-10 Project1/report/ms.tex
Updates from ShareLaTeX

@@
 %
-\begin{equation*}\label{eq:-u''}
+\begin{equation}\label{eq:-u''}
     -\frac{du^2}{dx^2} = f(x).
-\end{equation*}
+\end{equation}
 
@@
 %
-\begin{gather}\label{eq:f(x)}
+\begin{gather*}\label{eq:f(x)}
     f(x) = 100e^{-10x},
-\end{gather}
+\end{gather*}
 %
@@
 %
-\begin{gather*}
+\begin{gather}\label{eq:du/dx}
  -\frac{du^2}{dx^2} = -\left(\frac{u(x+h) + u(x-h) - 2u(x)}{h^2} + \mathcal{O}(h^2)\right),
-\end{gather*}
+\end{gather}
 %
-where h is the step size. By approximating the above equation and discretizing over x we get
+where h is the step size.
+
+\subsection{Numerical representation}
+By approximating equation \eqref{eq:du/dx} and discretizing over x we get
 %
@@
 %
-To simplify the calculations and reduce the FLOPS we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
+To simplify the calculations and reduce the FLOPS, we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
 %
 \begin{gather}\label{eq:u_discretized}
-    u_{i+1} + u_{i-1} - 2u_i = f'_i.
+    -\left(u_{i-1} - 2u_i + u_{i+1}\right) = f'_i.
 \end{gather}
 %
-The left hand side of the above equation can be represented as a matrix vector multiplication.
-
+The linear combination above can be represented by a matrix vector multiplication.
+The matrix will consist of the coefficients of the $u_i$'s along the central, upper and lower diagonal, making up a tridiagonal matrix.
+The coefficient, after taking the outer sign into  consideration, are -1, 2 and -1 respectively.
+The vector will consist of the values of the $u_i$'s.
+The matrix representation will therefore be on the following form:
+%
 \[ \boldsymbol{Au} =
@@
 \end{array}
-= \boldsymbol{f'_i}
-\].
-
+= \boldsymbol{f'_i}.
+\]
+%
+Since we are considering fixed end points for $u$ we do not include $i = 0$ and $i = n+1$ in the calculations for updating the $u_i$'s.
 %\begin{bmatrix}
@@
 %\end{array}
-
-\subsection{Generalized tridiagonal matrix}
 
@@
     \label{fig1: Generalized algorithm}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[size=0.25\textwidth]{plot1d.png}
+    \caption{}
+    \label{fig1:1d}
 \end{figure}

### 2018-09-10 Project1/report/ms.tex
Updates from ShareLaTeX

@@
 \tablecaption{\label{tab:results}}
-\tablecomments{This table contains average CPU times for all the algorithms}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 10^{6}}
 \tablecolumns{4}
-\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
+\tablehead{ & General & Taylored & Armadillo}
 \startdata
-Item 1 & Item 2 & Item 3 & Item 4
+Highest & Item 2 & Item 3 & Item 4 \\
+Lowest & Item 2 & Item 3 & Item 4 \\
+Average & Item 2 & Item 3 & Item 4 
 \enddata

### 2018-09-10 Project1/pplot1d.py
Updated 1d plot

@@ -15,5 +15,5 @@ plt.figure()
 plt.title('Logarithmic plot of error as a function of step size h',size=18)
 plt.plot(log10h,eps)
 plt.xlabel('$\\log_{10}(h)$', size=18);plt.ylabel('$\\epsilon_i$',size=18)
-plt.legend(['$\\epsilon_i$'],prop={'size': 18})
+plt.legend(['$\\log10(\\epsilon_i)$'],prop={'size': 18})
 plt.show()

### 2018-09-10 Project1/p1.cpp
Added plots for 1b

@@ -66,7 +66,7 @@ int main(int argc, char *argv[]){
         bt[0] = b[0];
         bt[1] = b[0];
 
-        for(int i = 2; i < n+2; i++) { //n+1 eller n???
+        for(int i = 2; i < n+2; i++) {
                 bt[i] = b[i] - a[i-1]*c[i-1]/bt[i-1];
                 ft[i] = fprime[i] - ft[i-1]*a[i-1]/bt[i-1];
         }

### 2018-09-10 Project1/pplot.py

@@ -14,8 +14,10 @@ for lines in open(infile, "r"):
     cf.append(float(col[2]))
 
 plt.figure()
+plt.title('Numerical and analytical solution of u(x)',size=18)
 plt.plot(x,u)
 plt.plot(x,cf)
-plt.legend(['u','cf'])
-
+plt.xlabel('x',size=18)
+plt.ylabel('u',size=18)
+plt.legend(['Numerical','Analytical'],prop={'size': 18})
 plt.show()

### 2018-09-10 Project1/pplot1d.py

@@ -1,7 +1,12 @@
 import matplotlib.pyplot as plt
 import sys
+import numpy as np
 
 infile = sys.argv[1]
+
+def y(h):
+    h = np.array(h)
+    return 2*h
 
 log10h = []
 eps = []
@@ -14,6 +19,7 @@ for lines in open(infile, "r"):
 plt.figure()
 plt.title('Logarithmic plot of error as a function of step size h',size=18)
 plt.plot(log10h,eps)
+plt.plot(log10h,y(log10h))
 plt.xlabel('$\\log_{10}(h)$', size=18);plt.ylabel('$\\epsilon_i$',size=18)
-plt.legend(['$\\log10(\\epsilon_i)$'],prop={'size': 18})
+plt.legend(['$log_{10}(\\epsilon_i)$', '$2log_{10}(h)$'],prop={'size': 18})
 plt.show()

### 2018-09-10 Project1/report/ms.tex
Updates from ShareLaTeX

@@
 \usepackage{fourier}
+\usepackage{listings}
+\newcommand{\R}{\mathbb{R}}
+
+\lstset{language=python,
+    showspaces=false,
+    extendedchars=true,
+    breaklines=true,
+    tabsize=2,
+    basicstyle=\small\ttfamily,
+    frame=single,
+    columns=flexible,
+    keepspaces=true,
+    }
 
@@
 Further rewriting this as a set of linear equation will allow us to solve it using our numerical algorithms.
+As generalized algorithm we will be using a pre-built function for LU-decomposition.
+Specializing our algorithm further for the problem we will make a generalized algorithm for any solving any tridiagonal matrix.
+Lastly we will make an algorithm completely tailored to our problem where the matrix elements are already specified.
 
@@
 %
+Where the step size $h = 1/(n+1)$. When calculating $x_i = i\times h$ the calculations runs over $1$ to $n+1$, such that the last
+%
+\begin{gather}
+    x_{n+1} = (n+1)h = (n+1)/(n+1) = 1,
+\end{gather}
+%
+which ensures that $x \in [0,1]$.
 To simplify the calculations and reduce the FLOPS, we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
@@
 The matrix will consist of the coefficients of the $u_i$'s along the central, upper and lower diagonal, making up a tridiagonal matrix.
-The coefficient, after taking the outer sign into  consideration, are -1, 2 and -1 respectively.
+The coefficient, after taking the outer sign of the left hand side of equation \eqref{eq:u_discretized} into consideration, are -1, 2 and -1 respectively.
 The vector will consist of the values of the $u_i$'s.
@@
 f'_1 \\
-f_2 \\
+f'_2 \\
 \dots \\
@@
 Since we are considering fixed end points for $u$ we do not include $i = 0$ and $i = n+1$ in the calculations for updating the $u_i$'s.
+
+\subsection{General algorithm}
+We consider the matrix $\boldsymbol{A}$ to consist of three vectors $\boldsymbol{a}$, $\boldsymbol{b}$ and $\boldsymbol{c}$ consisting of all $a_i$, $b_i$ and $c_i$ respectively.
+To solve the matrix equation we perform a forward and a backward substitution of $\boldsymbol{A}$ and $\boldsymbol{f'}$ respectively.
+
+In order to forward substitute $\boldsymbol{A}$, the matrix must be upper triangular, which is done by Gaussian elimination.
+An example with a $4x4$-matrix $\boldsymbol{E}$ and two $4x1$ vectors $x$ and $g$ follows to show the algorithm that are to be presented later.
+Consider
+%
+\[ \boldsymbol{Ex} =
+\begin{array}{c}
+\begin{bmatrix}\label{eq:Au=f}
+b_1     & c_1           & 0         & 0 \\
+a_1     & b_2           & c_2       & 0       \\
+0       & a_2           & b_3       & c_3    \\
+0  &  0            & a_3       & b_4     \\
+\end{bmatrix}
+\begin{bmatrix}
+x_1 \\
+x_2 \\
+x_3 \\
+x_4 \\
+\end{bmatrix}
+=
+\begin{bmatrix}
+g_1 \\
+g_2 \\
+g_3 \\
+g_4
+\end{bmatrix}
+\end{array}
+= \boldsymbol{g}.
+\]
+%
+Performing Gaussian elimination we multiply the first row of $\boldsymbol{E}$ by $a_1/b_1$
+%
+\begin{gather*}
+    \frac{a_1}{b_1}\begin{bmatrix} b_1 & c_1 & 0 & 0 \end{bmatrix} =  \begin{bmatrix} a_1 & \frac{a_1 c_1}{b_1} & 0 & 0\end{bmatrix},
+\end{gather*}
+and then subtract this from the second row:
+%
+\begin{gather*}
+    \begin{bmatrix} a_1 & b_2 & c_2 & 0 \end{bmatrix} - \begin{bmatrix} a_1 & \frac{a_1 c_1}{b_1} & 0 & 0\end{bmatrix} = \begin{bmatrix} 0 & b2 - \frac{a_1 c_1}{b_1} & c_2 & 0 \end{bmatrix}.
+\end{gather*}
+%
+With these operations the matrix $\boldsymbol{E}$ now reads
+%
+\[ \boldsymbol{E} =
+\begin{array}{c}
+\begin{bmatrix}
+b_1     & c_1           & 0         & 0 \\
+0 & b2 - \frac{a_1 c_1}{b_1} & c_2 & 0      \\
+0       & a_2           & b_3       & c_3    \\
+0  &  0            & a_3       & b_4    
+\end{bmatrix},
+\end{array}
+\]
+%
+where the substitution $\Tilde{b}_2 = b_2-\frac{a_1c_1}{b_1}$ is introduced. This is performed on each of the lower triangular elements such that the end result is a upper triangular matrix
+%
+\[ \boldsymbol{E} =
+\begin{array}{c}
+\begin{bmatrix}
+b_1     & c_1           & 0         & 0 \\
+0 & \Tilde{b}_2 & c_2 & 0      \\
+0       & 0           & \tilde{b}_3       & c_3    \\
+0  &  0            & 0       & \tilde{b}_4
+\end{bmatrix},
+\end{array}
+\]
+%
+where in general 
+%
+\begin{gather}\label{eq:b_i}
+\tilde{b}_i = b_i - \frac{a_{i-1} c_{i-1}}{\tilde{b}_{i-1}},
+\end{gather}
+for $i \in [2,n+1]$.
+The above equation applies directly to $\boldsymbol{A}$.
+Since $\boldsymbol{E}$ is altered, the same operations must be applied to $\boldsymbol{g}$.
+Multiplying the first element of $\boldsymbol{g}$ by $a_1/b_1$ and subtracting this from the second element
+%
+\begin{gather*}
+    \begin{bmatrix} g_2 \end{bmatrix} - \begin{bmatrix} \frac{g_1a_1}{b_1} \end{bmatrix}  = \begin{bmatrix} g_2 - \frac{g_1a_1}{b_1}\end{bmatrix}.
+\end{gather*}
+%
+We then substitute $\tilde{g}_2 = g_2 - \frac{g_{1} a_{1}}{b_{1}}$. Applying this to $f'$ gives the general expression
+%
+\begin{gather}\label{eq:f_i}
+    \tilde{f}_i = f'_i - \frac{\tilde{f}_{i-1} a_{i-1}}{\tilde{b}_{i-1}},
+\end{gather}
+
+for $i \in [2,n+1]$.
+The values for $\tilde{b}_i$ and $\tilde{f}_i$ are calculated through forward substitution.
+For $\tilde{b}_1$ and $\tilde{f}_1$ the value is not changed from the original $b_1$ and $f'_1$ respectively.
+Therefore the calculations are done over the interval $i \in [2,n+1]$ given the aforementioned initial conditions.
+
+To calculate the values for $u_i$ we perform a backward substitution.
+We return to the example $\boldsymbol{E}\boldsymbol{x} = \boldsymbol{g}$, considering the Gauss eliminated matrix equation.
+For $i=4$ the equation reads
+%
+\begin{gather*}
+    \tilde{b}_4 x_4 = \tilde{g}_4 \Rightarrow x_4 = \frac{\tilde{g}_4}{\tilde{b}_4},
+\end{gather*}
+%
+and for $i=3$
+%
+\begin{gather*}
+    \tilde{b}_3 x_3 + a_3 x_4 = \tilde{g_3} \Rightarrow \tilde{x}_3 = \frac{\tilde{g}_3 - a_3 x_4}{\tilde{b}_3}.
+\end{gather*}
+%
+In general and applied to $\boldsymbol{A} \boldsymbol{u} = \boldsymbol{f'}$ the equation for $u_i$ reads:
+%
+\begin{gather}\label{eq:u_i}
+    u_i = \frac{\tilde{f}_i - c_i u_{i+1}}{\tilde{b}_i}.
+\end{gather}
+%
+
+\begin{lstlisting}
+%Allocate dynamic memory for variables
+double *a = new double[n+1]
+%similar for other variables b, c, f-tilde, u and x (an array of step sizes)
+
+%Define the step size as h and make an array of steps from 0 to n+2
+for i = 0; i < n+2; i++{
+    x[i] = h*double(i)
+}
+
+%Updating b-tilde and f-tilde using the equations we found using forward substitution
+for i = 2; i < n+2; i++{
+    btilde[i] = update
+    ftilde[i] = update
+}
+
+%Updating u using the equations we found using backward substitution
+for i = n; i > 0; i--{
+    u[i] = update
+}
+%write the results to a file
+
+\end{lstlisting}
+%
+
+\subsection{Specific algorithm}
+
 %\begin{bmatrix}
@@
 %\end{array}
+%
 
-\subsection{Specific algorithm}
+%
+\begin{lstlisting}
+%Allocate dynamic memory for variables, but a, b and c now have specific values
+double a = -1; double b = 2; double c = -1
+\end{lstlisting}
 
-\subsection{Armadillo}
+\begin{lstlisting}
+%Updating b and f-tilde using the equations we found using forward substitution
+for i = 2; i < n+2; i++{
+    btilde[i] = update (a and c are now constants)
+    ftilde[i] = update (a and c are now constants)
+}
+
+\end{lstlisting}
+%
+
+\subsection{LU-decomposition}
+
+For a non-singular matrix $\boldsymbol{A} \in \R ^{nxn}$ we can also preform LU-decomposition, given by:
+%
+\[ \boldsymbol{A} = \boldsymbol{Lu} =
+\begin{array}{c}
+\begin{bmatrix}\label{eq:A = Lu}
+l_{11}     & 0           & \dots         & 0       & 0 \\
+l_{21}     & l_{22}           & 0       & \dots      & 0       \\
+\dots       & \dots           & \dots       & \dots      & \dots    \\
+\dots       & \dots           & \dots       & \dots      & 0    \\
+ l_{n1} &  l_{n2}            & \dots        & l_{nn-1}      & l_{nn}     \\
+\end{bmatrix}
+\begin{bmatrix}
+u_{11}     & u_{12}            & \dots         & \dots       & u_{1n} \\
+0     & u_{22}           & \dots       & \dots      & u_{2n}       \\
+\dots       & \dots           & \dots       & \dots      & \dots    \\
+0       & \dots           & \dots       & \dots      & u_{(n-1)n}    \\
+ 0 &  0            & \dots        & 0      & u_{nn}     \\
+\end{bmatrix}
+
+\end{array}.
+\]
+%
 
@@
 Which data are used (experiment,frequencies etc.)?
-Pixel resolution ($N_{\textrm{side}}$),$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysisfor other researchers.
+Pixel resolution ($N_{\textrm{side}}$),$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis for other researchers.
 
@@
 \begin{figure}[H]
-    \centering
-    \includegraphics[size=0.25\textwidth]{p1bplot.png}
-    \caption{The solution of the differential equation using a generalized algorithm, the three different plots show the $nxn$ matrix with $n=10$, $n=100$, $n=1000$ }
-    \label{fig1: Generalized algorithm}
+\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
+\caption{The solution of the differential equation using a generalized algorithm, the three different plots show the $nxn$ matrix with $n=10$, $n=100$, $n=1000$}
+\label{fig:figure_label}
 \end{figure}
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[size=0.25\textwidth]{plot1d.png}
-    \caption{}
-    \label{fig1:1d}
+\begin{figure}[t]
+\mbox{\epsfig{figure=plot1d.png,width=\linewidth,clip=}}
+\caption{Vi burde plotte til 10^{-7}}
+\label{fig:1d}
 \end{figure}
+%
+\begin{deluxetable}{lc}[h]
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{Table of log10 of the error, for stepsizes down to $10^{-7}$}
+\tablecolumns{4}
+\tablehead{Step size & Error}
+\startdata
+ \\
+$10^{-1}$ & $10^{-1}$ \\
+$10^{-2}$ & $10^{-1}$ \\
+$10^{-3}$ & $10^{-1}$ \\
+$10^{-4}$ & $10^{-1}$ \\
+$10^{-5}$ & $10^{-1}$ \\
+$10^{-6}$ & $10^{-1}$ \\
+$10^{-7}$ & $10^{-1}$ \\
+\enddata
+\end{deluxetable}
 
+%
 \subsection{Program CPU time}
-\begin{deluxetable}{lccc}
+
+Testing the time we ran the three different programs 10 times for different sizes $n$, listing longest, shortest and average CPU times.
+In [TABLE ] the results we have listed for timing our programs with $n = 1000$. Comparably for larger matrices $n=10^{6}$ the programs yield the timings presented in [TABLE ].
+%
+\begin{deluxetable}{lccc}[h]
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000}
+\tablecolumns{4}
+\tablehead{ & General & Tailored & Armadillo}
+\startdata
+Highest & 10.8*10^{-5}s & 8.4*10^{-5}s & 0.012942s \\
+Lowest & 3.4*10^{-5}s & 2.5*10^{-5}s & 0.007715s  \\
+Average & 7.8*10^{-5}s & 5.5*10^{-5}s & 0.010166s 
+\enddata
+\end{deluxetable}
+%
+%
+\begin{deluxetable}{lccc}[h]
 %\tablewidth{0pt}
@@
 \tablecolumns{4}
-\tablehead{ & General & Taylored & Armadillo}
+\tablehead{ & General & Tailored & Armadillo}
 \startdata
-Highest & Item 2 & Item 3 & Item 4 \\
-Lowest & Item 2 & Item 3 & Item 4 \\
-Average & Item 2 & Item 3 & Item 4 
+Highest & 0.029865s & 0.028870s & Out of memory \\
+Lowest & 0.027048s & 0.027058s & Out of memory  \\
+Average & 0.027544s & 0.027602s & Out of memory 
 \enddata
 \end{deluxetable}
+%
+
+Note that the LU-decomposition runs out of memory at $n = 10^{5}$ meaning that we will not be able to test the Armadillo made algorithm for larger matrices.
 
@@
 
-\begin{deluxetable}{lccc}
+%\lstset{language=python,
+%    showspaces=false,
+%    extendedchars=true,
+%    breaklines=true,
+%    tabsize=2,
+%    basicstyle=\small\ttfamily,
+%    frame=single,
+%    columns=flexible,
+%    keepspaces=true,
+%}
+%\begin{lstlisting}
+%\end{lstlisting}
+
+%\begin{deluxetable}{lccc}
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Summary of main results.}
-\tablecolumns{4}
-\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
-\startdata
-Item 1 & Item 2 & Item 3 & Item 4
-\enddata
-\end{deluxetable}
+%\tablecaption{\label{tab:results}}
+%\tablecomments{Summary of main results.}
+%\tablecolumns{4}
+%\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
+%\startdata
+%Item 1 & Item 2 & Item 3 & Item 4
+%\enddata
+%\end{deluxetable}
 

### 2018-09-11 Project1/report/ms.tex
Updates from ShareLaTeX

@@
 \newcommand{\R}{\mathbb{R}}
+\usepackage{subfig}
 
@@
 \begin{gather}\label{eq:u(x)}
-u(x) = 1-(1-e^{10})x-e^{-10x},
+u_{cf}(x) = 1-(1-e^{10})x-e^{-10x},
 \end{gather}
@@
 \subsection{Numerical representation}
+When creating matrices and vectors in C++ they are created as dynamic arrays, so that the length of the array is decided based upon the problem. This is done to reduce the memory usage.
+
 By approximating equation \eqref{eq:du/dx} and discretizing over x we get
@@
 
+Following is pseudo code for the solution of the general algorithm.
+%
 \begin{lstlisting}
-%Allocate dynamic memory for variables
+//Allocate dynamic memory for variables
 double *a = new double[n+1]
-%similar for other variables b, c, f-tilde, u and x (an array of step sizes)
+//similar for other variables b, c, f-tilde, u and x (an array of step sizes)
 
-%Define the step size as h and make an array of steps from 0 to n+2
+//Define the step size as h and make an array of steps from 0 to n+2
 for i = 0; i < n+2; i++{
@@
 
-%Updating b-tilde and f-tilde using the equations we found using forward substitution
+//Updating b-tilde and f-tilde using the equations we found using forward substitution
 for i = 2; i < n+2; i++{
@@
 
-%Updating u using the equations we found using backward substitution
+//Updating u using the equations we found using backward substitution
 for i = n; i > 0; i--{
@@
 }
-%write the results to a file
+//write the results to a file
 
@@
 \subsection{Specific algorithm}
+Further the goal is to improve upon the general algorithm by reducing the FLOPS and making sure there are only one calculation for each FLOP.
+
+Since $a_i$, $b_i$ and $c_i$ are constant for all $i$ the value can be inserted into equations \eqref{eq:b_i}, \eqref{eq:f_i} and \eqref{eq:u_i}:
+%
+\begin{gather*}
+    \tilde{b}_i = b_i - \frac{a_{i-1}c_{i-1}}{\tilde{b}_{i-1}} = 2 - \frac{1}{\tilde{b}_{i-1}}, \\
+    \tilde{f}_i = f'_i - \frac{\tilde{f}_{i-1} a_{i-1} }{\tilde{b}_{i-1}} = f'_i + \frac{\tilde{f}_{i-1}}{\tilde{b}_{i-1}}, \\
+    u_i = \frac{\tilde{f}_i - c_i u_{i+1}}{\tilde{b}_i} = \frac{\tilde{f}_i + u_{i+1}}{\tilde{b}_i}.
+\end{gather*}
+%
+In the above equations $1/\tilde{b}_{i-1}$ is calculated twice, which results in unnecessary calculations.
+Introducing a temporary variable $b'_i = 1/\tilde{b}_{i-1}$ changes the above equations for $\tilde{b}_i$ and $\tilde{f}_i$ to
+%
+\begin{gather*}
+    \tilde{b}_i = 2 - b'_{i-1}, \\
+    \tilde{f}_i = f'_i + b'_{i-1}\tilde{f}_{i-1}.
+\end{gather*}
+%
+
+Having reduced the FLOPS to minimize the CPU time the calculations are timed to compare with the general algorithm.
+Timing is done with a built in function within C++ and only ran over the calculations of $\boldsymbol{A}\boldsymbol{u} = \tilde{f}$.
+
+\subsection{Relative error}
+The relative error is calculated by
+%
+\begin{gather}
+    \epsilon_i = \log_{10}\left(\left|\frac{u_i - u_{cf,i}}{u_{cf,i}}\right|\right).
+\end{gather}
+%
+When computing the relative error the indices are reduced such that division by zero is avoided, that is when $u_{cf,i} \rightarrow 0$. Following is pseudo code:
+\begin{lstlisting}
+eps_max //calculating the initial value, assuming this is largest
+for i=n/10+1:(9*n/10)+1 //ignoring end points
+    calculate epsilon
+    if epsilon > eps_max
+        eps_max = epsilon
+\end{lstlisting}
 
@@
 \begin{lstlisting}
-%Allocate dynamic memory for variables, but a, b and c now have specific values
+//Allocate dynamic memory for variables, but a, b and c now have specific values
 double a = -1; double b = 2; double c = -1
@@
 \begin{lstlisting}
-%Updating b and f-tilde using the equations we found using forward substitution
+//Updating b and f-tilde using the equations we found using forward substitution
 for i = 2; i < n+2; i++{
@@
 \subsection{LU-decomposition}
+Comparing the results with a completely general method we solve $\boldsymbol{A}\boldsymbol{u} = \boldsymbol{f}$ with by LU decomposition with armadillo using a solve function.
 
@@
 \section{Results}
+
+%  6.61149997e-02   8.16507165e-04   8.31668016e-06   8.33105497e-08  1.43558860e-09   8.39730640e-07   2.74865353e-06]
+
 \label{sec:results}
@@
 \begin{figure}[H]
-\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
-\caption{The solution of the differential equation using a generalized algorithm, the three different plots show the $nxn$ matrix with $n=10$, $n=100$, $n=1000$}
-\label{fig:figure_label}
+\subfloat[The resolution of our step size is low and as a result there is a visible deviation from the analytical solution.]{{\includegraphics[scale=0.37]{plot1b_n10.png}}}
+\qquad
+\subfloat[Lowering the step size renders the deviation from the analytical solution invisible in the plot.]{{\includegraphics[scale=0.37]{plot1b_n100.png}}}\qquad
+\subfloat[Further lowering the step size the only visible difference is in the resolution of the curve in our plot.]{{\includegraphics[scale=0.37]{plot1b_n1000.png}}}
+\caption{The solution of the differential equation using a generalized algorithm at matrix sizes (a): $n = 10$, (b): $n = 100$ and (c): $n = 1000$.}
+\label{fig:fig1}
 \end{figure}
@@
 \mbox{\epsfig{figure=plot1d.png,width=\linewidth,clip=}}
-\caption{Vi burde plotte til 10^{-7}}
+\caption{Logarithmic plot of the error as function of the step size, to illustrate the $\mathcal{O}(h^{2})$ dependency we have plotted $log_{10}(h^{2})=2log_{10}(h)$, which in the plot doubles as $y = 2x$, the expected value of our error.}
 \label{fig:1d}
@@
 %
-\begin{deluxetable}{lc}[h]
+\begin{deluxetable}{lcc}[H]
 %\tablewidth{0pt}
 \tablecaption{\label{tab:results}}
-\tablecomments{Table of log10 of the error, for stepsizes down to $10^{-7}$}
+\tablecomments{Table of the error, for stepsizes down to $h = 10^{-7}$.
+The table reveals that the numerical error starts increasing after $h = 10^{-5}$, and before reaching that point we the the $\mathcal{O}(h^{2})$.}
 \tablecolumns{4}
-\tablehead{Step size & Error}
+\tablehead{ Step size & Error &}
 \startdata
  \\
-$10^{-1}$ & $10^{-1}$ \\
-$10^{-2}$ & $10^{-1}$ \\
-$10^{-3}$ & $10^{-1}$ \\
-$10^{-4}$ & $10^{-1}$ \\
-$10^{-5}$ & $10^{-1}$ \\
-$10^{-6}$ & $10^{-1}$ \\
-$10^{-7}$ & $10^{-1}$ \\
+ $10^{-1}$ & $6.6114*10^{-2}$ &\\
+ $10^{-2}$ & $8.1651*10^{-4}$ &\\
+ $10^{-3}$ & $8.3167*10^{-6}$ &\\
+ $10^{-4}$ & $8.3311*10^{-8}$ &\\
+ $10^{-5}$ & $1.4356*10^{-9}$ &\\
+ $10^{-6}$ & $8.3973*10^{-7}$ &\\
+ $10^{-7}$ & $2.7487*10^{-6}$ &\\
 \enddata
@@
 %
-\begin{deluxetable}{lccc}[h]
+\begin{deluxetable}{lccc}[H]
 %\tablewidth{0pt}
 \tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000, armadillo being overall, while our general and tailored algorithm timings both withing the same order of magnitude.}
 \tablecolumns{4}
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 10^{6}}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at $n = 10^{6}$, armadillo runs out of memory, while the general and tailored algorithms staying within the same order of magnitude.}
 \tablecolumns{4}
@@
 Note that the LU-decomposition runs out of memory at $n = 10^{5}$ meaning that we will not be able to test the Armadillo made algorithm for larger matrices.
+While the tailored and general algorithms are of same order of magnitude and even similar in size (on average) for $n = 10^{7}$, we see a clear difference for the smaller $n = 1000$.
+Our general algorithm runs at $8n$ FLOPS, while our specialized algorithm only runs at $6n$ FLOPS.
 

### 2018-09-11 Project1/p1.cpp
Update p1.cpp

@@ -10,9 +10,11 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
+        // The solution for f, including the step length f squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {
+        // Closed form solution
         return 1.0 - (1.0 - exp(-10.0))*x - exp(-10.0*x);
 }
 
@@ -24,23 +26,24 @@ int main(int argc, char *argv[]){
                 exit(1);
         }
         else{
+                // reading filename and value of n from command line
                 filename = argv[1];
-                n = atoi(argv[2]);
+                n = atoi(argv[2]);      // setting n from ascii to integer
         }
         double h = 1.0/(n+1); // Step size
 
         double*x = new double [n+2]; // x_i's
 
-        // Vectors forming the tridiagonal matrix. The upper and lower diagonal are n-1 long
+        // Dynamic arrays forming the tridiagonal matrix. The upper and lower diagonal are n+1 long
         double*a = new double [n+1];
         double*b = new double [n+2];
         double*c = new double [n+1];
 
-        // Vectors for calculations
+        // Dynamical arrays for calculations
         double*u = new double [n+2];
         double*bt = new double [n+2]; // b tilde, diagonal elements (NOT RHS of equation)
         double*ft = new double [n+2]; // f tilde (RHS of equation)
-        double*fprime = new double [n+2];
+        double*fprime = new double [n+2];       // the function h^2*f = f'
 
         clock_t start, finish;
         start = clock();
@@ -54,6 +57,7 @@ int main(int argc, char *argv[]){
         }
         //loop to update
         for(int i = 0; i < n+2; i++) {
+                // Updating x_i's and storing values for h^2*f in fprime array
                 x[i] = i*h;
                 fprime[i] = f(x[i],h);
         }
@@ -61,22 +65,26 @@ int main(int argc, char *argv[]){
         // Boundary conditions
         u[0] = 0;
         u[n+1] = 0;
-        ft[0] = fprime[0];
-        ft[1] = fprime[1];
-        bt[0] = b[0];
-        bt[1] = b[0];
+        ft[0] = fprime[0];      // ftilde_0 not included in calculations
+        ft[1] = fprime[1];      // ftilde_1 = h^2*f by the general equation
+        bt[0] = b[0];   // btilde_0 not included in calculations
[…]

### 2018-09-11 Project1/p1.cpp
Update p1.cpp

@@ -32,7 +32,6 @@ int main(int argc, char *argv[]){
         }
         double h = 1.0/(n+1); // Step size
 
-        double*x = new double [n+2]; // x_i's
 
         // Dynamic arrays forming the tridiagonal matrix. The upper and lower diagonal are n+1 long
         double*a = new double [n+1];
@@ -40,18 +39,21 @@ int main(int argc, char *argv[]){
         double*c = new double [n+1];
 
         // Dynamical arrays for calculations
+        double*x = new double [n+2];            // x_i's
         double*u = new double [n+2];
-        double*bt = new double [n+2]; // b tilde, diagonal elements (NOT RHS of equation)
-        double*ft = new double [n+2]; // f tilde (RHS of equation)
+        double*bt = new double [n+2];           // b tilde, diagonal elements (NOT RHS of equation)
+        double*ft = new double [n+2];           // f tilde (RHS of equation)
         double*fprime = new double [n+2];       // the function h^2*f = f'
 
         clock_t start, finish;
         start = clock();
 
         for(int i = 0; i < n+2; i++) {
+                // setting all b_i = 2
                 b[i] = 2.0;
         }
         for(int i = 0; i < n+1; i++) {
+                // setting all a_i, c_i = -1
                 a[i] = -1.0;
                 c[i] = -1.0;
         }
@@ -68,7 +70,7 @@ int main(int argc, char *argv[]){
         ft[0] = fprime[0];      // ftilde_0 not included in calculations
         ft[1] = fprime[1];      // ftilde_1 = h^2*f by the general equation
         bt[0] = b[0];   // btilde_0 not included in calculations
-        bt[1] = b[0];   // btilde_1 = -2 by the general equation
+        bt[1] = b[0];   // btilde_1 = 2 by the general equation
 
         for(int i = 2; i < n+2; i++) {
                 // updating btilde and ftilde
@@ -83,7 +85,7 @@ int main(int argc, char *argv[]){
         finish = clock();
         cout << 1.0*(finish - start)/CLOCKS_PER_SEC << endl;
         
-        // printing to file
+        // writing to file
         outfile.open(filename);
         for(int i=0; i < n+2; i++) {
                 outfile << x[i];

### 2018-09-11 Project1/p1c.cpp
Update p1c.cpp

@@ -12,9 +12,11 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
+        // The solution for f, including the step length f squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {
+        // Closed form solution
         return 1.0 - (1.0 - exp(-10.0))*x - exp(-10.0*x);
 }
 
@@ -26,23 +28,25 @@ int main(int argc, char *argv[]){
                 exit(1);
         }
         else{
-                filename = argv[1];
-                n = atoi(argv[2]);
+                // reading filename and value of n from command line
+                filename = argv[1];     
+                n = atoi(argv[2]);      // setting n from ascii to integer
         }
         double h = 1.0/(n+1); // Step size
 
-        // Vectors for calculations
-        double*x = new double [n+2]; // x_i's
-        double*k = new double [n+2];  // (f tilde)/(b tilde)
+        // Dynamical arrays for calculations
+        double*x = new double [n+2];            // x_i's
+        double*k = new double [n+2];            // (f tilde)/(b tilde)
         double*u = new double [n+2];
-        double*bt = new double [n+2]; // b tilde, diagonal elements (NOT RHS of equation)
-        double*ft = new double [n+2]; // f tilde (RHS of equation)
-        double*fprime = new double [n+2]; // our function f multiplied by h²
+        double*bt = new double [n+2];           // b tilde, diagonal elements (NOT RHS of equation)
+        double*ft = new double [n+2];           // f tilde (RHS of equation)
+        double*fprime = new double [n+2];       // the function h^2*f = f'
 
         clock_t start, finish;
         start = clock();
-        //loop to update
+       
         for(int i = 0; i < n+2; i++) {
+                // Updating x_i's and storing values for h^2*f in fprime array
                 x[i] = i*h;
                 fprime[i] = f(x[i],h);
         }
@@ -50,24 +54,26 @@ int main(int argc, char *argv[]){
         // Boundary conditions
         u[0] = 0;
         u[n+1] = 0;
-        ft[0] = fprime[0];
-        ft[1] = fprime[1];
-        bt[0] = 2.0;  // Diagonal values
-        bt[1] = 2.0;  // Diagonal values
+        ft[0] = fprime[0];      // ftilde_0 not included in calculations
[…]

### 2018-09-11 Project1/p1d.cpp
Update p1d.cpp

@@ -11,22 +11,24 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
+        // The solution for f, including the step length f squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {
+        // Closed form solution
         return 1.0 - (1.0 - exp(-10.0))*x - exp(-10.0*x);
 }
 
 int main(int argc, char *argv[]){
         string filename; // String filename for outputfile
-        int j_max;
+        int j_max;       // max exponent of 10
         if(argc <= 1) {
                 cout << "Missing arguments:" << argv[0] << " specify output filename and max exponent of n" << endl;
                 exit(1);
         }
         else{
-                filename = argv[1];
-                j_max = atoi(argv[2]);
+                filename = argv[1];  
+                j_max = atoi(argv[2]);  
         }
         double*eps_max_arr = new double [j_max];
         double*log10h = new double [j_max];
@@ -36,16 +38,18 @@ int main(int argc, char *argv[]){
                 int n = pow(10,j);
                 double h = 1.0/(n+1); // Step size
                 log10h[j-1] = log10(h);
-                // Vectors for calculations
+                
+                // Dynamical arrays for calculations
                 double*x = new double [n+2];     // x_i's
-                double*k = new double [n+2]; // (f tilde)/(b tilde)
+                double*k = new double [n+2];    // (f tilde)/(b tilde)
                 double*u = new double [n+2];
-                double*bt = new double [n+2]; // b tilde, diagonal elements (NOT RHS of equation)
-                double*ft = new double [n+2]; // f tilde (RHS of equation)
-                double*fprime = new double [n+2]; // our function f multiplied by h²
+                double*bt = new double [n+2];   // b tilde, diagonal elements (NOT RHS of equation)
+                double*ft = new double [n+2];   // f tilde (RHS of equation)
+                double*fprime = new double [n+2]; // the function h^2*f = f'
 
                 //loop to update
                 for(int i = 0; i < n+2; i++) {
+                // Updating x_i's and storing values for h^2*f in fprime array                        
                         x[i] = i*h;
                         fprime[i] = f(x[i],h);
                 }
@@ -53,31 +57,36 @@ int main(int argc, char *argv[]){
                 // Boundary conditions
                 u[0] = 0;
                 u[n+1] = 0;
-                ft[0] = fprime[0];
-                ft[1] = fprime[1];
[…]

### 2018-09-11 Project1/p1e.cpp
Update p1e.cpp

@@ -8,15 +8,17 @@ using namespace arma;
 ofstream outfile;
 
 inline double f(double x, double h) {
+        // The solution for f, including the step length h squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {
+        // Closed form solution
         return 1.0 - (1.0 - exp(-10.0))*x - exp(-10.0*x);
 }
 
 
 int main(int argc, char*argv[]){
-        int n; // number of points
+        int n;  // Defining number of points as an integer variable
         string filename;  // String filename for outputfile
 
         if(argc <= 1) {
@@ -24,31 +26,33 @@ int main(int argc, char*argv[]){
                 exit(1);
         }
         else{
+                // reading filename and value of n from command line
                 filename = argv[1];
-                n = atoi(argv[2]);
+                n = atoi(argv[2]);      // setting n from ascii to integer
         }
         double h = 1.0/(n+1);
 
         // Matrices and Vectors
-        mat fmat = zeros<vec>(n+2);            // matrix of f
-        mat A = zeros<mat>(n+2,n+2);
-        double*x = new double [n+2];
+        mat fmat = zeros<vec>(n+2);     // matrix of f
+        mat A = zeros<mat>(n+2,n+2);    // matrix for A
+        double*x = new double [n+2];    // dynamic array for x_i
 
         for(int i = 1; i < n+2; i++) {
+                // Updating x_i's and storing values for h^2*f in fprime array
                 x[i] = i*h;
                 fmat[i] = f(x[i],h);
         }
 
-        // Changing diagonal elements of A
-        A.diag() += double(2.0);
-        A.diag(1) -= double(1.0);
-        A.diag(-1) -= double(1.0);
+        // Changing elements along the diagonal of A
+        A.diag() += double(2.0);        // center diagonal
+        A.diag(1) -= double(1.0);       // upper diagonal
+        A.diag(-1) -= double(1.0);      // lower diagonal
 
-        vec u = zeros<vec>(n+2);
-        u = solve(A,fmat);
-
+        vec u = zeros<vec>(n+2);        // defining vector for solution to LU decomposition
+        u = solve(A,fmat);              // LU decomposing
+        
+        // writing to file
         outfile.open(filename);
-        //outfile << "  x:        approx:          exact:       relative error:" << endl;
         for(int i=0; i < n+2; i++) {
                 outfile << x[i];
                 outfile << " " << u[i];

### 2018-09-11 Project1/p1c.cpp
Update p1c.cpp

@@ -12,7 +12,7 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
-        // The solution for f, including the step length f squared
+        // The solution for f, including the step length h squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {

### 2018-09-11 Project1/p1.cpp
Update p1.cpp

@@ -10,7 +10,7 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
-        // The solution for f, including the step length f squared
+        // The solution for f, including the step length h squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {

### 2018-09-11 Project1/p1d.cpp
Update p1d.cpp

@@ -11,7 +11,7 @@ using namespace std;
 ofstream outfile;
 //Functions used.
 inline double f(double x, double h) {
-        // The solution for f, including the step length f squared
+        // The solution for f, including the step length h squared
         return pow(h,2)*100.0*exp(-10*x);
 }
 inline double closed_form(double x) {

### 2018-09-11 Project1/p1e.cpp
Update p1e.cpp

@@ -1,3 +1,5 @@
+// program must be run with the following addition terminal, after the commands to compile via terminal "-DARMA_DONT_USE_WRAPPER -lblas -llapack"
+
 #include <iostream>
 #include <cmath>
 #include <armadillo>

### 2018-09-11 Project1/report/ms.tex

@@
 \documentclass{emulateapj}
-%\documentclass[12pt,preprint]{aastex}
+\documentclass[12pt,preprint]{aastex}
 
@@
 
-\title{Project 1}
+\title{FYS3150 Computational Physics, project 1}
 
@@
 \begin{abstract}
-Comparing general and tailored methods for solving second order differential equations this study will present that producing a specialized method reduces the run time of a numerical solver without losing numerical accuracy.
-Reducing the floating point operations (FLOPS) will be proven efficient for large scale calculations.
+Comparing methods for solving second order differential equations this study will present that producing a specialized method reduces the run time of a numerical solver.
+When solving a second order differential equation we find that tailoring a method to the problem at hand will result in a shorter CPU time compared to a general method similar to the specialized.
+Further comparison proves that even slightly tailoring the algorithm (by the general method) will prove to be faster than a standard LU decomposition.
+For large scale calculations the LU decomposition is found to use more memory than the computer has available, while the general and specialized methods produces results, the latter being faster.
+The specialized method will focus on reducing the floating point operations (FLOPS) by tailoring the equations prior to the numerical calculations.
 
@@
 \end{abstract}
-\keywords{computational science --- numerical methods: error estimation, run time --- methods: numerical, tridiagonal matrices}
+\keywords{computational science --- numerical methods: error estimation, LU decomposition, forward/backward substitution --- methods: analytical, numerical, statistical}
 
@@
 \label{sec:introduction}
-This study on generalized and tailored numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations.
+This study on generalized and specialized numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations.
 By comparing the run time and observing the numerical error of the calculations .
@@
 Specializing our algorithm further for the problem we will make a generalized algorithm for any solving any tridiagonal matrix.
-Lastly we will make an algorithm completely tailored to our problem where the matrix elements are already specified.
+Lastly we will make an algorithm completely specialized to our problem where the matrix elements are already specified.
 
@@
 Having reduced the FLOPS to minimize the CPU time the calculations are timed to compare with the general algorithm.
-Timing is done with a built in function within C++ and only ran over the calculations of $\boldsymbol{A}\boldsymbol{u} = \tilde{f}$.
+Timing is done with a built in function within C++ and only ran over the calculations of $\boldsymbol{A}\boldsymbol{u} = \boldsymbol{\tilde{f}}$.
+Pseudo code is presented below
+
+%
+\begin{lstlisting}
+//Allocate dynamic memory for variables, but a, b and c now have specific values
+double a = -1; double b = 2; double c = -1
+\end{lstlisting}
+
+\begin{lstlisting}
+//Updating b and f-tilde using the equations we found using forward substitution
+for i = 2; i < n+2; i++{
+    btilde[i] = update (a and c are now constants)
+    ftilde[i] = update (a and c are now constants)
+}
+
+\end{lstlisting}
+%
 
@@
 
-%
-\begin{lstlisting}
-//Allocate dynamic memory for variables, but a, b and c now have specific values
-double a = -1; double b = 2; double c = -1
-\end{lstlisting}
-
-\begin{lstlisting}
-//Updating b and f-tilde using the equations we found using forward substitution
-for i = 2; i < n+2; i++{
-    btilde[i] = update (a and c are now constants)
-    ftilde[i] = update (a and c are now constants)
-}
-
-\end{lstlisting}
-%
-
 \subsection{LU-decomposition}
-Comparing the results with a completely general method we solve $\boldsymbol{A}\boldsymbol{u} = \boldsymbol{f}$ with by LU decomposition with armadillo using a solve function.
-
-For a non-singular matrix $\boldsymbol{A} \in \R ^{nxn}$ we can also preform LU-decomposition, given by:
+Using a package from
 %
@@
 %
-
-%Summarize properties of data.
-Which data are used (experiment,frequencies etc.)?
-Pixel resolution ($N_{\textrm{side}}$),$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis for other researchers.
 
@@
 \tablecolumns{4}
-\tablehead{ Step size & Error &}
+\tablehead{ n & Error &}
 \startdata
  \\
- $10^{-1}$ & $6.6114*10^{-2}$ &\\
- $10^{-2}$ & $8.1651*10^{-4}$ &\\
- $10^{-3}$ & $8.3167*10^{-6}$ &\\
- $10^{-4}$ & $8.3311*10^{-8}$ &\\
- $10^{-5}$ & $1.4356*10^{-9}$ &\\
- $10^{-6}$ & $8.3973*10^{-7}$ &\\
- $10^{-7}$ & $2.7487*10^{-6}$ &\\
+ $10^{1}$ & $6.6114 \cdot 10^{-2}$ &\\
+ $10^{2}$ & $8.1651 \cdot 10^{-4}$ &\\
+ $10^{3}$ & $8.3167\cdot 10^{-6}$ &\\
+ $10^{4}$ & $8.3311\cdot 10^{-8}$ &\\
+ $10^{5}$ & $1.4356\cdot 10^{-9}$ &\\
+ $10^{6}$ & $8.3973\cdot 10^{-7}$ &\\
+ $10^{7}$ & $2.7487\cdot 10^{-6}$ &\\
 \enddata
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000, armadillo being overall, while our general and tailored algorithm timings both withing the same order of magnitude.}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000, armadillo being overall, while our general and specialized algorithm timings both withing the same order of magnitude.}
 \tablecolumns{4}
-\tablehead{ & General & Tailored & Armadillo}
+\tablehead{ & General & Specialized & Armadillo}
 \startdata
-Highest & 10.8*10^{-5}s & 8.4*10^{-5}s & 0.012942s \\
-Lowest & 3.4*10^{-5}s & 2.5*10^{-5}s & 0.007715s  \\
-Average & 7.8*10^{-5}s & 5.5*10^{-5}s & 0.010166s 
+Highest & 0.000333s & 0.000237s & 0.021747s \\
+Lowest & 0.000191s & 0.000074s & 0.012255s  \\
+Average & 0.000249s & 0.000167s & 0.015978s 
 \enddata
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at $n = 10^{6}$, armadillo runs out of memory, while the general and tailored algorithms staying within the same order of magnitude.}
+\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at $n = 10^{6}$, armadillo runs out of memory, while the general and specialized algorithms staying within the same order of magnitude.}
 \tablecolumns{4}
-\tablehead{ & General & Tailored & Armadillo}
+\tablehead{ & General & Specialized & Armadillo}
 \startdata
-Highest & 0.029865s & 0.028870s & Out of memory \\
-Lowest & 0.027048s & 0.027058s & Out of memory  \\
-Average & 0.027544s & 0.027602s & Out of memory 
+Highest & 0.085026 & 0.073819s & Out of memory \\
+Lowest & 0.074668 & 0.064465s & Out of memory  \\
+Average & 0.079288s & 0.067805s & Out of memory 
 \enddata
@@
 Note that the LU-decomposition runs out of memory at $n = 10^{5}$ meaning that we will not be able to test the Armadillo made algorithm for larger matrices.
-While the tailored and general algorithms are of same order of magnitude and even similar in size (on average) for $n = 10^{7}$, we see a clear difference for the smaller $n = 1000$.
-Our general algorithm runs at $8n$ FLOPS, while our specialized algorithm only runs at $6n$ FLOPS.
+Our general algorithm runs at $8n$ FLOPS, while our specialized algorithm is optimized to only run at $6n$ FLOPS.
 
@@
 \label{sec:conclusions}
+
+Reviewing the plots from [FIG:1], there is an eye catching difference in error made between $n = 10$ and $n = 100$ points.
+Which is expected of our algorithm considering the quadratic dependency of the error with respect to step size $h^{2}$.
+However between these step sizes we already reach the maximum resolution of a side by side comparison, and between $n = 100$ and $n = 1000$ the only visible change is in the resolution of the plot itself.
+aiding our understanding of the results further by computing the relative error and plotting it in a logarithmic scale we can see that our error indeed has a quadratic dependency.
+For larger step sizes however the computer starts introducing its own error
+
+While the specialized and general algorithms are of same order of magnitude and though still not similar in size (on average) for $n = 10^{6}$, we see a clear difference for the smaller $n = 1000$.
+Analyzing the number of FLOPS in the different algorithms, it becomes clear that the specialized algorithm should be $8n/6n = 3/2$ times faster.
+Comparing average CPU times we see that this is highly accurate for $n=1000$, in which case its $0.000249/0.000167 = 1.49$times faster.
+However for $n = 10^{6}$ this does not seem to be the case entirely $0.079288/0.067805 = 1.17$times faster.
+This can partially be attributed to the small sample size of ten
 
@@
 
-\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
-  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
-  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89
+\bibitem{1}
+  M. H. Jensen,
+  \textit{Project 1},
+  \url{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Projects/2018/Project1/pdf/Project1.pdf} (10.09.18)
 

### 2018-09-11 Project1/report/ms.tex
Report finished

@@
 \subsection{LU-decomposition}
-Using a package from
+Lastly we compare the previous results to a solution of the matrix equation $\boldsymbol{Au} = \boldsymbol{f}$ by an LU decomposition.
+We decompose the matrix such that $\boldsymbol{A} = \boldsymbol{L u}$, where $\boldsymbol{L}$ is a lower triangular matrix and $\boldsymbol{u}$ is an upper triangular matrix:
 %
-\[ \boldsymbol{A} = \boldsymbol{Lu} =
+\[ \boldsymbol{Lu} =
 \begin{array}{c}
 \begin{bmatrix}\label{eq:A = Lu}
-l_{11}     & 0           & \dots         & 0       & 0 \\
-l_{21}     & l_{22}           & 0       & \dots      & 0       \\
-\dots       & \dots           & \dots       & \dots      & \dots    \\
-\dots       & \dots           & \dots       & \dots      & 0    \\
- l_{n1} &  l_{n2}            & \dots        & l_{nn-1}      & l_{nn}     \\
+l_{11}     & 0                 &\dots       & 0 \\
+l_{21}     & l_{22}              & \dots      & \dots       \\
+\dots       & \dots              & \dots      & 0    \\
+ l_{n1} &  l_{n2}                   & \dots      & l_{nn}     \\
 \end{bmatrix}
 \begin{bmatrix}
-u_{11}     & u_{12}            & \dots         & \dots       & u_{1n} \\
-0     & u_{22}           & \dots       & \dots      & u_{2n}       \\
-\dots       & \dots           & \dots       & \dots      & \dots    \\
-0       & \dots           & \dots       & \dots      & u_{(n-1)n}    \\
- 0 &  0            & \dots        & 0      & u_{nn}     \\
+u_{11}     & u_{12}                  & \dots       & u_{1n} \\
+0     & u_{22}                & \dots      & u_{2n}       \\
+\dots       & \dots             & \dots      & \dots    \\
+ 0 &  \dots                 & 0      & u_{nn}     \\
 \end{bmatrix}
-
 \end{array}.
@@
 %
+
+For the LU decomposition a function "solve" from the C++ package Armadillo is used.
 
@@
 \caption{Logarithmic plot of the error as function of the step size, to illustrate the $\mathcal{O}(h^{2})$ dependency we have plotted $log_{10}(h^{2})=2log_{10}(h)$, which in the plot doubles as $y = 2x$, the expected value of our error.}
-\label{fig:1d}
+\label{fig:fig2}
 \end{figure}
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{Table of the error, for stepsizes down to $h = 10^{-7}$.
-The table reveals that the numerical error starts increasing after $h = 10^{-5}$, and before reaching that point we the the $\mathcal{O}(h^{2})$.}
+\tablecomments{Table of the error, for $n \in [10^1,10^{7}]$ .
+The table reveals that the numerical error increases when $n > 10^{5}$, while the error for $n \leq 10^{5}$ follow the expected value $\mathcal{O}(h^{2})$.}
 \tablecolumns{4}
@@
 Testing the time we ran the three different programs 10 times for different sizes $n$, listing longest, shortest and average CPU times.
-In [TABLE ] the results we have listed for timing our programs with $n = 1000$. Comparably for larger matrices $n=10^{6}$ the programs yield the timings presented in [TABLE ].
-%
+In Table \ref{tab:n1000} the results we have listed for timing our programs with $n = 1000$.
+Comparably for larger matrices $n=10^{6}$ the programs yield the timings presented in Table \ref{tab:n10^6}.
+
 \begin{deluxetable}{lccc}[H]
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
+\tablecaption{\label{tab:n1000}}
 \tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000, armadillo being overall, while our general and specialized algorithm timings both withing the same order of magnitude.}
@@
 %
-%
+
 \begin{deluxetable}{lccc}[h]
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
+\tablecaption{\label{tab:n10^6}}
 \tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at $n = 10^{6}$, armadillo runs out of memory, while the general and specialized algorithms staying within the same order of magnitude.}
@@
 \label{sec:conclusions}
+\subsection{Error}
+Reviewing the plots from Fig.
+\ref{fig:fig1} we can see that the error is greater for $n = 10$ than it is for $n = 100$ points.
+Which is expected of our algorithm considering the error is of the order $h^2$, where $h$ is the step size.
+However between these step sizes we already reach the maximum resolution of a side by side comparison, and between $n = 100$ and $n = 1000$ the only visible change is in the resolution of the plot itself.
+Further we compute the relative error of the specialized algorithm compared to the closed form solution and plot it in a logarithmic scale as seen in Fig.
+\ref{fig:fig2}.
+By Fig.
+\ref{fig:fig2} we can see that the error goes as expected until $log_{10}(h) = -5$ and the error increases.
+This is the point where our numerical model becomes unstable due to the step size being too small, called loss of numerical accuracy, which is a direct result from n being too large.
 
-Reviewing the plots from [FIG:1], there is an eye catching difference in error made between $n = 10$ and $n = 100$ points.
-Which is expected of our algorithm considering the quadratic dependency of the error with respect to step size $h^{2}$.
-However between these step sizes we already reach the maximum resolution of a side by side comparison, and between $n = 100$ and $n = 1000$ the only visible change is in the resolution of the plot itself.
-aiding our understanding of the results further by computing the relative error and plotting it in a logarithmic scale we can see that our error indeed has a quadratic dependency.
-For larger step sizes however the computer starts introducing its own error
-
-While the specialized and general algorithms are of same order of magnitude and though still not similar in size (on average) for $n = 10^{6}$, we see a clear difference for the smaller $n = 1000$.
+\subsection{CPU time}
+While CPU time of the specialized and the general algorithms are of same order of magnitude, there is a clear difference for $n = 1000$.
 Analyzing the number of FLOPS in the different algorithms, it becomes clear that the specialized algorithm should be $8n/6n = 3/2$ times faster.
 Comparing average CPU times we see that this is highly accurate for $n=1000$, in which case its $0.000249/0.000167 = 1.49$times faster.
-However for $n = 10^{6}$ this does not seem to be the case entirely $0.079288/0.067805 = 1.17$times faster.
-This can partially be attributed to the small sample size of ten
+However for $n = 10^{6}$ this does not seem to be the case entirely, as we find it to be $0.079288/0.067805 = 1.17$ times faster.
+This can partially be attributed to the small sample size and the imprecise method when performing the calculations.
+The testing environment for calculating was a laptop, we expect that for large n the calculations take more time and there is a higher load on the CPU.
+This load results in a higher temperature which on a standard laptop could cause the performance to drop.
 
-%Summarize results.
-Discuss their importance, referring to the discovery to the initial seeds for structure formation.
-Mention that these results are in good agreement with expectations from inflationary theory.
+\subsection{LU decomposition}
+The LU decomposition is as expected a slower and more demanding algorithm.
+Since the algorithm stores the variables as matrices instead of vectors, the computer runs out of memory for larger n i.e $n > 10^{5}$.
+The reason that the LU decomposition runs out of memory is that it saves every value of the matrix, including the zeros on the lower diagonal $\boldsymbol{L}$ and the upper diagonal of $u$ which takes up a lot of memory when storing the $n\times n$ matrices.
+The number of FLOPS using LU decomposition is proportional to $n^{2}$ since we are dealing with $nxn$ matrices.
+This has a considerable effect and even for smaller n, we get a slower CPU time, due to the large number of FLOPS required in our calculations.
+
+\subsection{Comments}
+We both found this project to be great as an introduction to FYS3150 Computational Physics and as a way to start learning C++.
+We did not have many critiques, other than the fact that in section 1b the variables written in the matrix and the ones in the vector $b$ and $\tilde{b}$ was a bit unclear at first glance, and it is suggested that $\tilde{b}$ in the matrix $\boldsymbol{A}$ is renamed to $\tilde{f}$.
 

### 2018-09-28 Project2/2b.cpp
First draft of project 2 finished

@@ -14,7 +14,7 @@ void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
 
         // writing to file
         ofstream outfile;
-        outfile.open("2b_results.txt");
+        outfile.open(filename);
         for (int i = 0; i < n; i++) {
                 outfile << eigvals_jac[i];
                 outfile << " " << eigvals_arm[i];
@@ -24,7 +24,17 @@ void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
 }
 
 main(int argc, char *argv[]) {
-        int n = atoi(argv[1]); // size of matrix
+        string filename;
+        int n;
+        if(argc <= 1) {
+                cout << "Missing arguments:" << argv[0] << " specify output filename and value of n" << endl;
+                exit(1);
+        }
+        else{
+                // reading filename and value of n from command line
+                filename = argv[1];
+                n = atoi(argv[2]);      // setting n from ascii to integer
+        }
 
         double rho_0 = 0.0;
         double rho_n = 1.0;
@@ -39,6 +49,8 @@ main(int argc, char *argv[]) {
         A.diag(-1) += a; // lower diagonal
 
         mat R;  // matrix of eigenvectors
+
+        write_to_file(A,R,a,d,h,n);
 
         return 0;
 }

### 2018-09-28 Project2/2e.cpp

@@ -7,20 +7,15 @@
 using namespace std;
 using namespace arma;
 
-void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
-        vec eigvals_arm = eigvals_arma(n, h, a, d); // eigenvalues from Armadillo
+void write_to_file(mat &A, mat &R, double a, double d, double h, int n, double w_r){
+        //vec eigvals_arm = eigvals_arma(n, h, a, d); // eigenvalues from Armadillo
         vec eigvals_jac = jacobi(A, R, n, h);
 
         // writing to file
         ofstream outfile;
-        outfile.open("2d_results.txt");
+        outfile.open("2e_results.txt");
         for (int i = 0; i < n; i++) {
-                if(i < 4) {
-                        outfile << eigvals_jac[i] << endl;;
-                }
-                else{
-                        exit(1);
-                }
+                outfile << eigvals_jac[i] << endl;
         }
         outfile.close();
 }
@@ -28,18 +23,22 @@ void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
 main(int argc, char *argv[]) {
         string filename;
         int n;
+        double w_r, rho_n;
+
         if(argc <= 1) {
-                cout << "Missing arguments:" << argv[0] << " specify output filename and value of n" << endl;
+                cout << "Missing arguments:" << argv[0] << " specify output filename, value of n and omega_r" << endl;
                 exit(1);
         }
         else{
                 // reading filename and value of n from command line
                 filename = argv[1];
-                n = atoi(argv[2]); // setting n from ascii to integer
+                n = atoi(argv[1]); // size of matrix
+                w_r = atof(argv[2]); // angular frequency
+                rho_n = atof(argv[3]);
         }
 
         double rho_0 = 0.0;
-        double rho_n = 5.0;
+        //double rho_n = 10.0;
         double h = (rho_n - rho_0)/double(n);               // step length, preserving u(L) = 1
         double a = -1.0/pow(h,2);
         double d = 2.0/pow(h,2);
@@ -48,13 +47,13 @@ main(int argc, char *argv[]) {
 
         // Creating tridiagonal matrix
         mat A = zeros<mat>(n, n);        // matrix for A
-        A.diag() += d + pow(rho,2); // center diagonal
+        A.diag() += d + pow(w_r,2)*pow(rho,2) + 1.0/rho; // center diagonal
[…]

### 2018-09-28 Project2/eigvals.cpp

@@ -84,7 +84,8 @@ vec jacobi(mat &A, mat &R, int n, double h) {
                 iter++;
         }
         cout << "Number of iterations: " << iter << "\n";
-        vec eigvals = sort(A.diag()); // eigenvalues are the diagonal of A
+        vec eigvals = A.diag(); // eigenvalues are the diagonal of A
+        eigvals = sort(eigvals); // sorting eigenvalues
         return eigvals;
 }
 

### 2018-09-28 Project2/plot2e.py

@@ -1,0 +1,14 @@
+import matplotlib.pyplot as plt
+import sys
+
+infile = sys.argv[1]
+
+eigvals = []
+
+for lines in open(infile, "r"):
+    col = lines.split()
+    eigvals.append(float(col[0]))
+
+
+plt.plot(eigvals)
+plt.show()

### 2018-09-28 Project2/Project2.tex

--- 2018-09-11 Project1/report/ms.tex
@@
 \usepackage{epsfig,floatflt}
-\usepackage{fourier}
-\usepackage{listings}
-\newcommand{\R}{\mathbb{R}}
-\usepackage{subfig}
-
-\lstset{language=python,
-    showspaces=false,
-    extendedchars=true,
-    breaklines=true,
-    tabsize=2,
-    basicstyle=\small\ttfamily,
-    frame=single,
-    columns=flexible,
-    keepspaces=true,
-    }
-
-\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
-\SetMathAlphabet{\mathcal}{bold}{OMS}{cmsy}{b}{n}
 
@@
 
-\title{Project 1}
+\title{A spectacular title}
 
@@
 
-\altaffiltext{1}{Institute of physics, University of
+\altaffiltext{1}{Institute of Physics, University of
   Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}
 
-%\date{Received - / Accepted -}
-
 \begin{abstract}
-Comparing general and tailored methods for solving second order differential equations this study will present that producing a specialized method reduces the run time of a numerical solver without losing numerical accuracy.
-Reducing the floating point operations (FLOPS) will be proven efficient for large scale calculations.
-
-  %State problem. Briefly describe method and data. Summarize main results.
 \end{abstract}
-\keywords{computational science --- numerical methods: error estimation, run time --- methods: numerical, tridiagonal matrices}
+\keywords{cosmic microwave background --- cosmology: observations --- methods: statistical}
 
@@
 \label{sec:introduction}
-This study on generalized and tailored numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations.
-By comparing the run time and observing the numerical error of the calculations .
-To specify, the equation in question is the one dimensional Poisson equation:
-%
-\begin{equation*}
-    {\nabla^{2}} \Phi = -4\pi \rho(r),
-\end{equation*}
-%
-where $\Phi$ is the electrostatic potential generated by a localized charge distribution and $\rho$ is the charge density.
-This equation is used in electromagnetism, but in this case it was only necessary to analyze a normalized equation on the same form.
-Our normalized one dimensional Poisson equation can then be written as:
-%
-\begin{equation}\label{eq:-u''}
-    -\frac{du^2}{dx^2} = f(x).
-\end{equation}
-
-Further rewriting this as a set of linear equation will allow us to solve it using our numerical algorithms.
-As generalized algorithm we will be using a pre-built function for LU-decomposition.
-Specializing our algorithm further for the problem we will make a generalized algorithm for any solving any tridiagonal matrix.
-Lastly we will make an algorithm completely tailored to our problem where the matrix elements are already specified.
-
-%Discuss background, physical importance and possibly some history of
-%the problem that is being studied in this paper.
 
@@
 \label{sec:method}
-For the study we consider a solution for the source term $f(x)$ to be
-%
-\begin{gather*}\label{eq:f(x)}
-    f(x) = 100e^{-10x},
-\end{gather*}
-%
-which has a closed form solution 
-%
-\begin{gather}\label{eq:u(x)}
-u_{cf}(x) = 1-(1-e^{10})x-e^{-10x},
-\end{gather}
-%
-which yields $-f(x)$ when taking the second derivative.
 
-We consider the one dimensional Poisson equation to uphold Dirichlet boundary conditions; $u(0)=u(1)=0$, so $x \in [0,1]$. The second order derivative $u(x)$ can be represented as
-%
-\begin{gather}\label{eq:du/dx}
- -\frac{du^2}{dx^2} = -\left(\frac{u(x+h) + u(x-h) - 2u(x)}{h^2} + \mathcal{O}(h^2)\right),
-\end{gather}
-%
-where h is the step size.
-
-\subsection{Numerical representation}
-When creating matrices and vectors in C++ they are created as dynamic arrays, so that the length of the array is decided based upon the problem. This is done to reduce the memory usage.
-
-By approximating equation \eqref{eq:du/dx} and discretizing over x we get
-%
-\begin{gather*}
-    -\frac{du^2}{dx^2} = -\frac{u_{i+1} + u_{i-1} - 2u_i}{h^2} = f(x_i) = f_i.
-\end{gather*}
-%
-Where the step size $h = 1/(n+1)$. When calculating $x_i = i\times h$ the calculations runs over $1$ to $n+1$, such that the last
-%
-\begin{gather}
-    x_{n+1} = (n+1)h = (n+1)/(n+1) = 1,
-\end{gather}
-%
-which ensures that $x \in [0,1]$.
-To simplify the calculations and reduce the FLOPS, we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
-%
-\begin{gather}\label{eq:u_discretized}
-    -\left(u_{i-1} - 2u_i + u_{i+1}\right) = f'_i.
-\end{gather}
-%
-The linear combination above can be represented by a matrix vector multiplication.
-The matrix will consist of the coefficients of the $u_i$'s along the central, upper and lower diagonal, making up a tridiagonal matrix.
-The coefficient, after taking the outer sign of the left hand side of equation \eqref{eq:u_discretized} into consideration, are -1, 2 and -1 respectively.
-The vector will consist of the values of the $u_i$'s.
-The matrix representation will therefore be on the following form:
-%
-\[ \boldsymbol{Au} =
-\begin{array}{c}
-\begin{bmatrix}\label{eq:Au=f}
-b_1     & c_1           & 0         & \dots     & \dots     & 0 \\
-a_1     & b_2           & c_2       & 0         & \dots     & \dots \\
-0       & a_2           & b_3       & c_3       & 0     & \dots\\
-\dots  &  0            & a_3       & b_4       & \dots    & 0 \\
-\dots  & \dots        & 0 & \dots    & \dots    & c_{n-1} \\
-0       & \dots         & \dots         & 0         & a_{n-1}   & b_n   \\
-\end{bmatrix}
-\begin{bmatrix}
-u_1 \\
-u_2 \\
-\dots \\
-\dots \\
-\dots \\
-u_n
-\end{bmatrix}
-=
-\begin{bmatrix}
-f'_1 \\
-f'_2 \\
-\dots \\
-\dots \\
-\dots \\
-f'_n
-\end{bmatrix}
-\end{array}
-= \boldsymbol{f'_i}.
-\]
-%
-Since we are considering fixed end points for $u$ we do not include $i = 0$ and $i = n+1$ in the calculations for updating the $u_i$'s.
-
-\subsection{General algorithm}
-We consider the matrix $\boldsymbol{A}$ to consist of three vectors $\boldsymbol{a}$, $\boldsymbol{b}$ and $\boldsymbol{c}$ consisting of all $a_i$, $b_i$ and $c_i$ respectively.
-To solve the matrix equation we perform a forward and a backward substitution of $\boldsymbol{A}$ and $\boldsymbol{f'}$ respectively.
-
-In order to forward substitute $\boldsymbol{A}$, the matrix must be upper triangular, which is done by Gaussian elimination.
-An example with a $4x4$-matrix $\boldsymbol{E}$ and two $4x1$ vectors $x$ and $g$ follows to show the algorithm that are to be presented later.
-Consider
-%
-\[ \boldsymbol{Ex} =
-\begin{array}{c}
-\begin{bmatrix}\label{eq:Au=f}
-b_1     & c_1           & 0         & 0 \\
-a_1     & b_2           & c_2       & 0       \\
-0       & a_2           & b_3       & c_3    \\
-0  &  0            & a_3       & b_4     \\
-\end{bmatrix}
-\begin{bmatrix}
-x_1 \\
-x_2 \\
-x_3 \\
-x_4 \\
-\end{bmatrix}
-=
-\begin{bmatrix}
-g_1 \\
-g_2 \\
-g_3 \\
-g_4
-\end{bmatrix}
-\end{array}
-= \boldsymbol{g}.
-\]
-%
-Performing Gaussian elimination we multiply the first row of $\boldsymbol{E}$ by $a_1/b_1$
-%
-\begin{gather*}
-    \frac{a_1}{b_1}\begin{bmatrix} b_1 & c_1 & 0 & 0 \end{bmatrix} =  \begin{bmatrix} a_1 & \frac{a_1 c_1}{b_1} & 0 & 0\end{bmatrix},
-\end{gather*}
-and then subtract this from the second row:
-%
-\begin{gather*}
-    \begin{bmatrix} a_1 & b_2 & c_2 & 0 \end{bmatrix} - \begin{bmatrix} a_1 & \frac{a_1 c_1}{b_1} & 0 & 0\end{bmatrix} = \begin{bmatrix} 0 & b2 - \frac{a_1 c_1}{b_1} & c_2 & 0 \end{bmatrix}.
-\end{gather*}
-%
-With these operations the matrix $\boldsymbol{E}$ now reads
-%
-\[ \boldsymbol{E} =
-\begin{array}{c}
-\begin{bmatrix}
-b_1     & c_1           & 0         & 0 \\
-0 & b2 - \frac{a_1 c_1}{b_1} & c_2 & 0      \\
-0       & a_2           & b_3       & c_3    \\
-0  &  0            & a_3       & b_4    
-\end{bmatrix},
-\end{array}
-\]
-%
-where the substitution $\Tilde{b}_2 = b_2-\frac{a_1c_1}{b_1}$ is introduced. This is performed on each of the lower triangular elements such that the end result is a upper triangular matrix
-%
-\[ \boldsymbol{E} =
-\begin{array}{c}
-\begin{bmatrix}
-b_1     & c_1           & 0         & 0 \\
-0 & \Tilde{b}_2 & c_2 & 0      \\
-0       & 0           & \tilde{b}_3       & c_3    \\
-0  &  0            & 0       & \tilde{b}_4
-\end{bmatrix},
-\end{array}
-\]
-%
-where in general 
-%
-\begin{gather}\label{eq:b_i}
-\tilde{b}_i = b_i - \frac{a_{i-1} c_{i-1}}{\tilde{b}_{i-1}},
-\end{gather}
-for $i \in [2,n+1]$.
-The above equation applies directly to $\boldsymbol{A}$.
-Since $\boldsymbol{E}$ is altered, the same operations must be applied to $\boldsymbol{g}$.
-Multiplying the first element of $\boldsymbol{g}$ by $a_1/b_1$ and subtracting this from the second element
-%
-\begin{gather*}
-    \begin{bmatrix} g_2 \end{bmatrix} - \begin{bmatrix} \frac{g_1a_1}{b_1} \end{bmatrix}  = \begin{bmatrix} g_2 - \frac{g_1a_1}{b_1}\end{bmatrix}.
-\end{gather*}
-%
-We then substitute $\tilde{g}_2 = g_2 - \frac{g_{1} a_{1}}{b_{1}}$. Applying this to $f'$ gives the general expression
-%
-\begin{gather}\label{eq:f_i}
-    \tilde{f}_i = f'_i - \frac{\tilde{f}_{i-1} a_{i-1}}{\tilde{b}_{i-1}},
-\end{gather}
-
-for $i \in [2,n+1]$.
-The values for $\tilde{b}_i$ and $\tilde{f}_i$ are calculated through forward substitution.
-For $\tilde{b}_1$ and $\tilde{f}_1$ the value is not changed from the original $b_1$ and $f'_1$ respectively.
-Therefore the calculations are done over the interval $i \in [2,n+1]$ given the aforementioned initial conditions.
-
-To calculate the values for $u_i$ we perform a backward substitution.
-We return to the example $\boldsymbol{E}\boldsymbol{x} = \boldsymbol{g}$, considering the Gauss eliminated matrix equation.
-For $i=4$ the equation reads
-%
-\begin{gather*}
-    \tilde{b}_4 x_4 = \tilde{g}_4 \Rightarrow x_4 = \frac{\tilde{g}_4}{\tilde{b}_4},
-\end{gather*}
-%
-and for $i=3$
-%
-\begin{gather*}
-    \tilde{b}_3 x_3 + a_3 x_4 = \tilde{g_3} \Rightarrow \tilde{x}_3 = \frac{\tilde{g}_3 - a_3 x_4}{\tilde{b}_3}.
-\end{gather*}
-%
-In general and applied to $\boldsymbol{A} \boldsymbol{u} = \boldsymbol{f'}$ the equation for $u_i$ reads:
-%
-\begin{gather}\label{eq:u_i}
-    u_i = \frac{\tilde{f}_i - c_i u_{i+1}}{\tilde{b}_i}.
-\end{gather}
-%
-
-Following is pseudo code for the solution of the general algorithm.
-%
-\begin{lstlisting}
-//Allocate dynamic memory for variables
-double *a = new double[n+1]
-//similar for other variables b, c, f-tilde, u and x (an array of step sizes)
-
-//Define the step size as h and make an array of steps from 0 to n+2
-for i = 0; i < n+2; i++{
-    x[i] = h*double(i)
-}
-
-//Updating b-tilde and f-tilde using the equations we found using forward substitution
-for i = 2; i < n+2; i++{
-    btilde[i] = update
-    ftilde[i] = update
-}
-
-//Updating u using the equations we found using backward substitution
-for i = n; i > 0; i--{
-    u[i] = update
-}
-//write the results to a file
-
-\end{lstlisting}
-%
-
-\subsection{Specific algorithm}
-Further the goal is to improve upon the general algorithm by reducing the FLOPS and making sure there are only one calculation for each FLOP.
-
-Since $a_i$, $b_i$ and $c_i$ are constant for all $i$ the value can be inserted into equations \eqref{eq:b_i}, \eqref{eq:f_i} and \eqref{eq:u_i}:
-%
-\begin{gather*}
-    \tilde{b}_i = b_i - \frac{a_{i-1}c_{i-1}}{\tilde{b}_{i-1}} = 2 - \frac{1}{\tilde{b}_{i-1}}, \\
-    \tilde{f}_i = f'_i - \frac{\tilde{f}_{i-1} a_{i-1} }{\tilde{b}_{i-1}} = f'_i + \frac{\tilde{f}_{i-1}}{\tilde{b}_{i-1}}, \\
-    u_i = \frac{\tilde{f}_i - c_i u_{i+1}}{\tilde{b}_i} = \frac{\tilde{f}_i + u_{i+1}}{\tilde{b}_i}.
-\end{gather*}
-%
-In the above equations $1/\tilde{b}_{i-1}$ is calculated twice, which results in unnecessary calculations.
-Introducing a temporary variable $b'_i = 1/\tilde{b}_{i-1}$ changes the above equations for $\tilde{b}_i$ and $\tilde{f}_i$ to
-%
-\begin{gather*}
-    \tilde{b}_i = 2 - b'_{i-1}, \\
-    \tilde{f}_i = f'_i + b'_{i-1}\tilde{f}_{i-1}.
-\end{gather*}
-%
-
-Having reduced the FLOPS to minimize the CPU time the calculations are timed to compare with the general algorithm.
-Timing is done with a built in function within C++ and only ran over the calculations of $\boldsymbol{A}\boldsymbol{u} = \tilde{f}$.
-
-\subsection{Relative error}
-The relative error is calculated by
-%
-\begin{gather}
-    \epsilon_i = \log_{10}\left(\left|\frac{u_i - u_{cf,i}}{u_{cf,i}}\right|\right).
-\end{gather}
-%
-When computing the relative error the indices are reduced such that division by zero is avoided, that is when $u_{cf,i} \rightarrow 0$. Following is pseudo code:
-\begin{lstlisting}
-eps_max //calculating the initial value, assuming this is largest
-for i=n/10+1:(9*n/10)+1 //ignoring end points
-    calculate epsilon
-    if epsilon > eps_max
-        eps_max = epsilon
-\end{lstlisting}
-
-%\begin{bmatrix}
-%\Tilde{f}_1 \\
-%\Tilde{f}_2 \\
-%\dots \\
-%\dots \\
-%\dots \\
-%\Tilde{f}_n
-%\end{bmatrix}
-%\end{array}
-%
-
-%
-\begin{lstlisting}
-//Allocate dynamic memory for variables, but a, b and c now have specific values
-double a = -1; double b = 2; double c = -1
-\end{lstlisting}
-
-\begin{lstlisting}
-//Updating b and f-tilde using the equations we found using forward substitution
-for i = 2; i < n+2; i++{
-    btilde[i] = update (a and c are now constants)
-    ftilde[i] = update (a and c are now constants)
-}
-
-\end{lstlisting}
-%
-
-\subsection{LU-decomposition}
-Comparing the results with a completely general method we solve $\boldsymbol{A}\boldsymbol{u} = \boldsymbol{f}$ with by LU decomposition with armadillo using a solve function.
-
-For a non-singular matrix $\boldsymbol{A} \in \R ^{nxn}$ we can also preform LU-decomposition, given by:
-%
-\[ \boldsymbol{A} = \boldsymbol{Lu} =
-\begin{array}{c}
-\begin{bmatrix}\label{eq:A = Lu}
-l_{11}     & 0           & \dots         & 0       & 0 \\
-l_{21}     & l_{22}           & 0       & \dots      & 0       \\
-\dots       & \dots           & \dots       & \dots      & \dots    \\
-\dots       & \dots           & \dots       & \dots      & 0    \\
- l_{n1} &  l_{n2}            & \dots        & l_{nn-1}      & l_{nn}     \\
-\end{bmatrix}
-\begin{bmatrix}
-u_{11}     & u_{12}            & \dots         & \dots       & u_{1n} \\
-0     & u_{22}           & \dots       & \dots      & u_{2n}       \\
-\dots       & \dots           & \dots       & \dots      & \dots    \\
-0       & \dots           & \dots       & \dots      & u_{(n-1)n}    \\
- 0 &  0            & \dots        & 0      & u_{nn}     \\
-\end{bmatrix}
-
-\end{array}.
-\]
-%
-
-%Summarize properties of data.
-Which data are used (experiment,frequencies etc.)?
-Pixel resolution ($N_{\textrm{side}}$),$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis for other researchers.
+\section{Data}
+\label{sec:data}
 
 \section{Results}
-
-%  6.61149997e-02   8.16507165e-04   8.31668016e-06   8.33105497e-08  1.43558860e-09   8.39730640e-07   2.74865353e-06]
-
 \label{sec:results}
-\subsection{Numerical error}
-
-\begin{figure}[H]
-\subfloat[The resolution of our step size is low and as a result there is a visible deviation from the analytical solution.]{{\includegraphics[scale=0.37]{plot1b_n10.png}}}
-\qquad
-\subfloat[Lowering the step size renders the deviation from the analytical solution invisible in the plot.]{{\includegraphics[scale=0.37]{plot1b_n100.png}}}\qquad
-\subfloat[Further lowering the step size the only visible difference is in the resolution of the curve in our plot.]{{\includegraphics[scale=0.37]{plot1b_n1000.png}}}
-\caption{The solution of the differential equation using a generalized algorithm at matrix sizes (a): $n = 10$, (b): $n = 100$ and (c): $n = 1000$.}
-\label{fig:fig1}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=plot1d.png,width=\linewidth,clip=}}
-\caption{Logarithmic plot of the error as function of the step size, to illustrate the $\mathcal{O}(h^{2})$ dependency we have plotted $log_{10}(h^{2})=2log_{10}(h)$, which in the plot doubles as $y = 2x$, the expected value of our error.}
-\label{fig:1d}
-\end{figure}
-%
-\begin{deluxetable}{lcc}[H]
-%\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Table of the error, for stepsizes down to $h = 10^{-7}$.
-The table reveals that the numerical error starts increasing after $h = 10^{-5}$, and before reaching that point we the the $\mathcal{O}(h^{2})$.}
-\tablecolumns{4}
-\tablehead{ Step size & Error &}
-\startdata
- \\
- $10^{-1}$ & $6.6114*10^{-2}$ &\\
- $10^{-2}$ & $8.1651*10^{-4}$ &\\
- $10^{-3}$ & $8.3167*10^{-6}$ &\\
- $10^{-4}$ & $8.3311*10^{-8}$ &\\
- $10^{-5}$ & $1.4356*10^{-9}$ &\\
- $10^{-6}$ & $8.3973*10^{-7}$ &\\
- $10^{-7}$ & $2.7487*10^{-6}$ &\\
-\enddata
-\end{deluxetable}
-
-%
-\subsection{Program CPU time}
-
-Testing the time we ran the three different programs 10 times for different sizes $n$, listing longest, shortest and average CPU times.
-In [TABLE ] the results we have listed for timing our programs with $n = 1000$. Comparably for larger matrices $n=10^{6}$ the programs yield the timings presented in [TABLE ].
-%
-\begin{deluxetable}{lccc}[H]
-%\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at n = 1000, armadillo being overall, while our general and tailored algorithm timings both withing the same order of magnitude.}
-\tablecolumns{4}
-\tablehead{ & General & Tailored & Armadillo}
-\startdata
-Highest & 10.8*10^{-5}s & 8.4*10^{-5}s & 0.012942s \\
-Lowest & 3.4*10^{-5}s & 2.5*10^{-5}s & 0.007715s  \\
-Average & 7.8*10^{-5}s & 5.5*10^{-5}s & 0.010166s 
-\enddata
-\end{deluxetable}
-%
-%
-\begin{deluxetable}{lccc}[h]
-%\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{This table contains highest, lowest and average measured CPU times for the different algorithms used at $n = 10^{6}$, armadillo runs out of memory, while the general and tailored algorithms staying within the same order of magnitude.}
-\tablecolumns{4}
-\tablehead{ & General & Tailored & Armadillo}
-\startdata
-Highest & 0.029865s & 0.028870s & Out of memory \\
-Lowest & 0.027048s & 0.027058s & Out of memory  \\
-Average & 0.027544s & 0.027602s & Out of memory 
-\enddata
-\end{deluxetable}
-%
-
-Note that the LU-decomposition runs out of memory at $n = 10^{5}$ meaning that we will not be able to test the Armadillo made algorithm for larger matrices.
-While the tailored and general algorithms are of same order of magnitude and even similar in size (on average) for $n = 10^{7}$, we see a clear difference for the smaller $n = 1000$.
-Our general algorithm runs at $8n$ FLOPS, while our specialized algorithm only runs at $6n$ FLOPS.
 
@@
 \label{sec:conclusions}
-
-%Summarize results.
-Discuss their importance, referring to the discovery to the initial seeds for structure formation.
-Mention that these results are in good agreement with expectations from inflationary theory.
 
@@
 
-%\lstset{language=python,
-%    showspaces=false,
-%    extendedchars=true,
-%    breaklines=true,
-%    tabsize=2,
-%    basicstyle=\small\ttfamily,
-%    frame=single,
-%    columns=flexible,
-%    keepspaces=true,
-%}
-%\begin{lstlisting}
-%\end{lstlisting}
-
-%\begin{deluxetable}{lccc}
+\begin{deluxetable}{lccc}
 %\tablewidth{0pt}
-%\tablecaption{\label{tab:results}}
-%\tablecomments{Summary of main results.}
-%\tablecolumns{4}
-%\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
-%\startdata
-%Item 1 & Item 2 & Item 3 & Item 4
-%\enddata
-%\end{deluxetable}
+\tablecaption{\label{tab:results}}
+\tablecomments{Summary of main results.}
+\tablecolumns{4}
+\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
+\startdata
+Item 1 & Item 2 & Item 3 & Item 4
+\enddata
+\end{deluxetable}
 
 \begin{acknowledgements}
-
 \end{acknowledgements}
@@
 
-\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
-  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
-  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89
+\bibitem[]{}
 

### 2018-09-28 Project2/Project2.tex

@@
 \label{sec:method}
+\subsection{Buckling beam}
+Considering a beam that is fastened in both ends which moves up and down when a force F is exerted on it can be described by a differential equation
+%
+\begin{gather}
+    \gamma \frac{d²u(x)}{dx²} = -Fu(x),
+\end{gather}
+%
+where $u(x)$ is the displacement of the beam in the y-direction and $\gamma$ is a constant that is defined by the properties of the beam, such as the rigidity.
+Since the beam is fastened in both ends, the displacement $u(x) = 0$ and the sum outer forces on the beam is zero.
 

### 2018-09-29 Project1/p1.cpp
removed tex file for project 1

@@ -1,97 +1,98 @@
-#include <iostream>
-#include <cmath>
-#include <string>
-#include <iomanip>
-#include <fstream>
 #include <armadillo>
 #include <array>
+#include <cmath>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <string>
 
 using namespace std;
 ofstream outfile;
-//Functions used.
+// Functions used.
 inline double f(double x, double h) {
-        // The solution for f, including the step length h squared
-        return pow(h,2)*100.0*exp(-10*x);
+  // The solution for f, including the step length h squared
+  return pow(h, 2) * 100.0 * exp(-10 * x);
 }
 inline double closed_form(double x) {
-        // Closed form solution
-        return 1.0 - (1.0 - exp(-10.0))*x - exp(-10.0*x);
+  // Closed form solution
+  return 1.0 - (1.0 - exp(-10.0)) * x - exp(-10.0 * x);
 }
 
-int main(int argc, char *argv[]){
-        int n;  // Defining exponent as an integer variable
-        string filename;  // String filename for outputfile
-        if(argc <= 1) {
-                cout << "Missing arguments:" << argv[0] << " specify output filename and value of n" << endl;
-                exit(1);
-        }
-        else{
-                // reading filename and value of n from command line
-                filename = argv[1];
-                n = atoi(argv[2]);      // setting n from ascii to integer
-        }
-        double h = 1.0/(n+1); // Step size
+int main(int argc, char *argv[]) {
+  int n;           // Defining exponent as an integer variable
+  string filename; // String filename for outputfile
+  if (argc <= 1) {
+    cout << "Missing arguments:" << argv[0]
+         << " specify output filename and value of n" << endl;
+    exit(1);
+  } else {
+    // reading filename and value of n from command line
+    filename = argv[1];
+    n = atoi(argv[2]); // setting n from ascii to integer
+  }
+  double h = 1.0 / (n + 1); // Step size
 
+  // Dynamic arrays forming the tridiagonal matrix. The upper and lower diagonal
+  // are n+1 long
+  double *a = new double[n + 1];
+  double *b = new double[n + 2];
+  double *c = new double[n + 1];
 
[…]

### 2018-09-29 Project2/Project2.tex

--- 2018-09-28 Project2/Project2.tex
@@
 \label{sec:method}
-
-\section{Data}
-\label{sec:data}
+\subsection{Buckling beam}
+Considering a beam that is fastened in both ends which moves up and down when a force F is exerted on it can be described by a differential equation
+%
+\begin{gather}\label{eq:du2/dx2}
+    R \frac{d^2 u(x)}{dx^2} = -Fu(x),
+\end{gather}
+%
+where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity.
+Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper\ref{paper:P1}, and the sum of outer forces exerted on the beam is zero.
+Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$.
+Introducing this into Eq.
+\eqref{eq:du2/dx2} gives
+%
+\begin{gather*}
+    \frac{R}{L}\frac{d^2 u(\rho)}{d\rho^2} = - FLu(\rho)
+    \Rightarrow \frac{d^2 u(\rho)}{d\rho^2} = -\frac{FL^2}{R}u(\rho).
+\end{gather*}
+%
+Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with $\lambda$ as eigenvalues with corresponding eigenvectors.
 
@@
 
-\bibitem[]{}
+\bibitem[]{}\label{paper:P1}
 

### 2018-09-29 Project2/Project2.tex
Updates from ShareLaTeX

@@
 \usepackage{epsfig,floatflt}
+\setcitestyle{square}
+
+\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
+\SetMathAlphabet{\mathcal}{bold}{OMS}{cmsy}{b}{n}
 
@@
 where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity.
-Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper\ref{paper:P1}, and the sum of outer forces exerted on the beam is zero.
-Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$.
+Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper \cite{1}, and the sum of outer forces exerted on the beam is zero.
+Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$, where $\rho \in [0,1]$.
 Introducing this into Eq.
@@
 %
-Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with $\lambda$ as eigenvalues with corresponding eigenvectors.
+Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with discretized $\lambda$'s as eigenvalues with corresponding eigenvectors.
+The methods presented in the previous paper regarding the solution of a differential equation will be applied to this problem.
+The left hand size of the equation to solve,
+%
+\begin{gather}\label{eq:du/dx=lambda}
+    \frac{d^2 u(\rho)}{dx^2} = -\lambda u(\rho)
+\end{gather}
+%
+can be approximated as
+%
+\begin{gather*}
+    \frac{d^2 u(\rho)}{dx^2} = \frac{u(\rho + h) - 2u(\rho) + u(\rho - h)}{h^2} +  \mathcal{O}(h^2),
+\end{gather*}
+%
+where the step size $h = \left(\rho_{max} - \rho_{min}\right)/n$, which is the total length of the beam divided by the resolution of the calculations n.
+The maximum and minimum value of the dimensionless length variable $\rho$ is at the end points of the beam, such that $\rho_{max} = 1$ and $\rho_{min} = 0$.
+Following the method from the previous paper the discretized equation to solve is
+%
+\begin{gather*}
+    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} = \lambda u_i.
+\end{gather*}
+Which by the method of the previous paper can be transferred to a matrix eigenvalue problem;
+%
+\begin{equation*}
+    \begin{bmatrix} d& a & 0   & 0    & \dots  &0     & 0 \\
+                                a & d & a & 0    & \dots  &0     &0 \\
+                                0   & a & d & a  &0       &\dots & 0\\
+                                \dots  & \dots & \dots & \dots  &\dots      &\dots & \dots\\
+                                0   & \dots & \dots & \dots  &a  &d & a\\
+                                0   & \dots & \dots & \dots  &\dots       &a & d\end{bmatrix} 
+                                 \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} = \lambda \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} .
+\label{eq:matrixse} 
+\end{equation*}
+%
+Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this eigenvalue problem the analytical eigenvalues are
+%
+\begin{gather}\label{eq:eigenvals_analytical}
+    \lambda_j = d + 2a\cos\left(\frac{j\pi}{n+1}\right)
+\end{gather}
+for $j = 1,2,...,n-1$.
+
+\subsubsection{Jacobi's method for solving the buckling beam problem}
 
@@
 
-\begin{thebibliography}{}
+\begin{thebibliography}{9}
 
-\bibitem[]{}\label{paper:P1}
+\bibitem{1}
+Stig-Nicolai Foyn, Christer Dreierstad
+\textit{Project 1}
 

### 2018-09-29 Project2/report/Project2.tex
In the midst of Jacobi

--- 2018-09-29 Project2/Project2.tex
@@
 %
-Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with $\lambda$ as eigenvalues with corresponding eigenvectors.
+Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with discretized $\lambda$'s as eigenvalues with corresponding eigenvectors.
+The methods presented in the previous paper regarding the solution of a differential equation will be applied to this problem.
+The left hand size of the equation to solve,
+%
+\begin{gather}\label{eq:du/dx=lambda}
+    \frac{d^2 u(\rho)}{dx^2} = -\lambda u(\rho)
+\end{gather}
+%
+can be approximated as
+%
+\begin{gather*}
+    \frac{d^2 u(\rho)}{dx^2} = \frac{u(\rho + h) - 2u(\rho) + u(\rho - h)}{h^2} +  \mathcal{O}(h^2),
+\end{gather*}
+%
+where the step size $h = \left(\rho_{max} - \rho_{min}\right)/n$, which is the total length of the beam divided by the resolution of the calculations n.
+The maximum and minimum value of the dimensionless length variable $\rho$ is at the end points of the beam, such that $\rho_{max} = 1$ and $\rho_{min} = 0$.
+Following the method from the previous paper the discretized equation to solve is
+%
+\begin{gather*}
+    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} = \lambda u_i.
+\end{gather*}
+Which by the method of the previous paper can be transferred to a matrix eigenvalue problem;
+%
+\begin{equation*}
+    \begin{bmatrix} d& a & 0   & 0    & \dots  &0     & 0 \\
+                                a & d & a & 0    & \dots  &0     &0 \\
+                                0   & a & d & a  &0       &\dots & 0\\
+                                \dots  & \dots & \dots & \dots  &\dots      &\dots & \dots\\
+                                0   & \dots & \dots & \dots  &a  &d & a\\
+                                0   & \dots & \dots & \dots  &\dots       &a & d\end{bmatrix} 
+                                 \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} = \lambda \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} .
+\label{eq:matrixse} 
+\end{equation*}
+%
+Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this eigenvalue problem the analytical eigenvalues are
+%
+\begin{gather}\label{eq:eigenvals_analytical}
+    \lambda_j = d + 2a\cos\left(\frac{j\pi}{n+1}\right)
+\end{gather}
+for $j = 1,2,...,n-1$.
+
+\subsubsection{Jacobi's method for solving the buckling beam problem}
+Solving the eigenvalue problem with Jacobi's method means solving an equation on the form
+%
+\begin{gather}
+    \boldsymbol{S}^T\boldsymbol{A}\boldsymbol{S} = \boldsymbol{R},
+\end{gather}
+%
+where $\boldsymbol{S}$ is an orthogonal matrix, so it has the property $\boldsymbol{S}^{-1} = \boldsymbol{S}^T$.
+The resulting matrix $\boldsymbol{R}$ will consist of the eigenvectors, which together with the eigenvalues makes up the eigenpairs of the buckling beam problem.
+The matrix $\boldsymbol{S}$ is such that the elements
 

### 2018-09-29 Project2/report/Project2.tex
In the midst of Jacobi

@@
 The resulting matrix $\boldsymbol{R}$ will consist of the eigenvectors, which together with the eigenvalues makes up the eigenpairs of the buckling beam problem.
-The matrix $\boldsymbol{S}$ is such that the elements
+The matrix $\boldsymbol{S}$ is such that the elements along the diagonal are 1, and zero elsewhere, except four elements that are decided by the largest element of the off diagonal of $\boldsymbol{A}$.
+The four elements mentioned rotates $\boldsymbol{A}$ an angle $\theta$ in the Euclidean n-dimensional space.
+The largest element of $\boldsymbol{A}$ is found at a position $(k,l)$, and these indices are used to define the position of the mentioned four elements of $\boldsymbol{}$
 

### 2018-09-29 Project2/eigvals.cpp
Small change to analytical function for eigenvalues

@@ -105,7 +105,7 @@ vec eigvals_arma(int n, double h, double a, double d) {
 
 vec eigvals_analytical(int n, double a, double d) {
         // returns analytical eigenvalues of matrix A
-        vec lambda(n);
+        vec lambda = zeros<vec>(n);
         for(int j = 1; j < n+1; j++)
                 lambda[j-1] = d + 2.0*a*cos(j*M_PI/(n + 1.0));
         return lambda;

### 2018-09-29 Project2/report/Project2.tex
Still confined by Jacobi

--- 2018-09-29 Project2/report/Project2.tex
@@
 
-\subsubsection{Jacobi's method for solving the buckling beam problem}
+\subsubsection{Jacobi's method for solving an eigenvalue problem}
 Solving the eigenvalue problem with Jacobi's method means solving an equation on the form
 %
-\begin{gather}
+\begin{gather}\label{eq:B=SAS}
     \boldsymbol{S}^T\boldsymbol{A}\boldsymbol{S} = \boldsymbol{R},
@@
 The resulting matrix $\boldsymbol{R}$ will consist of the eigenvectors, which together with the eigenvalues makes up the eigenpairs of the buckling beam problem.
-The matrix $\boldsymbol{S}$ is such that the elements
+The matrix $\boldsymbol{S}$ is such that the elements along the diagonal are 1, and zero elsewhere, except four elements that are decided by the largest element of the off diagonal of $\boldsymbol{A}$.
+The largest element of $\boldsymbol{A}$ is found at a position $(k,l)$, and these indices are used to define the position of the mentioned four elements of $\boldsymbol{S}$, which are such that $s_{kk} = s_{ll} = \cos\theta$ and $s_{kl} = -s_{lk} = -\sin\theta$.
+This is considered as a rotation of an angle $\theta$ in the n-dimensional Euclidean space.
+When solving \eqref{eq:B=SAS} the resulting equations to solve are
+%
+\begin{align*}
+    &r_{ii} = a_{ii}, i \neq k, i \neq l \\
+    &r_{ik} = a_{ik}\cos\theta - a_{il}\sin\theta, i\neq k, i\neq l \\
+    &r_{il} = a_{il}\cos\theta + a_{ik}\sin\theta, i \neq k, i\neq l \\
+    &r_{kk} = a_{kk} \cos^2\theta - 2a_{kl}\cos\theta\sin\theta+ a_{ll}\sin^2\theta \\
+    &r_{ll} = a_{ll} \cos^2\theta+ 2a_{kl}\cos\theta\sin\theta a_{kk}\sin^2\theta \\
+    &r_{kl} = \left(a_{kk} - a_{ll}\right)\cos\theta\sin\theta + a_{kl}\left(\cos^2\theta - \sin^2\theta\right),
+\end{align*}
+which makes up the $\boldsymbol{R}$, which consists of the eigenvectors of $\boldsymbol{A}$.
+
+For the implimentation of Jacobi's method
 

### 2018-09-30 Project2/report/Project2.tex
Almost done with Jacobi's method

--- 2018-09-29 Project2/report/Project2.tex
@@
 \label{sec:method}
+The method presented will show how to solve an eigenvalue problem using Jacobi's method.
+The first problem is a classical one involving a beam fastened in both ends.
+Further the method is transformed to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to two electrons.
+
+\subsection{Jacobi's method for solving an eigenvalue problem}
+Solving the eigenvalue problem with Jacobi's method means solving an equation on the form
+%
+\begin{gather}\label{eq:B=SAS}
+    \bold{S}^T\bold{A}\bold{S} = \bold{B},
+\end{gather}
+%
+where $\bold{S}$ is an orthogonal matrix, so it has the property $\bold{S}^{-1} = \bold{S}^T$.
+The resulting matrix $\bold{B}$ will consist of the eigenvalues along the diagonal.
+The matrix $\bold{S}$ is such that the elements along the diagonal are 1, and zero elsewhere, except four elements which are such that their position are decided by the largest element of the off-diagonal of $\bold{A}$.
+The largest element of $\bold{A}$ is found at a position $(k,l)$, and these indices are used to define the position of the mentioned four elements of $\bold{S}$, which are such that $s_{kk} = s_{ll} = \cos\theta$ and $s_{kl} = -s_{lk} = -\sin\theta$.
+This is considered as a rotation of an angle $\theta$ in the n-dimensional Euclidean space.
+When solving \eqref{eq:B=SAS} the resulting equations to solve are
+%
+\begin{align*}
+    &b_{ii} = a_{ii}, i \neq k, i \neq l \\
+    &b_{ik} = a_{ik}c - a_{il}s, i\neq k, i\neq l \\
+    &b_{il} = a_{il}c + a_{ik}s, i \neq k, i\neq l \\
+    &b_{kk} = a_{kk} c^2 - 2a_{kl}cs+ a_{ll}s^2 \\
+    &b_{ll} = a_{ll} c^2 + 2a_{kl}cs a_{kk}s^2 \\
+    &b_{kl} = \left(a_{kk} - a_{ll}\right)cs + a_{kl}\left(c^2 - s^2\right),
+\end{align*}
+where $c = \cos\theta$ and $s = \sin\theta$.
+The main goal is to reduce the off-diagonal elements of $\bold{A}$ to zero in such a way that $\bold{B}$ has the eigenvalues along the diagonal, the reduction of the elements is done by choosing the angle $\theta$.
+We make a requirement that the off-diagonal elements of $\bold{B}$ is zero;
+%
+\begin{align*}
+    b_{kl} &= 0 = \left(a_{kk} - a_{ll}\right)cs + a_{kl} \left(c^2 - s^2\right),
+\end{align*}
+%
+which can then be written as
+%
+\begin{gather*}
+    \frac{\left(a_{ll} - a_{kk}\right)}{a_{kl}}cs = \left(c^2 - s^2\right).
+\end{gather*}
+%
+Defining $\tau = \left(a_{ll} - a_{kk}\right)/2a_{kl}$ and $\tan\theta = t = s/c$, and multiplying the equation above by $1/c^2$ we obtain a quadratic equation 
+%
+\begin{gather*}
+    t^2 + 2\tau t - 1 = 0,
+\end{gather*}
+%
+which is used to define the best fit $\theta$ to reduce the off-diagonal elements of $\bold{A}$.
+The solution to the quadratic equation is $t = -\tau \pm \sqrt{1 + \tau^2}$.
+If $\tau$ becomes large, the numerical accuracy will drop, as $t \approx -\tau \pm \tau$.
+By the following multiplication
+%
+\begin{align*}
+    t_{\pm} &= -\tau \pm \sqrt{1+\tau^2}\left(\frac{\tau \pm \sqrt{1+\tau^2}}{\tau \pm \sqrt{1+\tau^2}}\right) \\
+    &= \frac{-\tau^2 + \left(1 + \tau^2\right)}{\tau \pm \sqrt{1 + \tau^2}} = \frac{1}{\tau \pm \sqrt{1+\tau^2}}.
+\end{align*}
+%
+the issue of numerical accuracy is avoided by splitting the equation such that when $\tau < 0$ we choose the negative sign in front of the square root.
+
+\subsubsection{Numerical implementation}
+When implementing Jacobi's method numerically, $\bold{B}$ is not explicitly created, instead the elements of $\bold{A}$ are changed during each iteration.
+This allows the program to run with large matrices (large n), without taking up more memory thasn needed.
+This means the left hand side of the six previous equations are changed so that they are valid for the elements of $\bold{A}$ instead of $\bold{B}$, meaning $b_{ii} \rightarrow a_{ii}$, $b_{ik} \rightarrow a_{ik}$ and so forth.
+
+///EIGENVECTORS////
+
 \subsection{Buckling beam}
@@
 
-\subsubsection{Jacobi's method for solving the buckling beam problem}
-Solving the eigenvalue problem with Jacobi's method means solving an equation on the form
-%
-\begin{gather}
-    \boldsymbol{S}^T\boldsymbol{A}\boldsymbol{S} = \boldsymbol{R},
-\end{gather}
-%
-where $\boldsymbol{S}$ is an orthogonal matrix, so it has the property $\boldsymbol{S}^{-1} = \boldsymbol{S}^T$.
-The resulting matrix $\boldsymbol{R}$ will consist of the eigenvectors, which together with the eigenvalues makes up the eigenpairs of the buckling beam problem.
-The matrix $\boldsymbol{S}$ is such that the elements
-
 \section{Results}

### 2018-09-30 Project2/report/Project2.tex

@@
 The solution to the quadratic equation is $t = -\tau \pm \sqrt{1 + \tau^2}$.
-If $\tau$ becomes large, the numerical accuracy will drop, as $t \approx -\tau \pm \tau$.
-By the following multiplication
+If $\tau$ becomes large, the numerical accuracy will drop and $t \approx -\tau \pm \tau$.
+By the following transformation
 %
@@
 %
-the issue of numerical accuracy is avoided by splitting the equation such that when $\tau < 0$ we choose the negative sign in front of the square root.
+and by splitting the equation such that when $\tau < 0$ we choose the negative sign in front of the square root the aforementioned issue is avoided. The final equations for t are
+%
+\begin{align*}
+    t_+ = \frac{1}{\tau + \sqrt{1 + \tau^2}}, \tau \geq 0 \\
+    t_- = \frac{1}{\tau - \sqrt{1 + \tau^2}}, \tau < 0
+\end{align*}
+%
+From the definition of $t = \tan\theta$, s can be expressed as
+%
+\begin{align*}
+    s = \sin\theta = \tan\theta \cos\theta = tc,
+\end{align*}
+and c can be expressed as
+\begin{align*}
+    c &= \cos\theta = \frac{\cos\theta}{\sqrt{\cos^2\theta + \sin^2\theta}} = \frac{1}{\sqrt{1 + \tan^2\theta}}  = \frac{1}{\sqrt{1 + t^2}}.
+\end{align*}
 

### 2018-09-30 Project2/report/Project2.tex
Jacobi's method complete. Starting quantum dots
Middle of one electron

@@
 
-\title{A spectacular title}
+\title{Project 2}
 
@@
 \end{abstract}
-\keywords{cosmic microwave background --- cosmology: observations --- methods: statistical}
+\keywords{buckling beam --- quantum dots --- harmonic oscillator- and Coulomb potentials --- methods: scaling of differential equations, Jacobi's method}
 
@@
 %
-Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this eigenvalue problem the analytical eigenvalues are
+Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this specific eigenvalue problem the analytical eigenvalues are
 %
@@
 
+\subsection{Quantum dots, one electron}
+The method derived in the section for the buckling beam is general for a problem that can be described by a differential equation on the form of \eqref{eq:du/dx=lambda}.
+The method can be extended to a quantum mechanical problem, if the problem can be transformed into a differential equation as mentioned.
+The radial part of the Schr$\ddot{\mathrm{o}}$dinger equation with the orbital angular momentum number $l = 0$ is
+%
+\begin{equation*}
+    -\frac{\hbar^2}{2m}\left(\frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\right)R(r) + V(r)R(r) = ER(r),
+\end{equation*}
+%
+where $r \in [0,\infty)$, the potential $V(r)$ is the harmonic oscillator potential $V(r) = (1/2)kr^2$, where $k = mw^2$ and $\omega$ is the oscillator frequency. The allowed energies are 
+%
+\begin{equation*}
+    E_{n0} = \left(2n + \frac{3}{2}\right),
+\end{equation*}
+%
+for  n = 0, 1, 2,.... Substituting $R(r) = u(r)/r$:
+%
+\begin{equation*}
+    -\frac{\hbar^2}{2m}\frac{1}{r^2}\frac{d}{dr}r^2 \frac{d}{dr}\frac{u(r)}{r} + V(r)\frac{u(r)}{r} = E\frac{u(r)}{r}.
+\end{equation*}
+%
+Evaluating the derivatives in the above equation:
+%
+\begin{align*}
+    \frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\frac{u(r)}{r} &=  \frac{1}{r^2}\frac{d}{dr}r^2\left(\frac{\frac{du}{dr}r - u}{r^2}\right) \\
+    &= \frac{1}{r^2}\frac{d}{dr}\left(\frac{du}{dr}r - u\right) \\
+    &= \frac{1}{r^2}\left(\frac{d^2u}{dr^2}r + \frac{du}{dr} - \frac{du}{dr}\right) \\
+    &= \frac{1}{r}\frac{d^2u}{dr^2},
+\end{align*}
+%
+Inserting the above and multiplying by r yields
+%
+\begin{equation*}
+    -\frac{\hbar^2}{2m}\frac{d^2u}{dr^2} + V(r)u(r) = Eu(r).
+\end{equation*}
+%
+where the boundary conditions of $u$ are $u(0) = u(\infty) = 0$.
+Defining a dimensionless variable $\rho = r/\alpha$ transforms the potential to $V(\rho) = (1/2)k\alpha^2\rho^2$, and the Schr$\ddot{\mathrm{o}}$dinger's equation becomes
+%
+\begin{equation*}
+    -\frac{\hbar}{2m\alpha^2}\frac{d^2u(\rho)}{d\rho^2} + \frac{1}{2}k\alpha^2\rho^2u(\rho) = Eu(\rho).
+\end{equation*}
+%
+Multiplying the above equation by $2m\alpha^2 / \hbar^2$
+%
+\begin{equation*}
+    -\frac{d^2u(\rho)}{d\rho^2} + \frac{mk}{\hbar^2}\alpha^2 \rho^2u(\rho) = \frac{2m\alpha^2}{\hbar^2}
+\end{equation*}
+
 \section{Results}
@@
 
+\subsection{Jacobi's method}
+
+\subsection{Quantum dots, single electron}
+
+\subsection{Quantum dots, two electrons}
+
 \section{Conclusions}
 \label{sec:conclusions}
+
+%We have managed to reduce a 3 dimentional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyizing CM-system(?)
 

### 2018-09-30 Project2/report/Project2.tex
Adding plots
End of one e

@@
 \subsection{Quantum dots, one electron}
-The method derived in the section for the buckling beam is general for a problem that can be described by a differential equation on the form of \eqref{eq:du/dx=lambda}.
+The method derived in the section for the buckling beam is general for a problem that can be described by a differential equation on the form of Eq.
+\eqref{eq:du/dx=lambda}.
 The method can be extended to a quantum mechanical problem, if the problem can be transformed into a differential equation as mentioned.
@@
 \begin{equation*}
-    -\frac{d^2u(\rho)}{d\rho^2} + \frac{mk}{\hbar^2}\alpha^2 \rho^2u(\rho) = \frac{2m\alpha^2}{\hbar^2}
+    -\frac{d^2u(\rho)}{d\rho^2} + \frac{mk}{\hbar^2}\alpha^4 \rho^2u(\rho) = \frac{2m\alpha^2}{\hbar^2}Eu(\rho).
 \end{equation*}
+%
+Where the factor in front of $u(\rho$ on the right hand side is defined as
+%
+\begin{equation*}
+    \lambda = \frac{2m\alpha^2}{\hbar^2}E,
+\end{equation*}
+%
+which corresponds to the eigenvalues of the problem. To transform the equation to a form such that it can be solved as an eigenvalue problem, $\alpha$ is fixed such that
+%
+\begin{equation*}
+    \frac{mk}{\hbar^2}\alpha^4 = 1
+    \Rightarrow \alpha = \left(\frac{\hbar^2}{mk}\right)^{1/4}.
+\end{equation*}
+%
+The rewritten expression for the Schr$\ddot{\mathrm{o}}$dinger equation is then
+%
+\begin{equation*}
+    -\frac{d^2}{d\rho^2}u(\rho) + \rho^2u(\rho) = \lambda u(\rho),
+\end{equation*}
+%
+which has known analytical eigenvalues for the ground state (with $l=0)$ $\lambda_{n=0} = 3$ and for the two first excited states $\lambda_1 = 7, \lambda_2 = 11$.
 
@@
 
+\begin{figure}[t]
+\mbox{\epsfig{figure=iterationspersize.pdf,width=\linewidth,clip=}}
+\caption{Logarithmic plot of }
+\label{fig:fig1}
+\end{figure}
+
 \subsection{Quantum dots, single electron}
 
+\begin{figure}[t]
+\mbox{\epsfig{figure=GSEperOMGr.pdf,width=\linewidth,clip=}}
+\caption{}
+\label{fig:fig2}
+\end{figure}
+
 \subsection{Quantum dots, two electrons}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=ProbfuncperRhon.pdf,width=\linewidth,clip=}}
+\caption{Logarithmic plot of }
+\label{fig:fig3}
+\end{figure}
 

### 2018-09-30 Project2/eigvals.cpp
Adding plot to project folder

@@ -27,8 +27,8 @@ void rotate(mat &A, mat &R, int &k, int &l, int n) {
                 } else {
                         t = -1.0/(-tau + sqrt(1 + pow(tau,2))); // tau + sqrt(1 + pow(tau, 2));
                 }
-                c = 1.0 / sqrt(1 + pow(t, 2));
-                s = t * c;
+                c = 1.0/sqrt(1 + pow(t,2));
+                s = t*c;
         } else {
                 c = 1.0;
                 s = 0.0;

### 2018-09-30 Project2/report/Project2.tex
Adjusting plots
Done with the one electron quantum dot

@@
 %
-which has known analytical eigenvalues for the ground state (with $l=0)$ $\lambda_{n=0} = 3$ and for the two first excited states $\lambda_1 = 7, \lambda_2 = 11$.
+which has known analytical eigenvalues for the ground state (with $l=0)$ $\lambda_{n=0} = 3$ and for the two first excited states $\lambda_1 = 7, \lambda_2 = 11, \lambda_3 = 15$.
+
+By the same method as for the buckling beam we can discretize the Schr$\ddot{\mathrm{o}}$dinger equation:
+%
+\begin{equation*}
+    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} + \rho_i^2u_i = \lambda u_i,
+\end{equation*}
+%
+where the second term on the left comes from the $\rho_i^2 = V_i$, the harmonic oscillator potential.
+The step length is as previous $h = \left(\rho_n - \rho_0\right)/n$, where $\rho_n = \rho_{max} \rightarrow \infty$.
+In the calculations $\rho_n$ is approximated by a value that is large enough such that the wave function is zero at that point, and the physics are of no further interest.
+Evaluating the equation as a tridiagonal matrix the diagonal elements are defined as
+%
+\begin{equation*}
+    d_i = \frac{2}{h^2} + V_i,
+\end{equation*}
+%
+and the upper and lower diagonal elements as
+%
+\begin{equation*}
+    a_i = -\frac{1}{h^2},
+\end{equation*}
+%
+and the final discretized Schr$\ddot{\mathrm{o}}$dinger equation is
+%
+\begin{equation*}
+    d_iu_i + e_{i-1}u_{i-1} + e_{i+1}u_{i+1} = \lambda u_i,
+\end{equation*}
+%
+which is on the same form as the equation for the buckling beam, and the problem can be solved by the established Jacobi's method.
+
+\subsection{Quantum dots, two electrons}
 
@@
 \begin{figure}[t]
-\mbox{\epsfig{figure=iterationspersize.pdf,width=\linewidth,clip=}}
-\caption{Logarithmic plot of }
+\mbox{\epsfig{figure=iterationspersize15.pdf,width=\linewidth,clip=}}
+\caption{Logarithmic plot of the number of iteration needed as function of the matrix dimentionality n, to illustrate the quadratic dependency we have plotted $y = 2log_{10}(n)$, which in the plot is seen as the dotted $y=2x$ line.}
 \label{fig:fig1}
@@
 \begin{figure}[t]
-\mbox{\epsfig{figure=GSEperOMGr.pdf,width=\linewidth,clip=}}
-\caption{}
+\mbox{\epsfig{figure=GSEperOMGr15.pdf,width=\linewidth,clip=}}
+\caption{Plot of the ground state energy as function of $\omega_{r}$, it seemingly increases linearly with slope $a = 3.4369$, with error $da = 0.0511$}
 \label{fig:fig2}
@@
 \begin{figure}[t]
-\mbox{\epsfig{figure=ProbfuncperRhon.pdf,width=\linewidth,clip=}}
-\caption{Logarithmic plot of }
+\mbox{\epsfig{figure=ProbfuncperRhons15.pdf,width=\linewidth,clip=}}
+\caption{Plot of different probability distribution of two electrons in a harmonic oscillator potential, with and without Coulomb interaction, for different $\omega_{r}$, as function of our approximation of infinity $\rho_{n}$}
 \label{fig:fig3}

### 2018-10-01 Project2/report/Project2.tex
Starting work on quantum dots, two electron

@@
 
+Comparing the results from Jacobi's method and the analytic results from the provided in [SOURCE:Poject 2], we achieve sufficient numerical precision even for variyng sizes between $n = 5$ and $n = 400$.
+%
+\begin{deluxetable}{lcc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{Table comparing CPU-time between Armadillo's built in function (eigsym) and Jacobi's method, as visible from this data Jacobi's method increases in time spent considerably faster than that of Armadillo's function.}
+\tablecolumns{4}
+\tablehead{Size n  & Armadillo & Jacobi's method}
+\startdata
+10 & 0.000262s & 0.000155s \\
+50 & 0.000915s & 0.007585s \\
+100 & 0.005636s & 0.106488s \\
+200 & 0.018883s & 1.53787s \\
+300  & 0.04241s & 8.85018s \\
+400  & 0.085901s & 30.4565s
+\enddata
+\end{deluxetable}
+%
+%
 \begin{figure}[t]
@@
 \end{figure}
+%
+\subsection{Quantum dots, single electron}
+Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a Toepliz matrix with quantum harmonic oscillator potential term.
+For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$ .
 
-\subsection{Quantum dots, single electron}
+%
+CONSIDERING ADDING PLOTS OF INTEGRATION POINTS AND ENDPOINT ANALYSIS.
+%
+
+\subsection{Quantum dots, two electrons}
 
@@
 \end{figure}
-
-\subsection{Quantum dots, two electrons}
-
+%
+\begin{deluxetable}{lcc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{Table comparing ground stat analytic energy eigenvalues found in [SOURCE:] and numerical energy eigenvalues, the analytic eigenvalues displayed in the table of the source are halved, $\epsilon ' = 0.6250$, for $\omega_{r} = 0.25$ and $\epsilon ' = 0.1750$ for $\omega_{r} = 0.05$.}
+\tablecolumns{4}
+\tablehead{$\omega_{r}$ & Analytic $2\epsilon '$ & Numerical $E_{00}$}
+\startdata
+0.25 & 1.250 & 1.24999 \\
+0.05 & 0.350 & 0.349999
+\enddata
+\end{deluxetable}
+%
 \begin{figure}[t]
 \mbox{\epsfig{figure=ProbfuncperRhons15.pdf,width=\linewidth,clip=}}
-\caption{Plot of different probability distribution of two electrons in a harmonic oscillator potential, with and without Coulomb interaction, for different $\omega_{r}$, as function of our approximation of infinity $\rho_{n}$}
+\caption{Plot of probability distribution of the two electron quantum dot in a harmonic oscillator potential for different $\omega_{r}$, as well as for no Coulomb interaction.
+The probability is plotted as a function of the dimensionless relative particle position $\rho = \rho_n - \rho_0$.
+The maximum value of the $\rho_n$ is an approximation of an infinite value, and is selected as the value where the wave function has vanished}
 \label{fig:fig3}

### 2018-10-01 Project2/report/Project2.tex
Finishing up method and results

@@
 %
-Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this specific eigenvalue problem the analytical eigenvalues are
+Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this specific eigenvalue problem the analytical eigenvalues are, as stated in \cite{2}
 %
@@
 \subsection{Quantum dots, two electrons}
+Consider now two electrons in a harmonic oscillator potential.
+The electrons will interact through a repulsive Coulomb intereaction.
+For two non-interacting particles, the radial part of the Schr$\ddot{\mathrm{o}}$dinger equation is
+%
+\begin{equation*}
+    \left[\frac{-\hbar^2}{2m}\left[\frac{d^2}{dr_1^2} - \frac{d^2}{dr_2^2}\right] + \frac{k\left(r_1^2 + r_2^2\right)}{2}\right]u(r_1,r_2) = E^{(2)}u(r_1,r_2),
+\end{equation*}
+%
+where $r_1, r_2$ denotes the radii from the center of the harmonic oscillator potential to electron 1 and 2 respectively.
+The superscript of the energy corresponds to the energy of the two electrons, such that the energy of the single electron previously discussed would be denoted as $E^{(1)}$.
+Consider a relative coordinate $\bold{r} = \bold{r_1} - \bold{r_2}$ and a center of mass coordinate $\bold{R} = \left(\bold{r_1} + \bold{r_2}\right)/2$, which transforms the Schr$\ddot{\mathrm{o}}$dinger equation to
+%
+\begin{equation*}
+    \left[\frac{-\hbar^2}{m}\left[\frac{d^2}{dr^2} - \frac{1}{4}\frac{d^2}{dR^2}\right] + \frac{kr^2}{4} + kR^2\right]u(r,R) = E^{(2)}u(r,R)
+\end{equation*}
+%
+where the energy is now $E^{(2)} = E_r + E_R$.  Adding the Coulomb repulsion corresponds to adding a potential term
+%
+\begin{equation*}
+    V(r_1,r_2) = \frac{e^2}{4\pi \epsilon_0 |\bold{r_1} - \bold{r_2}|} = \frac{\beta e^2}{r}
+\end{equation*}
+%
+where $\epsilon_0$ is the permittivity of vacuum, and $\beta e^2 = 1.44$ eVnm.
 
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{Table comparing CPU-time between Armadillo's built in function (eigsym) and Jacobi's method, as visible from this data Jacobi's method increases in time spent considerably faster than that of Armadillo's function.}
+\tablecomments{Table comparing CPU-time between Armadillo's built in function \textbf{eigsym} and Jacobi's method, as visible from this data Jacobi's method increases in time spent considerably faster than that of Armadillo's function.}
 \tablecolumns{4}
@@
 Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a Toepliz matrix with quantum harmonic oscillator potential term.
-For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$ .
+For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$, where the analytic eigenvalue is $\lambda = 3$.
 
@@
 
+\bibitem{2}
+Morten Hjorth Jensen
+\textit{Project 2}
+\url{http://compphysics.github.io/ComputationalPhysics/doc/Projects/2018/
+Project2/pdf/Project2.pdf} (01.10.18)
+
 \end{thebibliography}

### 2018-10-01 Project2/report/Project2.tex
Added resolution plots

@@
 \begin{abstract}
+
 \end{abstract}
@@
 %
-where $r_1, r_2$ denotes the radii from the center of the harmonic oscillator potential to electron 1 and 2 respectively.
+where $r_1, r_2$ denotes the radii from the center of the harmonic oscillator potential to electron 1 and 2 respectively, and k is still $k = m\omega^2$.
 The superscript of the energy corresponds to the energy of the two electrons, such that the energy of the single electron previously discussed would be denoted as $E^{(1)}$.
@@
 %
-where the energy is now $E^{(2)} = E_r + E_R$.  Adding the Coulomb repulsion corresponds to adding a potential term
+where the energy is now $E^{(2)} = E_r + E_R$. By separation of variables $u(r,R)$ can be written as $\psi(r)\phi(R)$. Adding the Coulomb repulsion corresponds to adding a potential term
 %
@@
 %
-where $\epsilon_0$ is the permittivity of vacuum, and $\beta e^2 = 1.44$ eVnm.
+where $\epsilon_0$ is the permittivity of vacuum, and $\beta e^2 = 1.44$ eVnm. With the Coulomb potential the r-dependant part of the Schr$\ddot{\mathrm{o}}$dinger equation is given as
+%
+\begin{equation*}
+    \left[-\frac{\hbar^2}{m}\frac{d^2}{dr^2} + \frac{1}{4}kr^2 + \frac{\beta e^2}{r}\right]\psi(r) = E_r \psi(r).
+\end{equation*}
+%
+Introducing dimensionless variable for the position $\rho = r/\alpha$ as previous gives
+%
+\begin{equation*}
+    \left[-\frac{\hbar^2}{m\alpha^2}\frac{d^2}{d\rho^2} + \frac{1}{4}k\alpha^2\rho^2 + \frac{\beta e^2}{\alpha\rho}\right]\psi(\rho) = E_r\psi(\rho),
+\end{equation*}
+%
+multiplying by $m\alpha^2/\hbar^2$ yields
+%
+\begin{equation*}
+    \left[-\frac{d^2}{d\rho^2} + \frac{mk\alpha^4}{4\hbar^2}\rho^2 + \frac{m\alpha\beta e^2}{\rho\hbar^2} \right]\psi(\rho) = \frac{m\alpha^2}{\hbar^2} E_r\psi(\rho).
+\end{equation*}
+%
+Introducing variable $\omega_r$
+%
+\begin{equation*}
+    \omega_r^2 = \frac{1}{4}\frac{mk}{\hbar^2}\alpha^4,
+\end{equation*}
+and require the last term in the bracket is such that
+%
+\begin{equation*}
+    \frac{m\alpha \beta e^2}{\hbar^2} = 1,
+\end{equation*}
+%
+which corresponds to scaling the equation such that
+%
+\begin{equation*}
+     \alpha = \frac{\hbar^2}{m\beta e^2}.
+\end{equation*}
+%
+Further introducing
+%
+\begin{equation*}
+    \lambda = \frac{m\alpha^2}{\hbar^2}E_r,
+\end{equation*}
+%
+the equation then becomes
+%
+\begin{equation*}
+    -\frac{d^2}{d\rho^2}\psi(\rho) + \left(\omega_r^2\rho^2 + \frac{1}{\rho}\right)\psi(\rho) = \lambda \psi(\rho),
+\end{equation*}
+%
+where the sum in the brackets corresponds to the potential of the two electrons interacting via the Coulomb repulsion in a harmonic oscillator potential.
 
@@
 %
-CONSIDERING ADDING PLOTS OF INTEGRATION POINTS AND ENDPOINT ANALYSIS.
+\begin{figure}[H]
+\subfloat[At this matrix size $n = 80$ and interval size $\rho_{n} = 5$ the number of points per interval results in an acceptable resolution.]{{\includegraphics[scale=0.37]{Resolution1_15.pfd}}}
+\qquad
+\subfloat[Increasing the interval $\rho_{n} = 20$ while letting our matrix size stay the same significantly lowers the resolution of the plot because the number of points per interval decreases.]{{\includegraphics[scale=0.37]{Resolution2_15.png}}}\qquad
+\caption{Jacobi's method on a single quantum dot for matrix size $n = 80$ and two different intervals $\rho_n = {5}$ and $\rho_{n} = 20$}
+\label{fig:fig1}
+\end{figure}
+
 %

### 2018-10-01 Project2/report/Project2.tex

@@
 \setcitestyle{square}
+\usepackage{subfig}
 
@@
 \end{abstract}
-\keywords{buckling beam --- quantum dots --- harmonic oscillator- and Coulomb potentials --- methods: scaling of differential equations, Jacobi's method}
+\keywords{classical physics --- quantum mechanics --- electromagnetism --- methods: numerical, scaling of equations, Jacobi's method}
 
@@
 \label{sec:introduction}
+In this paper we will show how to transform problems such that it can be solved as an eigenvalue problem by Jacobi's method.
+We will start by considering a beam fastened in both ends while a force F acts on it, and further apply the method to a quantum mechanical system of an electron in a harmonic oscillator potential.
+Lastly we will be extending the method to a two electron system in a harmonic oscillator potential that interacts via Coulomb repulsion.
+By scaling the differential equation that describes the systems above, we obtain an equation that can be solved as an eigenvalue problem.
+The equation found is then discretized so that we transform the equation to matrix form, as discussed in our previous paper\cite{1}, which then allows us to apply the problem to Jacobi's method.
 
@@
 \label{sec:method}
-The method presented will show how to solve an eigenvalue problem using Jacobi's method.
+The method presented will show how to solve an eigenvalue problem using Jacobi's method and that by scaling the equations for a given problem, one can transform the problem to an eigenvalue problem on a matrix form with known analytical eigenpairs.
 The first problem is a classical one involving a beam fastened in both ends.
-Further the method is transformed to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to two electrons.
+Further the method is transformed to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to a problem with two electrons.
 
@@
 %
-where the sum in the brackets corresponds to the potential of the two electrons interacting via the Coulomb repulsion in a harmonic oscillator potential.
+where the sum in the brackets corresponds to the potential of the two electrons interacting via the Coulomb repulsion in a harmonic oscillator potential. When implementing
 
@@
 \tablecaption{\label{tab:results}}
-\tablecomments{Table comparing CPU-time between Armadillo's built in function \textbf{eigsym} and Jacobi's method, as visible from this data Jacobi's method increases in time spent considerably faster than that of Armadillo's function.}
+\tablecomments{Table comparing CPU-time between Armadillo's built in function \textbf{eig_sym} and Jacobi's method, as shown from these data Jacobi's method increases in CPU time faster than that of Armadillo's function.}
 \tablecolumns{4}
@@
 \end{deluxetable}
+%|
+Note here that Armadillo finished in less than a second of CPU-time even for $n = 400$, meanwhile CPU-time on Jacobi's method is already at 30 seconds.
 %
-%
-\begin{figure}[t]
+\begin{figure}[H]
 \mbox{\epsfig{figure=iterationspersize15.pdf,width=\linewidth,clip=}}
-\caption{Logarithmic plot of the number of iteration needed as function of the matrix dimentionality n, to illustrate the quadratic dependency we have plotted $y = 2log_{10}(n)$, which in the plot is seen as the dotted $y=2x$ line.}
+\caption{Logarithmic plot of the number of iteration needed as function of the matrix dimensionality n, to illustrate the quadratic dependency we have plotted $y = 2log_{10}(n)$, which in the plot is seen as the dotted $y=2x$ line.}
 \label{fig:fig1}
 \end{figure}
-%
+
 \subsection{Quantum dots, single electron}
-Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a Toepliz matrix with quantum harmonic oscillator potential term.
+Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a tridiagonal matrix with quantum harmonic oscillator potential term.
 For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$, where the analytic eigenvalue is $\lambda = 3$.
@@
 \begin{figure}[H]
-\subfloat[At this matrix size $n = 80$ and interval size $\rho_{n} = 5$ the number of points per interval results in an acceptable resolution.]{{\includegraphics[scale=0.37]{Resolution1_15.pfd}}}
-\qquad
-\subfloat[Increasing the interval $\rho_{n} = 20$ while letting our matrix size stay the same significantly lowers the resolution of the plot because the number of points per interval decreases.]{{\includegraphics[scale=0.37]{Resolution2_15.png}}}\qquad
-\caption{Jacobi's method on a single quantum dot for matrix size $n = 80$ and two different intervals $\rho_n = {5}$ and $\rho_{n} = 20$}
+\subfloat[At this matrix size $n = 80$ and interval size $\rho_{n} = 5$ the number of points per interval results in an acceptable resolution.]{{\includegraphics[scale=0.53]{Resolution1_15.pdf}}
+}\qquad
+\subfloat[Increasing the interval $\rho_{n} = 20$ significantly lowers the resolution of the plot because the number of points per interval decreases.]{{\includegraphics[scale=0.53]{Resolution2_15.pdf}}
+}\qquad
+\caption{Jacobi's method on a single quantum dot for matrix size $n = 80$ and two different intervals $\rho_n = {5}$ and $\rho_{n} = 20$.}
 \label{fig:fig1}
@@
 
+%\begin{figure}[H]
+%\subfloat[tekst.]{{\includegraphics[scale=0.37]{plot1b_n10.png}}%}
+%\qquad
+%\subfloat[tekst.]{{\includegraphics[scale=0.37]{plot1b_n100.png}%}}\qquad
+%\subfloat[tekst.]{{\includegraphics[scale=0.37]{plot1b_n1000.png%}}}
+%\caption{tekst.}
+%\label{fig:fig1}
+%\end{figure}
 %
@@
 
-\begin{figure}[t]
+\begin{figure}[H]
 \mbox{\epsfig{figure=GSEperOMGr15.pdf,width=\linewidth,clip=}}
@@
 %
-\begin{figure}[t]
+\begin{figure}[H]
 \mbox{\epsfig{figure=ProbfuncperRhons15.pdf,width=\linewidth,clip=}}
@@
 \label{sec:conclusions}
+
+\subsection{Jacobi's method}
+Utilizing Jacobi's method on scaled differential equations have proven to be a powerful tool for classical and quantum mechanics problems alike.
+When reviewing numerical accuracy of the method  compared to the analytic results given in \cite{2}, it gives results accurate down to at least 8 significant figures.
+This result is unique for Toepliz matrices, meaning for other tridiagonal matrices like the ones in the quantum mechanis part of this paper the numerical precision is no longer as high compared to analytical results.
+However in comparison to other methods it is quite demanding as the number of iterations increases along a quadratic dependency of $n$, the size of the matrix.
+In addition the method increases quite fast in CPU-time meaning that the method will not be very versatile if one would need a large $n$.
+Already at $n = 400$ the CPU time passes $30$ seconds, where Armadillo's method for solving an eigenvalue problem has a CPU time of less than a second.
+
+\subsection{Quantum dots, one electron}
+
+\subsection{Quantum dots, two electrons}
+
+\begin{figure}[H]
+\mbox{\epsfig{figure=Omg_r.pdf,width=\linewidth,clip=}}
+\caption{Plot of a quadratic potential scaled by $\omega_{r}$ like our harmonic oscillator potential, it is apparent that as $omega_{r}$ increases the potential becomes increasingly confined}
+\label{fig:fig6}
+\end{figure}
 
@@
 
-\begin{deluxetable}{lccc}
+%\begin{deluxetable}{lccc}
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Summary of main results.}
-\tablecolumns{4}
-\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
-\startdata
-Item 1 & Item 2 & Item 3 & Item 4
-\enddata
-\end{deluxetable}
+%\tablecaption{\label{tab:results}}
+%\tablecomments{Summary of main results.}
+%\tablecolumns{4}
+%\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
+%\startdata
+%Item 1 & Item 2 & Item 3 & Item 4
+%\enddata
+%\end{deluxetable}
 
@@
 
+\bibitem{3}
+ M. Taut
+\textit{Two electrons in an external oscillator potential: Particular analytic solutions of a Coulomb correlation problem}
+\url{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.48.3561} (01.10.18)
+
 \end{thebibliography}

### 2018-10-01 Project2/report/Project2.tex
Finishing intro

@@
 In this paper we will show how to transform problems such that it can be solved as an eigenvalue problem by Jacobi's method.
-We will start by considering a beam fastened in both ends while a force F acts on it, and further apply the method to a quantum mechanical system of an electron in a harmonic oscillator potential.
-Lastly we will be extending the method to a two electron system in a harmonic oscillator potential that interacts via Coulomb repulsion.
+We will start by considering a beam fastened in both ends with a force F acting on it, and further apply the method to a quantum mechanical system of an electron in a harmonic oscillator potential.
+Lastly we will be extending the method to a two electron system in a harmonic oscillator potential where the electrons interact via Coulomb repulsion.
 By scaling the differential equation that describes the systems above, we obtain an equation that can be solved as an eigenvalue problem.
-The equation found is then discretized so that we transform the equation to matrix form, as discussed in our previous paper\cite{1}, which then allows us to apply the problem to Jacobi's method.
+The scaled equation is then discretized and set up as a matrix, which method  was discussed in our previous paper\cite{1}, which allows us to apply Jacobi's method to solve the problem.
+We will be comparing the CPU time of Jacobi's method to Armadillo's method for solving an eigenvalue problem.
+Also comparing the calculated eigenvalues of Jacobi's method with Armadillo's function and the analytically calculated eigenvalues.
+When we extend the problem to a quantum mechanical one, the matrix no longer has identical values along the diagonal (i.e.
+no longer a Toeplitz matrix) as the harmonic osillator potential, and the Coulomb repulsion, is not constant as the electrons move in the system.
 
@@
 The first problem is a classical one involving a beam fastened in both ends.
-Further the method is transformed to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to a problem with two electrons.
+Further the method is applied to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to a problem with two electrons.
 
@@
 
-Comparing the results from Jacobi's method and the analytic results from the provided in [SOURCE:Poject 2], we achieve sufficient numerical precision even for variyng sizes between $n = 5$ and $n = 400$.
+Comparing the results from Jacobi's method and the analytic results from the provided in \cite{2}, we achieve sufficient numerical precision even for variyng sizes between $n = 5$ and $n = 400$.
 %
@@
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Table comparing CPU-time between Armadillo's built in function \textbf{eig_sym} and Jacobi's method, as shown from these data Jacobi's method increases in CPU time faster than that of Armadillo's function.}
-\tablecolumns{4}
+\tablecaption{\label{tab:results1}}
+\tablecomments{Table comparing CPU-time between Armadillo's built in function for solving an eigenvalue problem and Jacobi's method, as shown from these data Jacobi's method increases in CPU time faster than that of Armadillo's function.}
+\tablecolumns{3}
 \tablehead{Size n  & Armadillo & Jacobi's method}
@@
 
+\begin{deluxetable}{lcc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results2}}
+\tablecomments{Analytical energy eigenvalues provided in \cite{2} and the numerical ones found by Jacobi's method for $n=400$.}
+\tablecolumns{4}
+\tablehead{& Analytical energies & Numerical energies}
+\startdata
+& 3 & 2.99995 \\
+& 7 & 6.99976 \\
+& 11 & 10.9996 \\
+& 15 & 15.0043
+\enddata
+\end{deluxetable}
+
 %
@@
 \caption{Jacobi's method on a single quantum dot for matrix size $n = 80$ and two different intervals $\rho_n = {5}$ and $\rho_{n} = 20$.}
-\label{fig:fig1}
+\label{fig:fig2}
 \end{figure}
@@
 \caption{Plot of the ground state energy as function of $\omega_{r}$, it seemingly increases linearly with slope $a = 3.4369$, with error $da = 0.0511$}
-\label{fig:fig2}
+\label{fig:fig3}
 \end{figure}
@@
 %\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Table comparing ground stat analytic energy eigenvalues found in [SOURCE:] and numerical energy eigenvalues, the analytic eigenvalues displayed in the table of the source are halved, $\epsilon ' = 0.6250$, for $\omega_{r} = 0.25$ and $\epsilon ' = 0.1750$ for $\omega_{r} = 0.05$.}
+\tablecaption{\label{tab:results3}}
+\tablecomments{Table comparing ground stat analytic energy eigenvalues found in \cite{3} and numerical energy eigenvalues, the analytic eigenvalues displayed in the table of the source are halved, $\epsilon ' = 0.6250$, for $\omega_{r} = 0.25$ and $\epsilon ' = 0.1750$ for $\omega_{r} = 0.05$.}
 \tablecolumns{4}
@@
 The maximum value of the $\rho_n$ is an approximation of an infinite value, and is selected as the value where the wave function has vanished}
-\label{fig:fig3}
+\label{fig:fig4}
 \end{figure}
@@
 This result is unique for Toepliz matrices, meaning for other tridiagonal matrices like the ones in the quantum mechanis part of this paper the numerical precision is no longer as high compared to analytical results.
-However in comparison to other methods it is quite demanding as the number of iterations increases along a quadratic dependency of $n$, the size of the matrix.
+However in comparison to other methods it is quite demanding as the number of iterations increases along a quadratic dependency of $n$ [fig: \ref{fig:fig1}], the size of the matrix.
 In addition the method increases quite fast in CPU-time meaning that the method will not be very versatile if one would need a large $n$.
@@
 \caption{Plot of a quadratic potential scaled by $\omega_{r}$ like our harmonic oscillator potential, it is apparent that as $omega_{r}$ increases the potential becomes increasingly confined}
-\label{fig:fig6}
+\label{fig:fig5}
 \end{figure}

### 2018-10-01 Project2/report/Project2.tex

@@
 In this paper we will show how to transform problems such that it can be solved as an eigenvalue problem by Jacobi's method.
-We will start by considering a beam fastened in both ends with a force F acting on it, and further apply the method to a quantum mechanical system of an electron in a harmonic oscillator potential.
-Lastly we will be extending the method to a two electron system in a harmonic oscillator potential where the electrons interact via Coulomb repulsion.
+We will start by considering a beam fastened in both ends with a force F acting on it, described by
+%
+\begin{equation*}
+    R\frac{d^2u(x)}{dx^2} = -Fu(x),
+\end{equation*}
+%
+and further apply the method to a quantum mechanical system of an electron in a harmonic oscillator potential described by the radial part of the Schr$\ddot{\mathrm{o}}$dinger equation 
+%
+\begin{equation*}
+    \frac{-\hbar^2}{2m}\frac{d^2u(r)}{dr^2} + \frac{kr^2}{2}u(r) = Eu(r).
+\end{equation*}
+%
+Lastly we will be extending the method to a two electron system in a harmonic oscillator potential where the electrons interact via Coulomb repulsion, which by the radial part of the Schr$\ddot{\mathrm{o}}$dinger equation can be expressed as
+%
+\begin{equation*}
+    \left[\frac{-\hbar^2}{2m}\left[\frac{d^2}{dr_1^2} - \frac{d^2}{dr_2^2}\right] + \frac{k(r_1^2 + r_2^2)}{2}\right]u(r_1,r_2) = Eu(r_1,r_2)
+\end{equation*}
+%
+The three aforementioned equation can, by scaling, be transformed to an eigenvalue problem on as differential equation on the form
+%
+\begin{equation*}
+    -\frac{d^2u(\rho)}{d\rho^2} + C(\rho)u(\rho) = \lambda u(\rho),
+\end{equation*}
+%
+where $C(\rho)$ is some function which depends on the system at hand, which for the buckling beam problem is zero and for the quantum mechanical systems represents the potential and Coulomb repulsion.
 By scaling the differential equation that describes the systems above, we obtain an equation that can be solved as an eigenvalue problem.
@@
 no longer a Toeplitz matrix) as the harmonic osillator potential, and the Coulomb repulsion, is not constant as the electrons move in the system.
+We will also study what happens to the wave function as the strength of the harmonic oscillator potential is changed.
 
@@
 \subsection{Quantum dots, one electron}
+Studying a case of one electron by adding a quantum harmonic oscillator potential to our tridiagonal matrix.
+The numerical precision from applying Jacobi's method to Toepliz matrix, is reduced under the now more complex tridiagonal matrix.
+To achieve four leading digits it becomes necessary to increase matrix size to $n = 380$.
+This is a result of probability function calculated for the ground state energy, which now requires a stable balance between interval size $\rho$ and matrix size $n$.
+As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requi
 
@@
 \textit{Project 2}
-\url{http://compphysics.github.io/ComputationalPhysics/doc/Projects/2018/
-Project2/pdf/Project2.pdf} (01.10.18)
+\url{http://compphysics.github.io/ComputationalPhysics/doc/Projects/
+2018/Project2/pdf/Project2.pdf} (01.10.18)
 

### 2018-10-01 Project2/report/Project2.tex
Finished intro

@@
 This is a result of probability function calculated for the ground state energy, which now requires a stable balance between interval size $\rho$ and matrix size $n$.
-As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requi
+As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requires that matrix size increases along with the
 

### 2018-10-01 Project2/report/Project2.tex
Done with abstract

--- 2018-10-01 Project2/report/Project2.tex
@@
 \begin{abstract}
+In this study we will use Jacobi's method to solve eigenvalue problems.
+The problems that are solved are not eigenvalue problems to begin with, but are transformed to eigenvalue problems through scaling and methods discussed in our previous paper\cite{1}.
+When analyzing the results from Jacobi's method, we find it to be slow, but numerically accurate, when comparing the eigenvalues and CPU time to an eigenvalue solver from the C++ library Armadillo and comparing the eigenvalues with the results from the analytical solution.
+The number of iterations increase almost quadratically as $n^2$, with n being the dimension of an $n \times n$ matrix.
+With quantum mechanical potentials we find that approximations of maximum distance from the potential (allowed to be infinite) are crucial to avoid numerical errors.
 
 \end{abstract}
+
 \keywords{classical physics --- quantum mechanics --- electromagnetism --- methods: numerical, scaling of equations, Jacobi's method}
@@
 In this paper we will show how to transform problems such that it can be solved as an eigenvalue problem by Jacobi's method.
-We will start by considering a beam fastened in both ends with a force F acting on it, described by
+We will start by considering a beam fastened in both ends with a force acting on it, described by
 %
@@
 By scaling the differential equation that describes the systems above, we obtain an equation that can be solved as an eigenvalue problem.
-The scaled equation is then discretized and set up as a matrix, which method  was discussed in our previous paper\cite{1}, which allows us to apply Jacobi's method to solve the problem.
+The scaled equation is then discretized and set up as a matrix, which method  was discussed in our previous paper, which allows us to apply Jacobi's method to solve the problem.
 We will be comparing the CPU time of Jacobi's method to Armadillo's method for solving an eigenvalue problem.
@@
 The matrix $\bold{S}$ is such that the elements along the diagonal are 1, and zero elsewhere, except four elements which are such that their position are decided by the largest element of the off-diagonal of $\bold{A}$.
+For a unitary transformation the orthogonality and dot product are preserved, to show this, consider a vector $w$ and $v$ and a matrix $U$, we consider that $w_i = Uv_i$.
+We have
+%
+\begin{equation*}
+    w_i^T = \left(Uv_i\right)^T = v_i^TU^T,
+\end{equation*}
+%
+and $w_j = Uv_j$. This gives
+%
+\begin{equation*}
+    w_i^Tw_j = u_i^TU^T U v_j = v_iv_j = \delta_{ij}
+\end{equation*}
+%
+where $U^T U = I$, so as previously mentioned the dot product and orthgonality is preserved under a unitary transformation.
 The largest element of $\bold{A}$ is found at a position $(k,l)$, and these indices are used to define the position of the mentioned four elements of $\bold{S}$, which are such that $s_{kk} = s_{ll} = \cos\theta$ and $s_{kl} = -s_{lk} = -\sin\theta$.
@@
 
-///EIGENVECTORS////
-
 \subsection{Buckling beam}
@@
 where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity.
-Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper \cite{1}, and the sum of outer forces exerted on the beam is zero.
+Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper, and the sum of outer forces exerted on the beam is zero.
 Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$, where $\rho \in [0,1]$.
@@
 \subsection{Quantum dots, two electrons}
-
+%
 \begin{figure}[H]
@@
 \end{figure}
-
+%
 \section{Conclusions}
 \label{sec:conclusions}
-
+%
 \subsection{Jacobi's method}
@@
 Already at $n = 400$ the CPU time passes $30$ seconds, where Armadillo's method for solving an eigenvalue problem has a CPU time of less than a second.
-
+As conclusion to this I would recommend finding another method for larger matrices.
+%
 \subsection{Quantum dots, one electron}
@@
 The numerical precision from applying Jacobi's method to Toepliz matrix, is reduced under the now more complex tridiagonal matrix.
-To achieve four leading digits it becomes necessary to increase matrix size to $n = 380$.
-This is a result of probability function calculated for the ground state energy, which now requires a stable balance between interval size $\rho$ and matrix size $n$.
-As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requi
-
+To achieve four leading digits it becomes necessary to increase matrix size to $n = 380$, even on a reasonable interval $\rho$.
+The reduction in numerical precision is a result of probability function calculated for the ground state energy, which now requires a stable balance between interval size $\rho$ and matrix size $n$.
+As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requires that matrix size increase to preserve the points per interval (resolution).
+Meaning to avoid significant numerical errors in the eigenvalues and eigenvectors, we should select $\rho_{n}$ to where the function has for all practical purposes died down, and not much further.
+%
 \subsection{Quantum dots, two electrons}
-
+%
+By scaling the equations we have reduced a 3 dimensional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyzing relative radius between the electrons.
+Changing from harmonic oscillator potential to both harmonic oscillator- and Coulomb potential.
+From \cite{3} we have analytic values of the energy eigenvalues of the system for a selection of $\omega_{r}$.
+Using the same $\omega_{r}$, $n=400$ and appropriate values interval $\rho$ we managed to reproduce the analytical results to five digits behind the decimal point.
+Analyzing the energy eigenvalues further we also wound a seemingly linear dependency on $\omega_{r}$.
+%
 \begin{figure}[H]
 \mbox{\epsfig{figure=Omg_r.pdf,width=\linewidth,clip=}}
-\caption{Plot of a quadratic potential scaled by $\omega_{r}$ like our harmonic oscillator potential, it is apparent that as $omega_{r}$ increases the potential becomes increasingly confined}
+\caption{Plot of a quadratic potential scaled by $\omega_{r}$ like our harmonic oscillator potential, it is apparent that as $\omega_{r}$ increases the potential becomes increasingly confined}
 \label{fig:fig5}
 \end{figure}
-
-%We have managed to reduce a 3 dimentional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyizing CM-system(?)
+%
+Reviewing [fig:\ref{fig:fig4}] it is clear that increasing $\omega_{r}$ in turn makes the potential become increasingly confined as explained with [fig\ref{fig:fig5}], this is seen on the location of the probability function in the plot.
+Removing the Coulomb potential has the same effect as confining the electrons in a very strong potential since the significance between interaction with the potential and between the electrons is reduced.
+%
+\subsection{Comments}
+This has been an educational project, and to complete it we have been forced to improve in both C++ and implementation in general.
+There is also room for gaining more intuition into quantum mechanical systems.
+However while trying to write a good abstract and conclusion we find this project lacking in clear goals.
 
@@
 \textit{Project 1}
+\url{https://github.com/chdre/FYS3150/blob/master/Project1/report/
+FYS3150-Project1-stignicf-chrisdre.pdf} (01.10.18)
 

### 2018-10-02 Project2/report/Project2.tex
Lacks final statement in abstract and conclusion. Also needs proof reading

@@
 %
-Reviewing [fig:\ref{fig:fig4}] it is clear that increasing $\omega_{r}$ in turn makes the potential become increasingly confined as explained with [fig\ref{fig:fig5}], this is seen on the location of the probability function in the plot.
+Reviewing [fig:\ref{fig:fig4}] it is clear that increasing $\omega_{r}$ in turn makes the potential become increasingly confined as explained with [fig:\ref{fig:fig5}], this is seen on the location of the probability function in the plot.
 Removing the Coulomb potential has the same effect as confining the electrons in a very strong potential since the significance between interaction with the potential and between the electrons is reduced.

### 2018-10-02 Project2/2b.cpp
Added comments

@@ -8,6 +8,9 @@ using namespace std;
 using namespace arma;
 
 void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
+        /* Writing values to file. Takes the adress of matrix A and R, and the
+           values along the diagonal, upper diagonal and lower diagonal of A, theta
+           step size h and matrix dimension n.*/
         vec eigvals_arm = eigvals_arma(n, h, a, d); // eigenvalues from Armadillo
         vec eigvals_ana = eigvals_analytical(n, a, d); // analytical eigenvalues
         vec eigvals_jac = jacobi(A, R, n, h);
@@ -16,6 +19,7 @@ void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
         ofstream outfile;
         outfile.open(filename);
         for (int i = 0; i < n; i++) {
+                // writing all eigenvalues to file
                 outfile << eigvals_jac[i];
                 outfile << " " << eigvals_arm[i];
                 outfile << " " << eigvals_ana[i] << endl;
@@ -36,9 +40,9 @@ main(int argc, char *argv[]) {
                 n = atoi(argv[2]);      // setting n from ascii to integer
         }
 
-        double rho_0 = 0.0;
-        double rho_n = 1.0;
-        double h = (rho_n - rho_0)/double(n);               // step length, preserving u(L) = 1
+        double rho_0 = 0.0;   // start value of dimensionless length variable
+        double rho_n = 1.0;   // end value of dimensionless length variable
+        double h = (rho_n - rho_0)/double(n);  // step length
         double a = -1.0/pow(h,2);
         double d = 2.0/pow(h,2);
 

### 2018-10-02 Project2/2d.cpp

@@ -8,14 +8,18 @@ using namespace std;
 using namespace arma;
 
 void write_to_file(mat &A, mat &R, double a, double d, double h, int n){
+        /* Writing values to file. Takes the adress of matrix A and R, and the
+           values along the diagonal, upper diagonal and lower diagonal of A, theta
+           step size h and matrix dimension n.*/
         vec eigvals_arm = eigvals_arma(n, h, a, d); // eigenvalues from Armadillo
-        vec eigvals_jac = jacobi(A, R, n, h);
+        vec eigvals_jac = jacobi(A, R, n, h);       // eigenvalues from Jacobi
 
         // writing to file
         ofstream outfile;
         outfile.open("2d_results.txt");
         for (int i = 0; i < n; i++) {
                 if(i < 4) {
+                        // writing the first 4 eigenvalues to file
                         outfile << eigvals_jac[i] << endl;;
                 }
                 else{
@@ -38,13 +42,13 @@ main(int argc, char *argv[]) {
                 n = atoi(argv[2]); // setting n from ascii to integer
         }
 
-        double rho_0 = 0.0;
-        double rho_n = 5.0;
-        double h = (rho_n - rho_0)/double(n);               // step length, preserving u(L) = 1
+        double rho_0 = 0.0;     // start value of dimensionless length variable
+        double rho_n = 5.0;     // end value of dimensionless length variable
+        double h = (rho_n - rho_0)/double(n);  // step length
         double a = -1.0/pow(h,2);
         double d = 2.0/pow(h,2);
 
-        vec rho = rho_0 + linspace<vec>(1,n,n)*h;
+        vec rho = rho_0 + linspace<vec>(1,n,n)*h; // rho_i's
 
         // Creating tridiagonal matrix
         mat A = zeros<mat>(n, n);        // matrix for A

### 2018-10-02 Project2/2e.cpp

@@ -8,13 +8,16 @@ using namespace std;
 using namespace arma;
 
 void write_to_file(mat &A, mat &R, double a, double d, double h, int n, double w_r){
-        //vec eigvals_arm = eigvals_arma(n, h, a, d); // eigenvalues from Armadillo
+        /* Writing values to file. Takes the adress of matrix A and R, and the
+           values along the diagonal, upper diagonal and lower diagonal of A, theta
+           step size h and matrix dimension n.*/
         vec eigvals_jac = jacobi(A, R, n, h);
 
         // writing to file
         ofstream outfile;
         outfile.open("2e_results.txt");
         for (int i = 0; i < n; i++) {
+                // writing eigenvalues of the jacobis method
                 outfile << eigvals_jac[i] << endl;
         }
         outfile.close();
@@ -30,16 +33,15 @@ main(int argc, char *argv[]) {
                 exit(1);
         }
         else{
-                // reading filename and value of n from command line
+                // reading filename, value of n, omega_r and rho_n from command line
                 filename = argv[1];
                 n = atoi(argv[1]); // size of matrix
                 w_r = atof(argv[2]); // angular frequency
-                rho_n = atof(argv[3]);
+                rho_n = atof(argv[3]);  // end value of rho
         }
 
-        double rho_0 = 0.0;
-        //double rho_n = 10.0;
-        double h = (rho_n - rho_0)/double(n);               // step length, preserving u(L) = 1
+        double rho_0 = 0.0; // start value of dimensionless length variable
+        double h = (rho_n - rho_0)/double(n);  // step length
         double a = -1.0/pow(h,2);
         double d = 2.0/pow(h,2);
 

### 2018-10-02 Project2/eigvals.cpp

@@ -1,8 +1,8 @@
 #include "eigvals.h"
 
 double max_offdiag(mat &A, int n, int *l, int *k) {
-        // function to find the maximum value of the array A
-        double maxelm = 0.0;
+        // function to find the maximum value of the array A.
+        double maxelm = 0.0;  // temporary value of the max element
 
         for (int i = 0; i < n; i++) {
                 for (int j = i + 1; j < n; j++) {
@@ -29,7 +29,7 @@ void rotate(mat &A, mat &R, int &k, int &l, int n) {
                 }
                 c = 1.0/sqrt(1 + pow(t,2));
                 s = t*c;
-        } else {
+        } else {  // setting the angle theta to 90 degrees
                 c = 1.0;
                 s = 0.0;
         }
@@ -66,34 +66,38 @@ void rotate(mat &A, mat &R, int &k, int &l, int n) {
 }
 
 vec jacobi(mat &A, mat &R, int n, double h) {
-        int k, l;      // indices for largest off diagonal element
-        double eps = 1.0e-8; // tolerance
+        /* Jacobi's method for finding the eigenvalues of A. Function takes the
+           adresse of A and R, the matrix dimension n and the step size h. Returns
+           the eigenvalues (sorted by size) of A. */
+        int k, l;             // indices for largest off diagonal element
+        double eps = 1.0e-8;  // tolerance
 
-        R = zeros<mat>(n,n); // eigenvector matrix
+        R = zeros<mat>(n,n);  // eigenvector matrix
         R.diag() += double(1.0);
 
         double max_offdiagval = max_offdiag(A, n, &l, &k); // max offdiag element
         double max_iter = pow(double(n),3);         // max number of iterations
-        int iter = 0;                                // counter for iterations
+        int iter = 0;                               // counter for iterations
 
-        // creating a while loop that checks whether the off diagonal elements are
-        // larger than eps. Calling the function max_offdiag to find.
         while (abs(max_offdiagval) > eps && iter < max_iter) {
+                /*checking whether the off diagonal elements are larger than eps.
+                   Calling the function max_offdiag to find .*/
                 max_offdiagval = max_offdiag(A, n, &l, &k);
                 rotate(A, R, k, l, n);
                 iter++;
         }
         cout << "Number of iterations: " << iter << "\n";
-        vec eigvals = A.diag(); // eigenvalues are the diagonal of A
[…]

### 2018-10-02 Project2/report/Project2.tex
Report finished

--- 2018-10-01 Project2/report/Project2.tex
@@
 With quantum mechanical potentials we find that approximations of maximum distance from the potential (allowed to be infinite) are crucial to avoid numerical errors.
+In addition confining electrons in an increasingly larger potential has the same effect as omitting Coulomb interaction.
 
@@
 \end{deluxetable}
-%|
-Note here that Armadillo finished in less than a second of CPU-time even for $n = 400$, meanwhile CPU-time on Jacobi's method is already at 30 seconds.
 %
@@
 \subsection{Quantum dots, single electron}
-Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a tridiagonal matrix with quantum harmonic oscillator potential term.
-For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$, where the analytic eigenvalue is $\lambda = 3$.
+When analyzing how to achieve stable results by varying the integration points $n$ and the endpoint $\rho_{n}$, we found that for $\rho_{n} = 5$, the size n required to achieve four leading digits for the eigenvalue of the ground state is $n = 380$.
+This returns the value $E_{00} = 2.99995$, where the analytic eigenvalue is $\lambda = 3$.
+However $n=400$ provided (see Table \ref{tab:results2}) results that gives values closer to the analytical ones, and is more practical for testing purposes.
 
@@
 \tablecaption{\label{tab:results2}}
-\tablecomments{Analytical energy eigenvalues provided in \cite{2} and the numerical ones found by Jacobi's method for $n=400$.}
+\tablecomments{Analytical energy eigenvalues given in by the paper by M. H. Jensen\cite{2} and the numerical ones calculated by Jacobi's method for $n=400$.}
 \tablecolumns{4}
-\tablehead{& Analytical energies & Numerical energies}
+\tablehead{& Analytical  & Numerical }
 \startdata
@@
 \mbox{\epsfig{figure=GSEperOMGr15.pdf,width=\linewidth,clip=}}
-\caption{Plot of the ground state energy as function of $\omega_{r}$, it seemingly increases linearly with slope $a = 3.4369$, with error $da = 0.0511$}
+\caption{Plot of the ground state energy as function of $\omega_{r}$, it seemingly increases linearly with slope $a = 3.4369$, with error $da = 0.0511$.
+The slope is found by a matlab software (found in the github as vekt.m).}
 \label{fig:fig3}
@@
 \tablecaption{\label{tab:results3}}
-\tablecomments{Table comparing ground stat analytic energy eigenvalues found in \cite{3} and numerical energy eigenvalues, the analytic eigenvalues displayed in the table of the source are halved, $\epsilon ' = 0.6250$, for $\omega_{r} = 0.25$ and $\epsilon ' = 0.1750$ for $\omega_{r} = 0.05$.}
+\tablecomments{Table comparing the ground state analytic energy eigenvalues as provided in M.
+Taut's paper\cite{3} and the numerical energy eigenvalues found by Jacobi's method.
+The analytic eigenvalues displayed in the table of provided by M.
+Taut are halved.}
 \tablecolumns{4}
@@
 The probability is plotted as a function of the dimensionless relative particle position $\rho = \rho_n - \rho_0$.
-The maximum value of the $\rho_n$ is an approximation of an infinite value, and is selected as the value where the wave function has vanished}
+The maximum value of the $\rho_n$ is an approximation of an infinite value, and is selected as the value where the wave function has vanished.}
 \label{fig:fig4}
@@
 \subsection{Jacobi's method}
-Utilizing Jacobi's method on scaled differential equations have proven to be a powerful tool for classical and quantum mechanics problems alike.
-When reviewing numerical accuracy of the method  compared to the analytic results given in \cite{2}, it gives results accurate down to at least 8 significant figures.
-This result is unique for Toepliz matrices, meaning for other tridiagonal matrices like the ones in the quantum mechanis part of this paper the numerical precision is no longer as high compared to analytical results.
-However in comparison to other methods it is quite demanding as the number of iterations increases along a quadratic dependency of $n$ [fig: \ref{fig:fig1}], the size of the matrix.
-In addition the method increases quite fast in CPU-time meaning that the method will not be very versatile if one would need a large $n$.
+Utilizing Jacobi's method on scaled differential equations tranformed to an eigenvalue problem have proven to be a powerful tool for classical and quantum mechanics problems alike.
+When reviewing numerical accuracy of the method  compared to the analytic eigenvalues given in \cite{2}, it gives results accurate down to at least 8 significant figures.
+This result is unique for Toeplitz matrices, meaning for other tridiagonal matrices like the ones in the quantum mechanis part of this paper the numerical precision is no longer as high compared to analytical results.
+However in comparison to other methods it is quite demanding as the number of iterations increases along a quadratic dependency of $n$ (see Fig.
+\ref{fig:fig1}), the size of the matrix.
+In addition the CPU time of the method increases fast, meaning that the method will not be very usable if one would need a large $n$.
 Already at $n = 400$ the CPU time passes $30$ seconds, where Armadillo's method for solving an eigenvalue problem has a CPU time of less than a second.
-As conclusion to this I would recommend finding another method for larger matrices.
+For problems that require large matrices Jacobi's method is insufficient considering is high CPU time.
 %
 \subsection{Quantum dots, one electron}
-Studying a case of one electron by adding a quantum harmonic oscillator potential to our tridiagonal matrix.
-The numerical precision from applying Jacobi's method to Toepliz matrix, is reduced under the now more complex tridiagonal matrix.
-To achieve four leading digits it becomes necessary to increase matrix size to $n = 380$, even on a reasonable interval $\rho$.
-The reduction in numerical precision is a result of probability function calculated for the ground state energy, which now requires a stable balance between interval size $\rho$ and matrix size $n$.
-As seen in [fig:\ref{fig:fig2}] the interval size $\rho$ governs our approximation of infinity in the potential, however letting this interval increase requires that matrix size increase to preserve the points per interval (resolution).
-Meaning to avoid significant numerical errors in the eigenvalues and eigenvectors, we should select $\rho_{n}$ to where the function has for all practical purposes died down, and not much further.
+Studying a case of one electron by adding a quantum harmonic oscillator potential to the diagonal of the tridiagonal matrix.
+The numerical precision, compared to that of Jacobi's method on a Toeplitz matrix, is reduced as the matrix is more complex (the diagonal is no longer constant).
+To achieve four leading digits it becomes necessary to increase matrix size n, to a minimum of $n = 380$, even on a reasonable interval $\rho$.
+The reduction in numerical precision is a result of the potential added.
+The potential requires a stable balance between interval size $\rho$ and matrix size $n$, to achieve a good resolution of the probability function of the ground state energy.
+As seen in Fig.\ref{fig:fig2} the interval size $\rho$ decides the approximation of infinity in the potential, however letting this interval increase requires that the size of the matrix increase to preserve the resolution (points per interval).
+Meaning that to avoid significant numerical errors in the calculations of the eigenvalues and the eigenvectors, the interval end $\rho_{n}$ is required to be at the $\rho$ where the wave functions has collapsed.
+If $\rho_{n}$ is further out resolution will be sacrificed for diminishing returns in physical insight of our system.
 %
@@
 %
-By scaling the equations we have reduced a 3 dimensional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyzing relative radius between the electrons.
-Changing from harmonic oscillator potential to both harmonic oscillator- and Coulomb potential.
+By scaling the equations we have reduced a 3 dimensional two body problem to a analytically solvable problem.
+This is done by changing coordinate system from euclidean to spherical symmetric, and analyzing relative distance between the electrons.
 From \cite{3} we have analytic values of the energy eigenvalues of the system for a selection of $\omega_{r}$.
-Using the same $\omega_{r}$, $n=400$ and appropriate values interval $\rho$ we managed to reproduce the analytical results to five digits behind the decimal point.
-Analyzing the energy eigenvalues further we also wound a seemingly linear dependency on $\omega_{r}$.
+Using the $\omega_r$ in \ref{tab:results3}, $n=400$ and appropriate values interval $\rho$ we managed to reproduce the analytical results to five digits behind the decimal point.
+Analyzing the energy eigenvalues further we also wound a seemingly linear dependency on $\omega_{r}$ as seen in Fig.\ref{fig:fig3}.
 %
@@
 \mbox{\epsfig{figure=Omg_r.pdf,width=\linewidth,clip=}}
-\caption{Plot of a quadratic potential scaled by $\omega_{r}$ like our harmonic oscillator potential, it is apparent that as $\omega_{r}$ increases the potential becomes increasingly confined}
+\caption{Plot of a quadratic potential scaled by $\omega_{r}$ like the harmonic oscillator potential, it is apparent that as $\omega_{r}$ increases the potential becomes increasingly confined.}
 \label{fig:fig5}
@@
 %
-Reviewing [fig:\ref{fig:fig4}] it is clear that increasing $\omega_{r}$ in turn makes the potential become increasingly confined as explained with [fig\ref{fig:fig5}], this is seen on the location of the probability function in the plot.
+Reviewing Fig.\ref{fig:fig4} it is clear that increasing $\omega_{r}$ in turn makes the potential become increasingly confined as explained with Fig.\ref{fig:fig5}, this is seen on the location of the probability function in the plot.
 Removing the Coulomb potential has the same effect as confining the electrons in a very strong potential since the significance between interaction with the potential and between the electrons is reduced.

### 2018-10-22 Project3/main.cpp
Plots for various masses of Jupiter

@@ -13,8 +13,8 @@ int main(){
         //beta = atof(argv[1]);
         //vfac = atof(argv[2]);    // factor to change the velocity
 
-        Time = 10.0;    // time [years]
-        n = 1000;    // steps
+        Time = 7.5;    // time [years]
+        n = 100000;    // steps
         timestep = Time/n;
 
         // constants
@@ -24,7 +24,7 @@ int main(){
         M_sunval = 2e30;          // mass of sun
         M_sun = 1.0;              // relative mass of sun
         M_e = 6.0e24/M_sunval;    // relative mass of earth
-        M_j = 1.9e27/M_sunval;    // relative mass of jupiter
+        M_j = 1000*1.9e27/M_sunval;    // relative mass of jupiter
 
         planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
         planet Earth(M_e, 1, 0, 0, 0, 2.0*M_PI, 0);

### 2018-10-22 Project3/pplot.py

@@ -10,7 +10,9 @@ if data == "momentum":
 
 
     plt.figure()
+    plt.title("Momentum of Earth-Sun system using Velocity Verlet")
     plt.plot(time,angularMom)
+    plt.xlabel('Time'); plt.ylabel('p')
     plt.show()
 
 if data == "pos3d":
@@ -19,7 +21,7 @@ if data == "pos3d":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbit around sun using Velocity Verlet",size=15)
+    plt.title("Earth orbiting the Sun using Velocity Verlet",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
     plt.legend(['Initial $v = 1.41 * 2\\pi$'],prop={'size': 15})
     plt.show()
@@ -30,7 +32,7 @@ if data == "pos3dr3":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbit around sun using Velocity Verlet",size=15)
+    plt.title("Earth orbiting the Sun using Velocity Verlet",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
     plt.legend(['Earth'],prop={'size': 15})
     plt.show()
@@ -45,7 +47,7 @@ if data == "pos3e":
     plt.figure()
     plt.plot(x,y)
     plt.plot(x2,y2)
-    plt.title("Earth and Jupiter orbit around sun using Velocity Verlet",size=15)
+    plt.title("Earth and Jupiter orbiting the Sun using Velocity Verlet",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
     plt.legend(['Earth','Jupiter'],prop={'size': 15})
     plt.show()
@@ -60,9 +62,10 @@ if data == "pos3e2":
     plt.figure()
     plt.plot(x,y)
     plt.plot(x2,y2)
-    plt.title("Earth and Jupiter orbit around sun using Velocity Verlet",size=15)
+    plt.plot(0,0,"ko")
+    plt.title("Earth and Jupiter orbiting the Sun",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
-    plt.legend(['Earth','Jupiter'],prop={'size': 15})
+    plt.legend(['Earth','Jupiter', 'Sun'],prop={'size': 15},loc='upper right')
     plt.show()
 
 
[…]

### 2018-10-22 Project3/main.cpp
Working on final model

@@ -13,8 +13,8 @@ int main(){
         //beta = atof(argv[1]);
         //vfac = atof(argv[2]);    // factor to change the velocity
 
-        Time = 7.5;    // time [years]
-        n = 100000;    // steps
+        Time = 24.0;    // time [years]
+        n = 10000;    // steps
         timestep = Time/n;
 
         // constants
@@ -24,7 +24,7 @@ int main(){
         M_sunval = 2e30;          // mass of sun
         M_sun = 1.0;              // relative mass of sun
         M_e = 6.0e24/M_sunval;    // relative mass of earth
-        M_j = 1000*1.9e27/M_sunval;    // relative mass of jupiter
+        M_j = 1.9e27/M_sunval;    // relative mass of jupiter
 
         planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
         planet Earth(M_e, 1, 0, 0, 0, 2.0*M_PI, 0);
@@ -45,18 +45,26 @@ int main(){
            }
            outfile.close();*/
         ofstream outfile;
-        outfile.open("data/3e2.txt");
+        outfile.open("data/3f.txt");
         for(int i=0; i <= n; i++) {
-                solver earth(G, timestep, Sun, Jupiter);
-                earth.VelocityVerletSystem(G, timestep, Earth, Sun, Jupiter);
+                solver earth(G, timestep, Jupiter, Sun);
+                earth.VelocityVerletSystem(G, timestep, Earth, Jupiter, Sun);
 
-                solver jupiter(G, timestep, Sun, Earth);
-                jupiter.VelocityVerletSystem(G, timestep, Jupiter, Sun, Earth);
+                solver jupiter(G, timestep, Earth, Sun);
+                jupiter.VelocityVerletSystem(G, timestep, Jupiter, Earth, Sun);
+
+                solver sun(G, timestep, Earth, Jupiter);
+                sun.VelocityVerletSystem(G, timestep, Sun, Earth, Jupiter);
 
                 outfile << Earth.position[0] << " ";
                 outfile << Earth.position[1] << " ";
                 outfile << Jupiter.position[0] << " ";
-                outfile << Jupiter.position[1] << endl;
+                outfile << Jupiter.position[1] << " ";
+                outfile << Sun.position[0] << " ";
+                outfile << Sun.position[1] << " ";
+                outfile << Sun.angularMomentum(Earth, Jupiter) << " ";
[…]

### 2018-10-22 Project3/planet.cpp

@@ -8,7 +8,7 @@
 
 planet::planet(){
         // sun
-        mass = 1.0;
+        mass = 0.0;
         position = vec({1.0, 0.0, 0.0});
         velocity = vec({0.0, 0.0, 0.0});
         potential = 0;
@@ -78,6 +78,13 @@ double planet::angularMomentum(planet otherPlanet){
         return this->mass*vel*this->distance(otherPlanet);
 };
 
+double planet::angularMomentum(planet otherPlanet1, planet otherPlanet2){
+        double vel = sqrt(pow(this->velocity[0],2) + pow(this->velocity[1],2) + pow(this->velocity[2],2));
+        double p1 = this->mass*vel*this->distance(otherPlanet1);
+        double p2 = this->mass*vel*this->distance(otherPlanet2);
+        return p1 + p2;
+};
+
 vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
         double r1 = this->distance(otherPlanet1);
         double r2 = this->distance(otherPlanet2);
@@ -102,7 +109,6 @@ vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
         else{
                 F2 = vec({0,0,0});
         }
-
         return (F1 + F2)/currentMass;;
 }
 

### 2018-10-22 Project3/planet.h

@@ -22,6 +22,7 @@ vec acceleration(planet otherPlanet, double G);
 double kineticEnergy();
 double potentialEnergy(double G, planet otherPlanet);
 double angularMomentum(planet otherPlanet);
+double angularMomentum(planet otherPlanet1, planet otherPlanet2);
 vec newton(double G, planet otherPlanet1, planet otherPlanet2);
 };
 

### 2018-10-22 Project3/pplot.py

@@ -68,6 +68,23 @@ if data == "pos3e2":
     plt.legend(['Earth','Jupiter', 'Sun'],prop={'size': 15},loc='upper right')
     plt.show()
 
+if data == "pos3f":
+    x = np.loadtxt("data/3f.txt", usecols=0)
+    y = np.loadtxt("data/3f.txt", usecols=1)
+    x2 = np.loadtxt("data/3f.txt", usecols=2)
+    y2 = np.loadtxt("data/3f.txt", usecols=3)
+    x3 = np.loadtxt("data/3f.txt", usecols=4)
+    y3 = np.loadtxt("data/3f.txt", usecols=5)
+
+
+    plt.figure()
+    plt.plot(x,y)
+    plt.plot(x2,y2)
+    plt.plot(x3,y3)
+    plt.title("Solar system of Earth, Jupiter and Sun",size=15)
+    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
+    plt.legend(['Earth','Jupiter', 'Sun'],prop={'size': 15},loc='upper right')
+    plt.show()
 
 if data == "energy":
     kinetic = np.loadtxt("data/energy.txt", usecols=0)

### 2018-10-22 Project3/solver.cpp

@@ -6,6 +6,10 @@ solver::solver(double G, double h, planet &currentPlanet, planet &otherPlanet) {
         // empty solver object for use with methods
 }
 
+solver::solver(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &otherPlanet2) {
+// empty solver object for use with methods
+}
+
 void solver::euler(double G, double h, planet &currentPlanet, planet &otherPlanet) {
         /* calculating over the adresse of vel og pos, so that the values in the class
            "planet" are changed. This is again used when calculating the acceleration.

### 2018-10-22 Project3/solver.h

@@ -17,7 +17,7 @@ vec *vel, *pos;
 
 //contructor
 solver(double G, double h, planet &currentPlanet, planet &otherPlanet);
-
+solver(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &otherPlanet2);
 // function
 void euler(double G, double h, planet &otherPlanet, planet &currentPlanet);
 void VelocityVerlet(double G, double h, planet &currentPlanet, planet &otherPlanet);

### 2018-10-22 Project3/main.cpp
Working on angular mom

@@ -26,9 +26,9 @@ int main(){
         M_e = 6.0e24/M_sunval;    // relative mass of earth
         M_j = 1.9e27/M_sunval;    // relative mass of jupiter
 
-        planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
-        planet Earth(M_e, 1, 0, 0, 0, 2.0*M_PI, 0);
-        planet Jupiter(M_j, 5.2, 0, 0, 0, 0.7*M_PI, 0);
+        planet Sun(M_sun, -0.0049, 0, 0, 0, -0.01, 0);
+        planet Earth(M_e, 1-0.0049, 0, 0, 0, 2.0*M_PI, 0);
+        planet Jupiter(M_j, 5.2-0.0049, 0, 0, 0, 0.7*M_PI, 0);
 
         /*ofstream outfile;
            outfile.open("data/3e.txt");

### 2018-10-22 Project3/planet.cpp

@@ -78,12 +78,11 @@ double planet::angularMomentum(planet otherPlanet){
         return this->mass*vel*this->distance(otherPlanet);
 };
 
-double planet::angularMomentum(planet otherPlanet1, planet otherPlanet2){
+double planet::angularMomentum(planet otherPlanet,planet otherPlanet2){
         double vel = sqrt(pow(this->velocity[0],2) + pow(this->velocity[1],2) + pow(this->velocity[2],2));
-        double p1 = this->mass*vel*this->distance(otherPlanet1);
-        double p2 = this->mass*vel*this->distance(otherPlanet2);
-        return p1 + p2;
+        return this->mass*vel*this->distance(otherPlanet);
 };
+
 
 vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
         double r1 = this->distance(otherPlanet1);

### 2018-10-22 Project3/Project3.tex
Updates from ShareLaTeX

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}

\begin{document}

\title{Project 3}

\author{Hans Erlend Bakken Glad, Torbjørn Lode Gjerberg, Christer Dreierstad, Stig-Nicolai Foyn}

\email{, ,stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}

\altaffiltext{1}{Institute of Theoretical Astrophysics, University of
  Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}

%\date{Received - / Accepted -}

\begin{abstract}
  State problem. Briefly describe method and data. Summarize main results.
\end{abstract}
\keywords{planetary movement --- cosmology: observations --- methods: velocity verlet, forward euler}

\section{Introduction}
\label{sec:introduction}

Discuss background, physical importance and possibly some history of
the problem that is being studied in this paper.

\section{Method}
\label{sec:method}

Describe method. Define data model and likelihood. Outline how the
likelihood was computed (grid or MCMC).

Define the power law model in terms of $Q$ and $n$.

\section{Data}
\label{sec:data}

Summarize properties of data. Which data are used (experiment,
frequencies etc.)? Pixel resolution ($N_{\textrm{side}}$),
$\ell_{\textrm{max}}$ -- everything necessary to repeat the analysis
for other researchers.

Show a sky map of the smoothed data. Use the Healpix routine
``smoothing'' to do this; it works just like anafast. Smooth with a
$7^{\circ}$ beam, and plot with ``map2gif''. Show the RMS pattern as
well.

\section{Results}
\label{sec:results}

Show the 2D likelihood contours. Summarize constraints on $Q$ and
$n$.

\section{Conclusions}
\label{sec:conclusions}

Summarize results. Discuss their importance, referring to the
discovery to the initial seeds for structure formation. Mention that
these results are in good agreement with expectations from
inflationary theory.

%\begin{figure}[t]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

\begin{deluxetable}{lccc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Summary of main results.}
\tablecolumns{4}
\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
\startdata
Item 1 & Item 2 & Item 3 & Item 4
\enddata
\end{deluxetable}

\begin{acknowledgements}
  Who do you want to thank for helping out with this project?
\end{acknowledgements}

\begin{thebibliography}{}

\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89

\end{thebibliography}

\end{document}

### 2018-10-22 LaTeX/template/ms.tex

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}

\begin{document}

\title{Project 3}

\author{Hans Erlend Bakken Glad, Torbjørn Lode Gjerberg, Christer Dreierstad, Stig-Nicolai Foyn}

\email{ola.d.nordmann@astro.uio.no}

\altaffiltext{1}{Institute of Physics, University of
  Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}

%\date{Received - / Accepted -}

\begin{abstract}

\end{abstract}
\keywords{planetary movement --- cosmology: observations --- methods: velocity verlet, forward euler}

\section{Introduction}
\label{sec:introduction}

Simulating the orbits of the planets in the Solar System is a suitable problem for making use of object-oriented code. While the two-body problem is straightforward to implement numerically, introducing multiple planets can quickly become tedious if the code is not written for this purpose. This is where object-orientation excels, as it is based on writing code once and then running it several times for different cases. In practice, this is done by writing classes in C++.

Our procedure for creating an object-oriented code starts with solving the motion of Earth around the Sun with no object-orientation in mind. This gives us an idea of which parts of our code that is suited for object-orientation. The calculation of the orbits is performed with Euler's method and then the velocity Verlet method, which allows us to compare their stability and properties such as energy conservation.

\section{Theory and method}
\label{sec:method}
a
\subsection{Integration methods}
The Euler method is a first-order method to approximate the solution to a differential equation. Given an analytical expression for the acceleration $a$ and an initial velocity $v(t)$ and position $x(t)$, the new values at a time $t + \Delta t$ can be found with

\begin{gather}
    v(t + \Delta t) = v(t) + a(t) \Delta t \\
    x(t + \Delta t) = x(t) + v(t) \Delta t \\
    a(t + \Delta t) = f( ... ).
\end{gather}
In our case, $f$ is a function of position only.

\subsection{Object-orientation}
a

\subsection{Scaling the equations governing the Solar System}
We will consider a model of our Solar System where gravity is the only force present. By Newton's law of gravitation, two celestial bodies with mass $M_1$ and $M_2$ separated by a distance $r$ will both experience a force with magnitude

\begin{equation}\label{eq:gforce}
    F_G = \frac{G M_1 M_2}{r^2},
\end{equation}
where $G = 6.67408 \times 10^{-11} \ \textrm{N} \ \textrm{kg}^{-2} \textrm{m}^2$. The acceleration experienced by body 1 is then
\begin{equation}
    a_1 = \frac{F_G}{M_1} = \frac{G M_2}{r^2},
\end{equation}
and similarly for body 2.

If we consider a 'flat' Solar System with only two dimensions (naming the axes $x$ and $y$), the acceleration can be decomposed into

\begin{equation}
    a_x = \frac{G M_2}{r^2}\cdot\frac{\Delta x}{r}, \quad a_y = \frac{G M_2}{r^2}\cdot\frac{\Delta y}{r},
\end{equation}
where $\Delta_x$ and $\Delta_y$ are the differences in position coordinates $x$ and $y$ between the two bodies. This can easily be extended to include the third dimension (the 'height' of the Solar System) as well.

In order to make our values easier to work with, we can scale the equations accordingly. We define the Sun to have a mass $M_\odot = 2\times 10^{30}$ kg and express the other masses (i. e. planets, moons) as a fraction of $M_\odot$. Distance is measured in terms of astronomical units AU, where $1 \ \textrm{AU} = 1.496 \times 10^{11} \ \textrm{meters}$ and time is measured in years. This is appropriate for the scales we are looking at, since the Earth takes one year (surprise) to orbit the Sun once at an average distance of 1 AU. The gravitational constant can now be expressed as

\begin{equation}
G = 4\pi^2 \ \frac{\textrm{AU}^3}{\textrm{yr}^{2}M} 
\end{equation}
where $M$ is the total mass in the system. From this we get the expression that will be used to calculate the acceleration $a$ induced by a body with mass $M$:

\begin{equation}
a = \frac{4\pi^2 M}{r^2}, \ \textrm{with} \ [a] = \frac{\textrm{AU}}{\textrm{yr}^2}.
\end{equation}

\subsection{Our simplest model: Earth orbiting a stationary Sun}

To get an idea on how to produce a realistic model of the Solar System, we will first look at a simple model with only the Earth and the Sun present. Earth's orbit is set to be perfectly circular, while in reality it is slightly elliptical. The velocity of a circular orbit expressed in terms of the radius (or orbital distance) $r$ and period $T$ is

\begin{equation}
    v_r = \frac{2\pi r}{T}.
\end{equation}
For Earth, we have $r = 1 \ \textrm{AU}$ and $T = 1 \ \textrm{yr}$. Earth's orbital velocity is thus always $2\pi$ AU/yr.

In our algorithm we position the Sun in the origin of our coordinate system. This further simplifies our calculations, as the distance between the Sun and Earth is simply defined by the coordinates $(x, y)$ of Earth. To test this simple case, we initialize the position of Earth at $x = 1 \ \textrm{AU}$ and $y = 0$. This means that the initial velocity in the $x$-direction has to be zero, otherwise we would not get a circular orbit. For a counter-clockwise orbit around the Sun, we then need to set the initial velocities to

\begin{equation}
    v_x = 0, \quad v_y = 2\pi.
\end{equation}

The stability of our Euler and velocity Verlet algorithms are then tested using different time steps $\Delta t$. By looking at the kinetic and potential energy of the system, we can also determine how well the two algorithms conserve energy. Total energy and angular momentum should be conserved in this system since gravity is the only force present, and because gravity is a conservative force we should not lose any energy. In theory, the Verlet algorithm should perform better (why?) than the Euler algorithm in this regard.

After this, we will not be using Euler's method due to it's lack of precision.

\subsection{Expanding our model: Adding Jupiter to the system}

After object-orienting our code, adding more planets should only be a matter of introducing new objects that represent the physical parameters of the planets. Adding the planet Jupiter will let us study a three-body problem, though we will still keep the Sun motionless. Due to the significant mass of Jupiter ($\approx 300$ times more massive than Earth) we should see some changes in Earth's orbit.

Introducing another planet means adding another term contributing to the total force on Earth. Jupiter will also experience this force, but the effect will be much less significant due the planet's high mass. The magnitude of this force is

\begin{equation}
    F_{\textrm{Earth-Jupiter}} = \frac{GM_{\textrm{Jupiter}} M_{\textrm{Earth}}}{r_{\textrm{Earth-Jupiter}}^2} = \frac{GM_E M_J}{r_{E-J}^2}
\end{equation}
where the masses of Earth and Jupiter are $M_{\textrm{E}} = 6 \times 10^{24}$ kg and $M_{\textrm{J}} = 1.9 \times 10^{27}$ kg. The distance between the two planets $r_{\textrm{E-J}}$ is calculated using the $x$ and $y$ coordinates of the planets,

\begin{equation}
    r_{\textrm{E-J}} = \sqrt{\left(x_E - x_j\right)^2 + \left(y_E - y_J\right)^2 } = \sqrt{(\Delta x_{E-J})^2 + (\Delta y_{E-J})^2}.
\end{equation}
Jupiter has an orbital radius of 5.2 AU with respect to the Sun, so that the distance between Earth and Jupiter can vary from 4.2 AU to 6.2 AU. Because $F \propto 1/r^2$, this means that the gravitational forces between the two planets vary by a factor

\begin{equation}
    \frac{1}{(4.2 \ \textrm{AU})^2} \bigg/  \frac{1}{(6.2 \ \textrm{AU})^2} \approx 2.2.
\end{equation}
At a distance of 4.2 AU away from Earth, the forces between the two are as high as possible. Comparing this to the force exerted by the Sun on Earth, we get

\begin{equation}
    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{G M_E M_{Sun}/r_{Earth-Sun}^2}{G M_E M_J/r_{Earth-Jupiter}^2} = \frac{M_{Sun} \ r_{Earth-Jupiter}^2}{M_J \ r_{Earth-Sun}^2}
\end{equation}
Using that $M_{J} = 9.5 \times 10^{-4} M_\odot$, we get

\begin{equation}
    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{M_\odot \cdot (4.2 \ \textrm{AU})^2}{9.5 \times 10^{-4} M_\odot \cdot (1 \ \textrm{AU})^2} \approx 18600.
\end{equation}
The gravitational force from Jupiter is around 18600 times weaker than the gravitational force from the Sun, even at the lowest possible Earth-Jupiter distance. This implies that the effects of Jupiter on Earth's orbit should be minimal. However, we will be scaling up the mass of Jupiter in order to examine the stability of our Verlet solver.

\subsection{Final model: Include all planets and Pluto}

Our final model of the Solar System will include all planets and the dwarf planet Pluto. Additionally, we will allow the gravitational attraction from the planets to also affect the Sun. We define the origin to be the center of mass of the system, then give the Sun an initial velocity that gives a total momentum of zero for the whole system. This ensures that the center of mass remains at the origin.

\section{Data}
\label{sec:data}

\section{Results}
\label{sec:results}

\section{Conclusions}
\label{sec:conclusions}

%\begin{figure}[t]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

\begin{deluxetable}{lccc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Summary of main results.}
\tablecolumns{4}
\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
\startdata
Item 1 & Item 2 & Item 3 & Item 4
\enddata
\end{deluxetable}

\begin{acknowledgements}
  Who do you want to thank for helping out with this project?
\end{acknowledgements}

\begin{thebibliography}{}

\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89

\end{thebibliography}

\end{document}

### 2018-10-22 Project3/main.cpp
clean up of files

@@ -10,9 +10,6 @@ int main(){
         double M_e, M_sun, M_j, M_sunval, G, Time, timestep, vfac, v0_e;
         int n;
 
-        //beta = atof(argv[1]);
-        //vfac = atof(argv[2]);    // factor to change the velocity
-
         Time = 24.0;    // time [years]
         n = 10000;    // steps
         timestep = Time/n;

### 2018-10-22 Project3/pplot_change.py

@@ -1,14 +1,0 @@
-import matplotlib.pyplot as plt
-import sys
-
-
-x = np.loadtxt("test.txt", usecols=0)
-y = np.loadtxt("test.txt", usecols=1)
-
-
-plt.figure()
-plt.plot(x,y)
-plt.title("Earth orbit around sun using Velocity Verlet",size=15)
-plt.xlabel("x",size=15); plt.ylabel("y",size=15)
-plt.legend(['Eearth'],prop={'size': 15})
-plt.show()

### 2018-10-22 Project3/pplot_energy.py

@@ -1,29 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-
-data = sys.argv[1]
-
-if data == "momentum":
-    angularMom = np.loadtxt("data/test.txt", usecols=0)
-    time = np.loadtxt("data/energy.txt", usecols=1)
-
-
-    plt.figure()
-    plt.plot(time,angularMom)
-    plt.show()
-
-
-if data == "energy":
-    kinetic = np.loadtxt("data/energy.txt", usecols=0)
-    potential = np.loadtxt("data/energy.txt", usecols=1)
-    time = np.loadtxt("data/energy.txt", usecols=2)
-
-    plt.figure()
-    plt.plot(time,kinetic)
-    plt.plot(time,potential)
-
-    plt.title("Energy of Earth in orbit around sun",size=15)
-    plt.xlabel("time [years]",size=15); plt.ylabel("Energy",size=15)
-    plt.legend(['Kinetic', 'Potential'],prop={'size': 15})
-    plt.show()

### 2018-10-22 Project3/pplot_to-merge.py

@@ -1,2 +1,0 @@
-import matplotlib.pyplot as plt
-import sys

### 2018-10-22 Project3/main.cpp
Implementing Mercury

@@ -7,61 +7,99 @@
 using namespace std;
 
 int main(){
-        double M_e, M_sun, M_j, M_sunval, G, Time, timestep, vfac, v0_e;
+        double M_e, M_sun, M_j, M_sunval, M_m, v_sun0, v_j0, v_e0, initPos_sun;
+        double totalMass, CenterOfMass, G, Time, timestep, vfac, v0_e;
         int n;
 
-        Time = 24.0;    // time [years]
-        n = 10000;    // steps
+        Time = 100;    // time [years]
+        n = 1e9;    // steps
         timestep = Time/n;
 
         // constants
         G = 4.0*pow(M_PI,2);    // gravitational constant [AU³/yr²]
 
         // masses
-        M_sunval = 2e30;          // mass of sun
-        M_sun = 1.0;              // relative mass of sun
-        M_e = 6.0e24/M_sunval;    // relative mass of earth
-        M_j = 1.9e27/M_sunval;    // relative mass of jupiter
+        M_sunval = 2.0e30;          // mass of sun
+        M_sun = 1.0;              // relative mass of Sun
+        M_e = 6.0e24/M_sunval;    // relative mass of Earth
+        M_j = 1.9e27/M_sunval;    // relative mass of Jupiter
+        M_m = 3.3e23/M_sunval;    // relative mass of Mercury
+        totalMass = M_sun + M_e + M_j;
 
-        planet Sun(M_sun, -0.0049, 0, 0, 0, -0.01, 0);
-        planet Earth(M_e, 1-0.0049, 0, 0, 0, 2.0*M_PI, 0);
-        planet Jupiter(M_j, 5.2-0.0049, 0, 0, 0, 0.7*M_PI, 0);
+        v_e0 = 2.0*M_PI;    // initial velocity of Earth
+        v_j0 = 0.7*M_PI;    // initial velocity of Jupiter
+        v_sun0 = -(v_e0*M_e + v_j0*M_j)/M_sun;  // initial velocuty of Sun
+
+        // CenterOfMass = (M_e*1.0 + M_j*5.2 + M_sun*0)/totalMass;
+        // initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
+
+        planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
[…]

### 2018-10-22 Project3/planet.cpp

@@ -78,9 +78,11 @@ double planet::angularMomentum(planet otherPlanet){
         return this->mass*vel*this->distance(otherPlanet);
 };
 
-double planet::angularMomentum(planet otherPlanet,planet otherPlanet2){
+double planet::angularMomentum(planet otherPlanet1,planet otherPlanet2){
         double vel = sqrt(pow(this->velocity[0],2) + pow(this->velocity[1],2) + pow(this->velocity[2],2));
-        return this->mass*vel*this->distance(otherPlanet);
+        double p1 = this->mass*vel*this->distance(otherPlanet1);
+        double p2 = this->mass*vel*this->distance(otherPlanet2);
+        return p1 + p2;
 };
 
 
@@ -103,7 +105,6 @@ vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
                 vec currentPos2 = this->position;
                 vec otherPos2 = otherPlanet2.position;
                 F2 = -G*otherMass*currentMass*(currentPos2 - otherPos2)/pow(r2,3);
-
         }
         else{
                 F2 = vec({0,0,0});
@@ -111,7 +112,16 @@ vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
         return (F1 + F2)/currentMass;;
 }
 
+vec planet::einstein(double G, planet otherPlanet) {
+        double r = this->distance(otherPlanet);
+        vec currentPos = this->position;
+        vec currentVel = this->velocity;
+        vec otherPos = otherPlanet.position;
 
+        double F = (-G*otherMass*(currentPos-otherPos)/pow(r,3))%
+                   (1+(3*abs((currentPos - otherPos)%(currentVel)))/(pow(r,2)*pow(c,2)));
+        return F;
+}
 
 
 //

### 2018-10-22 Project3/pplot.py

@@ -75,15 +75,36 @@ if data == "pos3f":
     y2 = np.loadtxt("data/3f.txt", usecols=3)
     x3 = np.loadtxt("data/3f.txt", usecols=4)
     y3 = np.loadtxt("data/3f.txt", usecols=5)
+    x4 = np.loadtxt("data/3f.txt", usecols=6)
+    y4 = np.loadtxt("data/3f.txt", usecols=7)
 
 
     plt.figure()
     plt.plot(x,y)
     plt.plot(x2,y2)
     plt.plot(x3,y3)
+    plt.plot(x4,y4)
     plt.title("Solar system of Earth, Jupiter and Sun",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
-    plt.legend(['Earth','Jupiter', 'Sun'],prop={'size': 15},loc='upper right')
+    plt.legend(['Earth','Jupiter', 'Sun', 'Center of Mass'],prop={'size': 15},loc='upper right')
+    plt.show()
+
+if data == "pos3g":
+    x = np.loadtxt("data/3g.txt", usecols=0)
+    y = np.loadtxt("data/3g.txt", usecols=1)
+    x2 = np.loadtxt("data/3g.txt", usecols=2)
+    y2 = np.loadtxt("data/3g.txt", usecols=3)
+    x3 = np.loadtxt("data/3g.txt", usecols=4)
+    y3 = np.loadtxt("data/3g.txt", usecols=5)
+
+
+    plt.figure()
+    plt.plot(x,y)
+    plt.plot(x2,y2)
+    plt.plot(x3,y3)
+    plt.title("Solar system of Earth, Mercury and Sun",size=15)
+    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
+    plt.legend(['Earth','Mercury', 'Sun'],prop={'size': 15},loc='upper right')
     plt.show()
 
 if data == "energy":

### 2018-10-22 Project3/solver.cpp

@@ -47,3 +47,13 @@ void solver::VelocityVerletSystem(double G, double h, planet &currentPlanet, pla
         accel_new = currentPlanet.newton(G, otherPlanet1, otherPlanet2);
         (*vel) += h/2.0*(accel_new + accel);
 };
+
+void solver::VelocityVerletEinstein(double G, double h, planet &currentPlanet, planet &otherPlanet) {
+        pos = &(currentPlanet.position);
+        vel = &(currentPlanet.velocity);
+        accel = currentPlanet.einstein(G, otherPlanet);
+
+        (*pos) += (*vel)*h + h*h/2.0*accel;
+        accel_new = currentPlanet.einstein(G, otherPlanet);
+        (*vel) += h/2.0*(accel_new + accel);
+}

### 2018-10-22 Project3/solver.h

@@ -23,6 +23,7 @@ void euler(double G, double h, planet &otherPlanet, planet &currentPlanet);
 void VelocityVerlet(double G, double h, planet &currentPlanet, planet &otherPlanet);
 void energy(double G, planet &currentPlanet, planet &otherPlanet);
 void VelocityVerletSystem(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &otherPlanet2);
+void VelocityVerletEinstein(double G, double h, planet &currentPlanet, planet &otherPlanet);
 };
 
 #endif /*solver_H*/

### 2018-10-22 Project3/main.cpp
Working on mercury

@@ -8,7 +8,7 @@ using namespace std;
 
 int main(){
         double M_e, M_sun, M_j, M_sunval, M_m, v_sun0, v_j0, v_e0, initPos_sun;
-        double totalMass, CenterOfMass, G, Time, timestep, vfac, v0_e;
+        double totalMass, CenterOfMass, G, Time, timestep, vfac, v0_e, c;
         int n;
 
         Time = 100;    // time [years]
@@ -29,6 +29,8 @@ int main(){
         v_e0 = 2.0*M_PI;    // initial velocity of Earth
         v_j0 = 0.7*M_PI;    // initial velocity of Jupiter
         v_sun0 = -(v_e0*M_e + v_j0*M_j)/M_sun;  // initial velocuty of Sun
+
+        c = 173*365.25;   // speed of light [AU/yr]
 
         // CenterOfMass = (M_e*1.0 + M_j*5.2 + M_sun*0)/totalMass;
         // initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
@@ -87,19 +89,18 @@ int main(){
 
            }
            outfile.close();*/
-
         ofstream outfile;
         outfile.open("data/3g.txt");
-        for(int i=0; i <= n; i++) {
+        for(int i=0; i <= n; i+=1000) {
                 solver mercury(G, timestep, Mercury, Sun);
-                mercury.VelocityVerletEinstein(G, timestep, Mercury, Sun);
+                mercury.VelocityVerletEinstein(G, c, timestep, Mercury, Sun);
                 solver sun(G, timestep, Sun, Mercury);
-                sun.VelocityVerletEinstein(G, timestep, Sun, Mercury);
+                sun.VelocityVerletEinstein(G, c, timestep, Sun, Mercury);
 
-                outfile << Earth.position[0] << " ";
-                outfile << Earth.position[1] << " ";
                 outfile << Mercury.position[0] << " ";
-                outfile << Mercury.position[1] << endl;
+                outfile << Mercury.position[1] << " ";
+                outfile << Sun.position[0] << " ";
+                outfile << Sun.position[1] << endl;
         }
         outfile.close();
 };

### 2018-10-22 Project3/planet.cpp

@@ -112,15 +112,17 @@ vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
         return (F1 + F2)/currentMass;;
 }
 
-vec planet::einstein(double G, planet otherPlanet) {
+vec planet::einstein(double G, double c, planet otherPlanet) {
         double r = this->distance(otherPlanet);
+        double currentMass = this->mass;
+        double otherMass = otherPlanet.mass;
         vec currentPos = this->position;
         vec currentVel = this->velocity;
         vec otherPos = otherPlanet.position;
+        vec l = cross((currentPos - otherPos),currentVel);
 
-        double F = (-G*otherMass*(currentPos-otherPos)/pow(r,3))%
-                   (1+(3*abs((currentPos - otherPos)%(currentVel)))/(pow(r,2)*pow(c,2)));
-        return F;
+        vec F = (-G*otherMass*(currentPos-otherPos)/pow(r,3))%(1 + (3*pow(l,2)/(pow(r,2)*pow(c,2))));
+        return F/currentMass;
 }
 
 

### 2018-10-22 Project3/planet.h

@@ -9,7 +9,7 @@ using namespace arma;
 class planet {
 public:
 
-double mass, potential, kinetic, angularMom;
+double mass, potential, kinetic, angularMom, c;
 vec position, velocity, F1, F2;
 
 // initializers
@@ -24,6 +24,7 @@ double potentialEnergy(double G, planet otherPlanet);
 double angularMomentum(planet otherPlanet);
 double angularMomentum(planet otherPlanet1, planet otherPlanet2);
 vec newton(double G, planet otherPlanet1, planet otherPlanet2);
+vec einstein(double G, double c, planet otherPlanet);
 };
 
 #endif /*planet_H*/

### 2018-10-22 Project3/pplot.py

@@ -94,17 +94,14 @@ if data == "pos3g":
     y = np.loadtxt("data/3g.txt", usecols=1)
     x2 = np.loadtxt("data/3g.txt", usecols=2)
     y2 = np.loadtxt("data/3g.txt", usecols=3)
-    x3 = np.loadtxt("data/3g.txt", usecols=4)
-    y3 = np.loadtxt("data/3g.txt", usecols=5)
 
 
     plt.figure()
     plt.plot(x,y)
     plt.plot(x2,y2)
-    plt.plot(x3,y3)
-    plt.title("Solar system of Earth, Mercury and Sun",size=15)
+    plt.title("Solar system of Mercury and Sun",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
-    plt.legend(['Earth','Mercury', 'Sun'],prop={'size': 15},loc='upper right')
+    plt.legend(['Mercury', 'Sun'],prop={'size': 15},loc='upper right')
     plt.show()
 
 if data == "energy":

### 2018-10-22 Project3/solver.cpp

@@ -48,12 +48,12 @@ void solver::VelocityVerletSystem(double G, double h, planet &currentPlanet, pla
         (*vel) += h/2.0*(accel_new + accel);
 };
 
-void solver::VelocityVerletEinstein(double G, double h, planet &currentPlanet, planet &otherPlanet) {
+void solver::VelocityVerletEinstein(double G, double c, double h, planet &currentPlanet, planet &otherPlanet) {
         pos = &(currentPlanet.position);
         vel = &(currentPlanet.velocity);
-        accel = currentPlanet.einstein(G, otherPlanet);
+        accel = currentPlanet.einstein(G, c, otherPlanet);
 
         (*pos) += (*vel)*h + h*h/2.0*accel;
-        accel_new = currentPlanet.einstein(G, otherPlanet);
+        accel_new = currentPlanet.einstein(G, c, otherPlanet);
         (*vel) += h/2.0*(accel_new + accel);
 }

### 2018-10-22 Project3/solver.h

@@ -11,7 +11,7 @@ class solver {
 
 public:
 
-double h, G;
+double h, G, c;
 vec accel, accel_new;
 vec *vel, *pos;
 
@@ -23,7 +23,7 @@ void euler(double G, double h, planet &otherPlanet, planet &currentPlanet);
 void VelocityVerlet(double G, double h, planet &currentPlanet, planet &otherPlanet);
 void energy(double G, planet &currentPlanet, planet &otherPlanet);
 void VelocityVerletSystem(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &otherPlanet2);
-void VelocityVerletEinstein(double G, double h, planet &currentPlanet, planet &otherPlanet);
+void VelocityVerletEinstein(double G, double c, double h, planet &currentPlanet, planet &otherPlanet);
 };
 
 #endif /*solver_H*/

### 2018-10-24 Project3/main.cpp
preload

@@ -40,7 +40,6 @@ int main(){
         //initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
 
         planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
-        //planet Mercury(M_m, -1.323E-01, -4.393E-01, -2.444E-02, 2.132E-02*365.25, 6.577E-03*365.25, -2.494E-03*365.25);
         planet Mercury(M_m, 0.3075, 0, 0, 0, 12.44, 0);
 
 

### 2018-10-24 Project3/pplot.py

@@ -4,6 +4,20 @@ import sys
 
 data = sys.argv[1]
 
+params = {'legend.fontsize':'x-large',
+        'axes.labelsize': 'x-large',
+        'axes.titlesize': 'x-large',
+        'xtick.labelsize': 'x-large',
+        'ytick.labelsize': 'x-large'}
+plt.rcParams.update(params)
+
+params = {'legend.fontsize': 'x-large',
+        'axes.labelsize': 'x-large',
+        'axes.titlesize': 'x-large',
+        'xtick.labelsize': 'x-large',
+        'ytick.labelsize': 'x-large'}
+plt.rcParams.update(params)
+
 if data == "momentum":
     angularMom = np.loadtxt("data/test.txt", usecols=0)
     time = np.loadtxt("data/test.txt", usecols=1)
@@ -21,7 +35,7 @@ if data == "pos3d":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbiting the Sun using Velocity Verlet",size=15)
+    plt.title("Earth orbiting the Sun",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
     plt.legend(['Initial $v = 1.41 * 2\\pi$'],prop={'size': 15})
     plt.show()

### 2018-10-24 Project3/main.cpp
For the stig

@@ -6,24 +6,28 @@
 
 using namespace std;
 
+void Write3cToFile(string method, double G, double h, int n, double M_e, double M_sun);
+void Write3eToFile(double G, double h, int n, double M_e, double M_j, double M_sun);
+void Write3dToFile(double G, double h, double beta, int n, double M_sun, double M_e);
+void Write3fToFile(double G, double h, int n, double M_e, double M_m, double M_sun);
+void Write3gToFile(double G, double h, int n, double c, double M_sun, double M_m);
+
 int main(){
         double M_e, M_sun, M_j, M_sunval, M_m, v_sun0, v_j0, v_e0, initPos_sun;
-        double totalMass, CenterOfMass, G, Time, timestep, vfac, v0_e, c;
-        double r_old2, r_old, r_new;
+        double totalMass, CenterOfMass, G, Time, timestep, vfac, v0_e, c, beta;
         int n;
-        vec r_pos;
-
-        vector<double> arcsec;
 
         Time = 100.0;    // time [years]
         n = 1e9;    // steps
         timestep = Time/n;
+        beta = 2.0;
 
         // constants
-        G = 4.0*pow(M_PI,2);    // gravitational constant [AU³/yr²]
+        G = 4.0*pow(M_PI,2);        // gravitational constant [AU³/yr²]
+        c = 63240.08034;      // speed of light [AU/yr]
 
         // masses
-        M_sunval = 2.0e30;          // mass of sun
+        M_sunval = 2.0e30;        // mass of sun
         M_sun = 1.0;              // relative mass of Sun
         M_e = 6.0e24/M_sunval;    // relative mass of Earth
         M_j = 1.9e27/M_sunval;    // relative mass of Jupiter
@@ -34,70 +38,152 @@ int main(){
         v_j0 = 0.7*M_PI;    // initial velocity of Jupiter
         v_sun0 = -(v_e0*M_e + v_j0*M_j)/M_sun;  // initial velocuty of Sun
 
-        c = 63240.08034;   // speed of light [AU/yr]
-
-
         //initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
 
+        Write3cToFile(string euler, G, timestep, n, M_e, M_sun);
+        //Write3eToFile(G, timestep, n, M_e, M_j, M_sun);
[…]

### 2018-10-24 Project3/planet.cpp

@@ -61,6 +61,20 @@ vec planet::acceleration(planet otherPlanet, double G) {
         }
 };
 
+vec planet::accelerationAlt(planet otherPlanet, double G, double beta) {
+        // calculating the acceleration a = -GM_sun/r²
+        double r = this->distance(otherPlanet);
+        if (r != 0) {
+                double otherMass = otherPlanet.mass;
+                vec currentPos = this->position;
+                vec otherPos = otherPlanet.position;
+                return -G*otherMass*(currentPos - otherPos)/pow(r,beta);
+        }
+        else {
+                return vec({0,0,0});
+        }
+};
+
 
 double planet::kineticEnergy(){
         double vel2 = pow(this->velocity[0],2) + pow(this->velocity[1],2) + pow(this->velocity[2],2);
@@ -112,15 +126,12 @@ vec planet::newton(double G, planet otherPlanet1, planet otherPlanet2) {
 
 vec planet::einstein(double G, double c, planet otherPlanet) {
         double r = this->distance(otherPlanet);
+        double m = this->mass;
+        vec F;
         vec v = this->velocity;
         vec rvec = this->position;
 
-
-        double m = this->mass;
-
         double l = abs(norm(cross(rvec,v)));
-
-        vec F;
 
         if (r != 0) {
                 F = rvec*(-G*m/pow(r,3)*(1 + 3.0*pow(l,2)/(pow(r,2)*pow(c,2))));

### 2018-10-24 Project3/planet.h

@@ -19,6 +19,7 @@ planet(double M, double x, double y, double z, double vx, double vy, double vz);
 // functions
 double distance(planet otherPlanet);
 vec acceleration(planet otherPlanet, double G);
+vec accelerationAlt(planet otherPlanet, double G, double beta);
 double kineticEnergy();
 double potentialEnergy(double G, planet otherPlanet);
 double angularMomentum(planet otherPlanet);

### 2018-10-24 Project3/pplot.py

@@ -4,7 +4,7 @@ import sys
 
 data = sys.argv[1]
 
-params = {'legend.fontsize':'x-large',
+"""params = {'legend.fontsize':'x-large',
         'axes.labelsize': 'x-large',
         'axes.titlesize': 'x-large',
         'xtick.labelsize': 'x-large',
@@ -16,7 +16,7 @@ params = {'legend.fontsize': 'x-large',
         'axes.titlesize': 'x-large',
         'xtick.labelsize': 'x-large',
         'ytick.labelsize': 'x-large'}
-plt.rcParams.update(params)
+plt.rcParams.update(params)"""
 
 if data == "momentum":
     angularMom = np.loadtxt("data/test.txt", usecols=0)
@@ -26,7 +26,7 @@ if data == "momentum":
     plt.figure()
     plt.title("Momentum of Earth-Sun system using Velocity Verlet")
     plt.plot(time,angularMom)
-    plt.xlabel('Time'); plt.ylabel('p')
+    plt.xlabel('Time [yr]'); plt.ylabel('p [solar mass AU$^2$/yr$^2$]')
     plt.show()
 
 if data == "pos3d":
@@ -36,7 +36,7 @@ if data == "pos3d":
     plt.figure()
     plt.plot(x,y)
     plt.title("Earth orbiting the Sun",size=15)
-    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
+    plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
     plt.legend(['Initial $v = 1.41 * 2\\pi$'],prop={'size': 15})
     plt.show()
 
@@ -62,7 +62,7 @@ if data == "pos3e":
     plt.plot(x,y)
     plt.plot(x2,y2)
     plt.title("Earth and Jupiter orbiting the Sun using Velocity Verlet",size=15)
-    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
+    plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
     plt.legend(['Earth','Jupiter'],prop={'size': 15})
     plt.show()
 
@@ -78,7 +78,7 @@ if data == "pos3e2":
     plt.plot(x2,y2)
     plt.plot(0,0,"ko")
     plt.title("Earth and Jupiter orbiting the Sun",size=15)
-    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
[…]

### 2018-10-24 Project3/solver.cpp

@@ -33,6 +33,16 @@ void solver::VelocityVerlet(double G, double h, planet &currentPlanet, planet &o
         (*vel) += h/2.0*(accel_new + accel);
 };
 
+void solver::VelocityVerletAlt(double G, double h, double beta, planet &currentPlanet, planet &otherPlanet) {
+        pos = &(currentPlanet.position);
+        vel = &(currentPlanet.velocity);
+        accel = currentPlanet.accelerationAlt(otherPlanet, G, beta);
+
+        (*pos) += (*vel)*h + h*h/2.0*accel;
+        accel_new = currentPlanet.accelerationAlt(otherPlanet, G, beta);
+        (*vel) += h/2.0*(accel_new + accel);
+};
+
 void solver::energy(double G, planet &currentPlanet, planet &otherPlanet) {
         currentPlanet.kinetic = currentPlanet.kineticEnergy();
         currentPlanet.potential = currentPlanet.potentialEnergy(G, otherPlanet);

### 2018-10-24 Project3/solver.h

@@ -21,6 +21,7 @@ solver(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &
 // function
 void euler(double G, double h, planet &otherPlanet, planet &currentPlanet);
 void VelocityVerlet(double G, double h, planet &currentPlanet, planet &otherPlanet);
+void VelocityVerletAlt(double G, double h, double beta, planet &currentPlanet, planet &otherPlanet);
 void energy(double G, planet &currentPlanet, planet &otherPlanet);
 void VelocityVerletSystem(double G, double h, planet &currentPlanet, planet &otherPlanet1, planet &otherPlanet2);
 void VelocityVerletEinstein(double G, double c, double h, planet &currentPlanet, planet &otherPlanet);

### 2018-10-24 Project3/report/Project3.tex

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}
\usepackage{url}
\setcitestyle{square}

\begin{document}

\title{Project 3}

\author{Hans Erlend Bakken Glad, Torbjørn Lode Gjerberg, Christer Dreierstad, Stig-Nicolai Foyn}

\email{heglad@student.matnat.uio.o,  SKRIV EPOSTEN DIN HER TORBJØRN,stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}

\altaffiltext{1}{Institute of Physics, University of
  Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}

%\date{Received - / Accepted -}

\begin{abstract}

\end{abstract}
\keywords{planetary movement --- cosmology: observations --- methods: velocity verlet, forward euler}

\section{Introduction}
\label{sec:introduction}

Simulating the orbits of the planets in the Solar System is a suitable problem for making use of object-oriented code. While the two-body problem is straightforward to implement numerically, introducing multiple planets can quickly become tedious if the code is not written for this purpose. This is where object-orientation excels, as it is built upon writing code once and then running it several times for different cases. In practice, this is done by writing classes in C++ and using them to solve various problems.

Our procedure for creating an object-oriented code starts with solving the motion of Earth around the Sun, but without implementing any classes first. This gives us an idea of which parts of our code is suited for object-orientation. The calculation of the orbits is performed with numerical integration, using the Euler method and the Velocity Verlet method. We will compare the stability of the two methods and how well they conserve physical quantities. After writing a proper object-oriented code, we will be able to extend our model to include all planets in the Solar System.
%(Demonstrate usefulness of object-oriented code)

\section{Theory and method}
\label{sec:method}

\subsection{Integration methods}
The Euler method (or Forward Euler) is a first-order approximation to the solution of a differential equation. We require an analytical expression $f$ for the acceleration $a$, an initial velocity $v$ and position $x$. In the one-dimensional case and with a step length of $h$, the values at the time $t + h$ are then
%
\begin{gather*}
    v_{t + h} = v_{t} + a_{t} h \\
    x_{t + h} = x_{t} + v_{t} h \\
    a_{t + h} = f(x_{t+1}) 
\end{gather*}
%
or on discretized form where $i+1$ represents the next point in time:
%
\begin{gather*}
    v(i + 1) = v(i) + a(i) h \\
    x(i + 1) = x(i) + v(i) h \\
    a(i + 1) = f(x(i+1))
\end{gather*}
%
In our program, this procedure is repeated until $t$ reaches the desired simulation time. This method can easily be extended to 2 or 3 spatial dimensions.

The problem with the Euler method is that it does not conserve the mechanical energy of the system, i. e. the sum of the kinetic and potential energy,
%
\begin{equation}
E_{tot} = E_k + U,
\end{equation}
%
is not constant over time. Because we are looking at a system with gravity as the only force, the total energy should be conserved because gravity is a conservative force. This issue may not be noticeable for the first few orbits, but this will eventually cause orbits to deviate significantly from their original shape.

A more suitable way of calculating planetary orbits is through using the Velocity Verlet method. This integration method is symmetric in time, which means that it should conserve the total energy of the system when $h \rightarrow 0$. We implement the Velocity Verlet method with
%
\begin{align*}
    x(i + 1) & = x(i) + v(i) h + a(i) h^2 \\
    a(i + 1) & = f(x(i+1)) \\
    v(i + 1) & = v(i) + \frac{h}{2}\left(a(i+1) + a(i)\right),
\end{align*}
%
and these values are calculated $n$ times with a given precision $h$.

\subsection{Object-orientation}

\subsection{Scaling the equations governing the Solar System}
We will consider a model of our Solar System where gravity is the only force present. By Newton's law of gravitation, two celestial bodies with mass $M_1$ and $M_2$ separated by a distance $r$ will both experience a force with magnitude
%
\begin{equation}\label{eq:gforce}
    F_G = \frac{G M_1 M_2}{r^2},
\end{equation}
where $G = 6.67408 \times 10^{-11} \ \textrm{N} \ \textrm{kg}^{-2} \textrm{m}^2$. The acceleration experienced by body 1 is then
%
\begin{equation*}
    a_1 = \frac{F_G}{M_1} = \frac{G M_2}{r^2},
\end{equation*}
%
and similarly for body 2.

If we consider a 'flat' Solar System with only two dimensions (naming the axes $x$ and $y$), the acceleration can be decomposed into
%
\begin{equation*}
    a_x = \frac{G M_2}{r^2}\cdot\frac{\Delta x}{r}, \quad a_y = \frac{G M_2}{r^2}\cdot\frac{\Delta y}{r},
\end{equation*}
%
where $\Delta_x$ and $\Delta_y$ are the differences in position coordinates $x$ and $y$ between the two bodies. This can easily be extended to include the third dimension (the 'height' of the Solar System) as well.

In order to make our values easier to work with, we will scale the equations accordingly. We define the Sun to have a mass $M_S = M_\odot = 2\times 10^{30}$ kg and express the other masses (i. e. planets, moons) as a fraction of $M_\odot$. Distance is measured in terms of astronomical units AU, where $1 \ \textrm{AU} = 1.496 \times 10^{11} \ \textrm{meters}$ and time is measured in years. This is appropriate for the scales we are looking at, since the Earth takes one year (surprise) to orbit the Sun once at an average distance of 1 AU. The gravitational constant can now be expressed as 
%
\begin{equation*}
G = 4\pi^2 \ \frac{\textrm{AU}^3}{\textrm{yr}^{2}M} 
\end{equation*}
%
where $M$ is the total mass in the system. From this we get the expression that will be used to calculate the acceleration $a$ induced by a body with mass $M$:
%
\begin{equation*}
a = \frac{4\pi^2 M}{r^2}, \ \textrm{with} \ [a] = \frac{\textrm{AU}}{\textrm{yr}^2}.
\end{equation*}

\subsection{Our simplest model: Earth orbiting a stationary Sun}

To get an idea on how to produce a realistic model of the Solar System, we will first look at a simple model with only the Earth and the Sun present. Earth's orbit is set to be perfectly circular, while in reality it is slightly elliptical. The velocity of a circular orbit expressed in terms of the radius (or orbital distance) $r$ and period $T$ is
%
\begin{equation*}
    v_r = \frac{2\pi r}{T}.
\end{equation*}
%
For Earth, we have $r = 1 \ \textrm{AU}$ and $T = 1 \ \textrm{yr}$. Earth's orbital velocity is thus always $2\pi$ AU/yr.

In our algorithm we position the Sun in the origin of our coordinate system. This further simplifies our calculations, as the distance between the Sun and Earth is simply defined by the coordinates $(x, y)$ of Earth. To test this simple case, we initialize the position of Earth at $x = 1 \ \textrm{AU}$ and $y = 0$. This means that the initial velocity in the $x$-direction has to be zero, otherwise we would not get a circular orbit. For a counter-clockwise orbit around the Sun, we then need to set the initial velocities to
%
\begin{equation*}
    v_x = 0, \quad v_y = 2\pi.
\end{equation*}
%
The stability of our Euler and velocity Verlet algorithms are then tested using different time steps $\Delta t$. By looking at the kinetic and potential energy of the system, we can also determine how well the two algorithms conserve energy. Total energy and angular momentum should be conserved in this system since gravity is the only force present, and because gravity is a conservative force we should not lose any energy. In theory, the Verlet algorithm should perform better (why?) than the Euler algorithm in this regard.

After this, we will not be using Euler's method due to its lack of precision.

\subsection{Expanding our model: Adding Jupiter to the system}

After object-orienting our code, adding more planets should only be a matter of introducing new objects that represent the physical parameters of the planets. Adding the planet Jupiter will let us study a three-body problem, though we will still keep the Sun motionless. Due to the significant mass of Jupiter ($\approx 300$ times more massive than Earth) we should see some changes in Earth's orbit.

Introducing another planet means adding another term contributing to the total force on Earth. Jupiter will also experience this force, but the effect will be much less significant due the planet's high mass. The magnitude of this force is
%
\begin{equation*}
    F_{\textrm{Earth-Jupiter}} = \frac{GM_{\textrm{Jupiter}} M_{\textrm{Earth}}}{r_{\textrm{Earth-Jupiter}}^2} = \frac{GM_E M_J}{r_{E-J}^2}
\end{equation*}
%
where the masses of Earth and Jupiter are $M_{\textrm{E}} = 6 \times 10^{24}$ kg and $M_{\textrm{J}} = 1.9 \times 10^{27}$ kg. The distance between the two planets $r_{\textrm{E-J}}$ is calculated using the $x$ and $y$ coordinates of the planets,
%
\begin{align*}
    r_{\textrm{E-J}} &= \sqrt{\left(x_E - x_j\right)^2 + \left(y_E - y_J\right)^2 } \\
    &= \sqrt{(\Delta x_{E-J})^2 + (\Delta y_{E-J})^2}.
\end{align*}
%
Jupiter has an orbital radius of 5.2 AU with respect to the Sun, so that the distance between Earth and Jupiter can vary from 4.2 AU to 6.2 AU. Because $F \propto 1/r^2$, this means that the gravitational forces between the two planets vary by a factor
%
\begin{equation*}
    \frac{1}{(4.2 \ \textrm{AU})^2} \bigg/  \frac{1}{(6.2 \ \textrm{AU})^2} \approx 2.2.
\end{equation*}
%
At a distance of 4.2 AU away from Earth, the forces between the two are as high as possible. Comparing this to the force exerted by the Sun on Earth, we get
%
\begin{equation*}
    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{G M_E M_{\odot}/r_{E-S}^2}{G M_E M_J/r_{E-J}^2} = \frac{M_{\odot} \ r_{E-J}^2}{M_J \ r_{E-S}^2},
\end{equation*}
%
where the indices representing Earth, Sun and Jupiter are shortened to E, S and J respectively.
Using that $M_{J} = 9.5 \times 10^{-4} M_\odot$, we get
%
\begin{equation*}
    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{M_\odot \cdot (4.2 \ \textrm{AU})^2}{9.5 \times 10^{-4} M_\odot \cdot (1 \ \textrm{AU})^2} \approx 18600.
\end{equation*}
%
The gravitational force from Jupiter is around 18600 times weaker than the gravitational force from the Sun, even at the lowest possible Earth-Jupiter distance. This implies that the effects of Jupiter on Earth's orbit should be minimal. However, we will be scaling up the mass of Jupiter in order to examine the stability of our Verlet solver, as a higher Jupiter mass introduces stronger forces. This in turn causes less stability in the system.

\subsection{Final model: Include all planets and Pluto}

Until now we have not considered the gravitational effects from the planets on the Sun. In reality, the Sun will also gain a velocity of its own, though this velocity will be low compared to the planet velocities. For this reason it is best to center our reference system at the center of mass of the Solar System. We define the center of mass to be at the origin, then give the Sun an initial velocity that gives a total momentum of zero for the whole system. This ensures that the center of mass remains at the origin.

By using these conditions, we find the initial position of the Sun by defining the center of mass $R$ as:
%
\begin{equation*}
    R = \frac{1}{M}(M_S r_S + M_J r_J + M_E r_E),
\end{equation*}
%
where $M$ is the sum of all masses present, $M = M_S + M_J + M_E$.
We find the initial position of the Sun by setting $R=0$ and solving for $r_S$:
%
\begin{equation*}
    r_S = - \frac{1}{M_S M}\left( M_J r_J + M_E r_E \right) = -0.0049 \ \textrm{AU}.
\end{equation*}
%
This calculation assumes that the Sun, Earth and Jupiter all lie along the x-axis in that order. The initial position of the Sun is thus
%
\begin{equation}
    x_0 = -0.0049 \ \textrm{AU}, \quad y_0 = 0.
\end{equation}
%
The Sun's initial velocity is determined from the condition that the center of mass remains stationary, i. e. the momentum $P$ of the center of mass is zero:
%
\begin{equation*}
    P = M_S v_S + M_J v_J + M_E v_J = 0.
\end{equation*}
%
Solving for $v_S$ this gives (again assuming that the Solar System lies along the x-axis):
%
\begin{equation*}
    v_S = -\frac{1}{M_S}\left( M_E v_E + M_J v_J \right) = -0.0021 \ \textrm{AU/yr},
\end{equation*}
%
so that the initial velocity components are 
%
\begin{equation*}
    v_{x} = 0, \quad v_{y} =  -0.0021 \ \textrm{AU/yr}.
\end{equation*}
%
Our final model of the Solar System will include all planets and the dwarf planet Pluto. By using data from NASA \cite{bib:nasa}, we are able to set the initial conditions of the bodies in the Solar System to their values at the current time.

\subsection{Escape velocity of the Earth}

The escape velocity of an object in a gravitational field is found by setting up the total energy before and after escaping. If the Earth were to have a high enough velocity, the planet would escape the gravitational influence of the Sun. We consider a hypothetical scenario where the gravitational force between the Sun and Earth is defined by
%
\begin{equation*}
    F_G = \frac{G M_\odot M_E}{r^\beta}, \quad \beta \in [2,3].
\end{equation*}
%
For conservative forces, the potential is defined as the integral of the force along a path. This gives us 
%
\begin{align*}
    U & = \int F_G \ dr \\
    & = \int \frac{G M_\odot M_E}{r^\beta} dr \\
    & = \frac{1}{1-\beta} \frac{G M_\odot M_E}{r^{\beta-1}}.
\end{align*}
%
The minimum velocity that the Earth needs to escape is found by setting the final velocity (after escaping) to zero. We can also set the potential energy to zero after the planet has escaped, since $r \rightarrow \infty$. The total energy before and after escaping is then 
%
\begin{equation*}
    \frac{1}{2}M_E v^2 + \frac{1}{1-\beta} \frac{G M_\odot M_E}{r^{\beta-1}} = \frac{1}{2}M_E \cdot 0 + 0
\end{equation*}
%
so that 
%
\begin{gather*}
    v^2 = -\frac{2}{1-\beta} \frac{G M_\odot}{r^{\beta-1}} \\
  \Rightarrow \ v_{escape} = \sqrt{ -\frac{1}{1 - \beta} \frac{8\pi^2}{r^{\beta - 1}} }.
\end{gather*}
%
Since $\beta$ is always greater than 1 we avoid getting complex values. This lets us study the effects on the escape velocity by changing the strength of the gravitational force, as a higher value of $\beta$ corresponds to a weaker force.

\subsection{Perihelion precession of Mercury}

The observed perihelion precession of Mercury is

\section{Results}
\label{sec:results}

\section{Discussion}
\label{sec:discussion}

\section{Conclusions}
\label{sec:conclusions}

%\begin{figure}[t]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

\begin{deluxetable}{lccc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Summary of main results.}
\tablecolumns{4}
\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
\startdata
Item 1 & Item 2 & Item 3 & Item 4
\enddata
\end{deluxetable}

\begin{acknowledgements}

\end{acknowledgements}

\begin{thebibliography}{}
\bibitem{bib:lectures} Hjorth-Jensen, Morten, 2015, 
\textit{Computational Physics - Lecture Notes Fall 2015}

\bibitem{bib:nasa} \textit{HORIZONS Web Interface}, retrieved from \url{https://ssd.jpl.nasa.gov/horizons.cgi#top} (22.10.18)

\bibitem{1}
 Stig-Nicolai Foyn, Christer Dreierstad
\textit{Project 1} \url{https://github.com/chdre/FYS3150/blob/master/Project1/report/FYS3150-Project1-stignicf-chrisdre.pdf} (01.10.18)
\end{thebibliography}

%\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
  %Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
  %A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89

\end{document}

### 2018-10-24 Project3/main.cpp
For the stig

@@ -39,17 +39,15 @@ int main(){
         v_sun0 = -(v_e0*M_e + v_j0*M_j)/M_sun;  // initial velocuty of Sun
 
         //initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
-
-        Write3cToFile(string euler, G, timestep, n, M_e, M_sun);
+        string euler;
+        Write3cToFile(euler, G, timestep, n, M_e, M_sun);
         //Write3eToFile(G, timestep, n, M_e, M_j, M_sun);
         //Write3dToFile(G, timestep, beta, n, M_sun, M_e);
         //Write3fToFile(G, timestep, n, M_e, M_m, M_sun);
         //Write3gToFile(G, timestep, n, c, M_sun, M_m)M
-
-
-
-
 }
+
+
 void Write3cToFile(string method, double G, double h, int n, double M_e, double M_sun){
         planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
         planet Earth(M_e, 1, 0, 0, 0, 2.0*M_PI, 0);
@@ -133,7 +131,6 @@ void Write3eToFile(double G, double h, int n, double M_e, double M_j, double M_s
         }
         outfile.close();
 }
-
 
 void Write3fToFile(double G, double h, int n, double M_e, double M_m, double M_sun){
         double v0_j = sqrt(4.0*pow(M_PI,2)/5.2);

### 2018-10-24 Project3/pplot.py

@@ -24,7 +24,7 @@ if data == "momentum":
 
 
     plt.figure()
-    plt.title("Momentum of Earth-Sun system using Velocity Verlet")
+    plt.title("Momentum of Earth-Sun system")
     plt.plot(time,angularMom)
     plt.xlabel('Time [yr]'); plt.ylabel('p [solar mass AU$^2$/yr$^2$]')
     plt.show()
@@ -46,7 +46,7 @@ if data == "pos3dr3":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbiting the Sun using Velocity Verlet",size=15)
+    plt.title("Earth orbiting the Sun",size=15)
     plt.xlabel("x",size=15); plt.ylabel("y",size=15)
     plt.legend(['Earth'],prop={'size': 15})
     plt.show()
@@ -61,7 +61,7 @@ if data == "pos3e":
     plt.figure()
     plt.plot(x,y)
     plt.plot(x2,y2)
-    plt.title("Earth and Jupiter orbiting the Sun using Velocity Verlet",size=15)
+    plt.title("Earth and Jupiter orbiting the Sun",size=15)
     plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
     plt.legend(['Earth','Jupiter'],prop={'size': 15})
     plt.show()
@@ -140,7 +140,7 @@ if data == "pos":
 
         plt.figure()
         plt.plot(x,y)
-        plt.title("Earth orbiting the Sun using Velocity Verlet",size=15)
+        plt.title("Earth orbiting the Sun (Velocity Verlet)",size=15)
         plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
         plt.legend(['Earth'],prop={'size': 15})
         plt.show()
@@ -150,7 +150,7 @@ if data == "pos":
 
         plt.figure()
         plt.plot(x,y)
-        plt.title("Earth orbit around sun using Euler-Cromer",size=15)
+        plt.title("Earth orbit around Sun (Euler-Cromer)",size=15)
         plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
         plt.legend(['Earth'],prop={'size': 15})
         plt.show()
@@ -161,7 +161,7 @@ if data == "pos":
 
         plt.figure()
         plt.plot(x,y)
-        plt.title("Earth orbiting the Sun using Euler",size=15)
[…]

### 2018-10-24 Project3/3fplot.py
Perhelion not solved. All else finished

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 x = np.loadtxt("data/testdata3f.txt")
-t = np.linspace(0,100,415)
+t = np.linspace(0,100,432)
 
 plt.plot(t,x)
 plt.show()

### 2018-10-24 Project3/main.cpp

@@ -6,7 +6,8 @@
 
 using namespace std;
 
-void Write3cToFile(string method, double G, double h, int n, double M_e, double M_sun);
+void Write3cToFile(int method, double G, double h, int n, double M_e, double M_sun);
+void WriteEnergynMomentumToFile(double G, double h, int n, double M_sun, double M_e);
 void Write3eToFile(double G, double h, int n, double M_e, double M_j, double M_sun);
 void Write3dToFile(double G, double h, double beta, int n, double M_sun, double M_e);
 void Write3fToFile(double G, double h, int n, double M_e, double M_m, double M_sun);
@@ -18,9 +19,9 @@ int main(){
         int n;
 
         Time = 100.0;    // time [years]
-        n = 1e9;    // steps
+        n = 1e8;    // steps
         timestep = Time/n;
-        beta = 2.0;
+        beta = 3;
 
         // constants
         G = 4.0*pow(M_PI,2);        // gravitational constant [AU³/yr²]
@@ -38,30 +39,51 @@ int main(){
         v_j0 = 0.7*M_PI;    // initial velocity of Jupiter
         v_sun0 = -(v_e0*M_e + v_j0*M_j)/M_sun;  // initial velocuty of Sun
 
-        //initPos_sun = -(5.2*M_j + M_e)/(totalMass*M_sun);   // initial position of Sun found by requiring center of mass = 0
-        string euler;
-        Write3cToFile(euler, G, timestep, n, M_e, M_sun);
-        //Write3eToFile(G, timestep, n, M_e, M_j, M_sun);
+        //Write3cToFile(1, G, timestep, n, M_e, M_sun);
+        //WriteEnergynMomentumToFile(G, timestep, n, M_sun, M_e);
         //Write3dToFile(G, timestep, beta, n, M_sun, M_e);
+        //Write3eToFile(G, timestep, n, M_e, M_j, M_sun);
         //Write3fToFile(G, timestep, n, M_e, M_m, M_sun);
-        //Write3gToFile(G, timestep, n, c, M_sun, M_m)M
+        Write3gToFile(G, timestep, n, c, M_sun, M_m);
 }
 
-
-void Write3cToFile(string method, double G, double h, int n, double M_e, double M_sun){
+void WriteEnergynMomentumToFile(double G, double h, int n, double M_sun, double M_e){
         planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
[…]

### 2018-10-24 Project3/planet.cpp

@@ -62,13 +62,12 @@ vec planet::acceleration(planet otherPlanet, double G) {
 };
 
 vec planet::accelerationAlt(planet otherPlanet, double G, double beta) {
-        // calculating the acceleration a = -GM_sun/r²
         double r = this->distance(otherPlanet);
         if (r != 0) {
                 double otherMass = otherPlanet.mass;
                 vec currentPos = this->position;
                 vec otherPos = otherPlanet.position;
-                return -G*otherMass*(currentPos - otherPos)/pow(r,beta);
+                return -G*otherMass*(currentPos - otherPos)/pow(r,beta+1);
         }
         else {
                 return vec({0,0,0});

### 2018-10-24 Project3/pplot.py

@@ -19,14 +19,14 @@ params = {'legend.fontsize': 'x-large',
 plt.rcParams.update(params)"""
 
 if data == "momentum":
-    angularMom = np.loadtxt("data/test.txt", usecols=0)
-    time = np.loadtxt("data/test.txt", usecols=1)
+    angularMom = np.loadtxt("data/3c-energy.txt", usecols=2)
+    time = np.loadtxt("data/3c-energy.txt", usecols=3)
 
 
     plt.figure()
-    plt.title("Momentum of Earth-Sun system")
+    plt.title("Angular momentum of Earth-Sun system")
     plt.plot(time,angularMom)
-    plt.xlabel('Time [yr]'); plt.ylabel('p [solar mass AU$^2$/yr$^2$]')
+    plt.xlabel('Time [yr]'); plt.ylabel('L $M_{\\odot} AU$^2$/yr]')
     plt.show()
 
 if data == "pos3d":
@@ -35,9 +35,10 @@ if data == "pos3d":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbiting the Sun",size=15)
+    plt.plot(0,0, "C1o")
+    plt.title("Earth escaping orbit of Sun",size=15)
     plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
-    plt.legend(['Initial $v = 1.41 * 2\\pi$'],prop={'size': 15})
+    plt.legend(['Earth', 'Sun'],prop={'size': 15})
     plt.show()
 
 if data == "pos3dr3":
@@ -46,9 +47,10 @@ if data == "pos3dr3":
 
     plt.figure()
     plt.plot(x,y)
-    plt.title("Earth orbiting the Sun",size=15)
-    plt.xlabel("x",size=15); plt.ylabel("y",size=15)
-    plt.legend(['Earth'],prop={'size': 15})
+    plt.plot(0,0, "C1o")
+    plt.title("Earth-Sun system with modified gravity",size=15)
+    plt.xlabel("x [AU]",size=15); plt.ylabel("y [AU]",size=15)
+    plt.legend(['Earth', 'Sun'],prop={'size': 15})
     plt.show()
 
 if data == "pos3e":
@@ -59,27 +61,12 @@ if data == "pos3e":
 
 
     plt.figure()
-    plt.plot(x,y)
-    plt.plot(x2,y2)
+    plt.plot(x,y,)
+    plt.plot(x2,y2,'C5')
[…]

### 2018-10-24 Project3/solver.cpp

@@ -18,9 +18,10 @@ void solver::euler(double G, double h, planet &currentPlanet, planet &otherPlane
         pos = &(currentPlanet.position);
         accel = currentPlanet.acceleration(otherPlanet, G);
 
+        (*pos) += (*vel)*h;
         // euler-cromer. Change position of vel and pos below to use euler
         (*vel) += accel*h;
-        (*pos) += (*vel)*h;
+
 };
 
 void solver::VelocityVerlet(double G, double h, planet &currentPlanet, planet &otherPlanet) {

### 2018-10-24 Project3/testcase.cpp
Finished

@@ -1,0 +1,46 @@
+#define CATCH_CONFIG_MAIN
+#include "catch.hpp"
+#include "planet.h"
+#include "solver.h"
+
+
+TEST_CASE("Total energy of system"){
+
+        double M_e, M_sun, M_sunval, G, Time, timestep;
+        double vfac, v0_e, t_0, k1, p1, k2, p2, energy1, energy2;
+        int n;
+
+        Time = 5.0; // time [years]
+        n = 1e2; // steps
+        timestep = Time/n;
+
+        // constants
+        G = 4.0*pow(M_PI,2); // gravitational constant [AU³/yr²]
+
+        // masses
+        M_sunval = 2e30;    // mass of sun
+        M_sun = 1.0;        // relative mass of sun
+        M_e = 6.0e24/M_sunval; // relative mass of earth
+
+        planet Sun(M_sun, 0, 0, 0, 0, 0, 0);
+        planet Earth(M_e, 1, 0, 0, 0, 2.0*M_PI, 0);
+
+
+        k1 = Earth.kineticEnergy();
+        p1 = Earth.potentialEnergy(G, Sun);
+
+
+        energy1 = k1 + p1;
+        for(int i=0; i <= n; i++) {
+                solver earth(G, timestep, Earth, Sun);
+                earth.VelocityVerlet(G, timestep, Earth, Sun);
+                k2 = Earth.kineticEnergy();
+                p2 = Earth.potentialEnergy(G, Sun);
+        }
+
+
+        energy2 = k2 + p2;
+        cout << energy1 << " " << energy2 << endl;
+
+        REQUIRE(energy1 == Approx(energy2));
+}

### 2018-11-05 Project4/p4.cpp
Done with metro

@@ -1,0 +1,115 @@
+#include <cmath>
+#include <iostream>
+#include <fstream>
+#include <random>
+#include <armadillo>
+#include <random>
+#include <map>
+
+using namespace std;
+using namespace arma;
+
+inline int PB(int index, int spins, int correction) {
+        // returns the index of the element
+        return (index + spins + correction)%(spins);
+}
+
+void Energy(int n, mat &SMat, double &energy);
+void Metropolis(int n, int mcs, double T);
+map<double, double> energies(double T);
+
+
+int main(int argc, char *argv[]){
+        int n, mcc;
+        double T;
+
+        n = atoi(argv[1]);    // number of spins
+        mcc = atoi(argv[2]);  // number of MC cycles
+        T = atof(argv[3]);    // temperature [kT/J]
+
+}
+
+void Metropolis(int n, int mcc, double T, map<double, double> energyDiff){
+        random_device rd;
+        mt19937_64 generator(rd());
+        uniform_int_distribution<> random_spin(0, 1);
+        uniform_real_distribution<> RNG(0.0,1.0);
+
+        mat SMat = ones<mat>(n,n);    // ground state
+        mat ExpectVals = zeros(n,5);  // expectation values
+
+        double energy = 0;  // energy
+        double magMoment = pow(double(n),2); // magnetic moment (initial value, of ground state = n²)
+
+        // calculating energy of lattice
+        Energy(n, SMat, energy);
+
+        for(int m = 1; m < mcc; m++) {
+                for (int x = 0; x < n; x++) {
+                        for (int y = 0; y < n; y++) {
+                                SMat(x,y) = 2*random_spin(generator) - 1;   // randomizing spin at x,y
+                                int dE = 2*SMat(x,y)*(SMat(x, PB(x, n, 1)) + SMat(x, PB(x, n, -1))
+                                                      + SMat(PB(y, n, 1),y) + SMat(PB(y, n, -1),y));
+
+                                if (RNG(generator) <= energyDiff.find(dE)->second) {
+                                        SMat(x,y) *= -1;   // flipping spin
+                                        magMoment += (double) 2.0*SMat(x,y);
+                                        energy += (double) dE;
+                                }
+                                else{
+                                        if (RNG(generator) <= energyDiff.find(dE)->second) {
+                                                SMat(x,y) *= -1; // flipping spin
+                                                magMoment += (double) 2.0*SMat(x,y);
[…]

### 2018-11-05 Project3/report/Project3.tex

@@
 
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
@@
 
-\title{Project 3}
+\title{Modeling the Solar System using object-oriented code to solve differential equations}
 
@@
 
-\email{heglad@student.matnat.uio.o,  SKRIV EPOSTEN DIN HER TORBJØRN,stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}
+\email{heglad@student.matnat.uio.o,  torbjolg@student.matnat.uio.no ,stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}
 
@@
 \begin{abstract}
+Solving differential equations describing the motion of a planet orbiting a star as a two-body problem is easily solved using simple numerical methods, such as the Euler method.
+When considering the energy of the system we observe that the Euler method does not conserve energy, as time goes the system gains energy.
+Further we observe that the angular momentum increases with time using the Euler method.
+Implementing the Velocity Verlet method we get an energy conserving method that produces better results on fewer steps, at the cost of FLOPS.
+We observe that Velocity Verlet is slower than the Euler method due to the amount of FLOPS needed by Velocity Verlet, but only by a factor close to two.
+Expanding on the model adding more planets in orbit is most easily done by object-orientation, making the program do what would have had to be hard coded otherwise.
+
+When changing the $1/r^2$ dependency of the force in the two-body Earth-Sun system to $1/r^{\beta}$ where $\beta \in [2,3]$ we observe that the Earth is in a bound state for $\beta < 3$.
+When $\beta = 3$ the Earth leaves the orbit around the Sun.
+Introducing the three-body problem as a system of the Earth, Jupiter and the Sun, where the latter's position is fixed at the origin we study the stability of the Velocity Verlet method.
+The stability is studied by changing the mass of Jupiter such that we dramatically change the effect it has on the Earth.
+With the mass 10 times larger the method is still stable, increasing the mass by a factor of 1000 we observe that the Earth is flung out of orbit around the Sun, with a dependency upon the resolution of the calculations.
+By increasing the amount of steps Earth stays in orbit.
+For a solar system consisting of Mercury and the Sun we observe that the perihelion precession is 26'' over a century (100 years).
 
 \end{abstract}
-\keywords{planetary movement --- cosmology: observations --- methods: velocity verlet, forward euler}
+\keywords{planetary movement --- computational science --- methods: Velocity Verlet, Forward Euler}
 
@@
 \label{sec:introduction}
-
 Simulating the orbits of the planets in the Solar System is a suitable problem for making use of object-oriented code.
-While the two-body problem is straightforward to implement numerically, introducing multiple planets can quickly become tedious if the code is not written for this purpose.
+While the two-body problem is straightforward to solve numerically, introducing multiple planets can quickly become tedious if the program is not written for this purpose.
 This is where object-orientation excels, as it is built upon writing code once and then running it several times for different cases.
@@
 
-Our procedure for creating an object-oriented code starts with solving the motion of Earth around the Sun, but without implementing any classes first.
-This gives us an idea of which parts of our code is suited for object-orientation.
+By solving differential equations for the motion of the Earth around a static Sun, we conclude that implementing more bodies will require the same operation for each body.
+Setting up an object-oriented code and treating the bodies of the Solar System as objects will allow the the program to handle the repetitive calculations using a single constructor or function.
 The calculation of the orbits is performed with numerical integration, using the Euler method and the Velocity Verlet method.
 We will compare the stability of the two methods and how well they conserve physical quantities.
-After writing a proper object-oriented code, we will be able to extend our model to include all planets in the Solar System.
+
+Having a fully object-oriented code, we will be able to extend our model to include all planets in the Solar System.
 %(Demonstrate usefulness of object-oriented code)
@@
 The Euler method (or Forward Euler) is a first-order approximation to the solution of a differential equation.
-We require an analytical expression $f$ for the acceleration $a$, an initial velocity $v$ and position $x$.
+In our case, we want to solve for the position of a planet given its acceleration.
+This means that we require an analytical expression $f$ for the acceleration $a$, an initial velocity $v$ and position $x$.
 In the one-dimensional case and with a step length of $h$, the values at the time $t + h$ are then
@@
 \begin{gather*}
-    v_{t + h} = v_{t} + a_{t} h \\
-    x_{t + h} = x_{t} + v_{t} h \\
+    v_{t + h} = v_{t} + a_{t} \cdot h \\
+    x_{t + h} = x_{t} + v_{t} \cdot h \\
     a_{t + h} = f(x_{t+1}) 
@@
 %
-or on discretized form where $i+1$ represents the next point in time:
+or on discretized form with $n$ points, so that a point in time is represented by $ i = \left[0, 1, \dots n-1 \right]$:
 %
 \begin{gather*}
-    v(i + 1) = v(i) + a(i) h \\
-    x(i + 1) = x(i) + v(i) h \\
-    a(i + 1) = f(x(i+1))
+    v_{i + 1} = v_i + a_i h \\
+    x_{i + 1} = x_i + v_i h \\
+    a_{i + 1} = f(x_{i+1}).
 \end{gather*}
@@
 Because we are looking at a system with gravity as the only force, the total energy should be conserved because gravity is a conservative force.
-This issue may not be noticeable for the first few orbits, but this will eventually cause orbits to deviate significantly from their original shape.
+This issue may not be noticeable for the first few orbits, but this will eventually (depending on the precision used) cause orbits to deviate significantly from their original shape.
 
@@
 \begin{align*}
-    x(i + 1) & = x(i) + v(i) h + a(i) h^2 \\
-    a(i + 1) & = f(x(i+1)) \\
-    v(i + 1) & = v(i) + \frac{h}{2}\left(a(i+1) + a(i)\right),
+    x_{i + 1} & = x_i + v_i h + a_i h^2 \\
+    a_{i + 1} & = f(x_{i+1}) \\
+    v_{i + 1} & = v_i + \frac{h}{2}\left(a_{i+1} + a_i\right),
 \end{align*}
@@
 and these values are calculated $n$ times with a given precision $h$.
-
-\subsection{Object-orientation}
 
@@
 \end{equation}
-where $G = 6.67408 \times 10^{-11} \ \textrm{N} \ \textrm{kg}^{-2} \textrm{m}^2$. The acceleration experienced by body 1 is then
+where $G = 6.67408 \times 10^{-11} \ \textrm{N} \ \textrm{kg}^{-2} \textrm{m}^2$. The direction of this force is then determined by the direction of the position vector:
 %
 \begin{equation*}
-    a_1 = \frac{F_G}{M_1} = \frac{G M_2}{r^2},
+    \Vec{e_r} = \Vec{r}/ r ,
 \end{equation*}
+%
+so that the force vector becomes
+%
+\begin{equation}
+    \Vec{F_G} = \frac{G M_1 M_2}{r^2}\Vec{e_r} = \frac{G M_1 M_2}{r^3}\Vec{r}.
+\end{equation}
+%
+The acceleration experienced by the body with mass $M_1$ is then
+%
+\begin{equation}\label{eq:accel}
+    \Vec{a} = \frac{\vec{F_G}}{M_1} = \frac{G M_2}{r^3}\Vec{r},
+\end{equation}
 %
@@
 In order to make our values easier to work with, we will scale the equations accordingly.
-We define the Sun to have a mass $M_S = M_\odot = 2\times 10^{30}$ kg and express the other masses (i.
-e.
-planets, moons) as a fraction of $M_\odot$.
+We define the Sun to have a mass $M_S = M_\odot = 2\times 10^{30}$ kg and express the other masses (such as planets, moons) as a fraction of $M_\odot$.
 Distance is measured in terms of astronomical units AU, where $1 \ \textrm{AU} = 1.496 \times 10^{11} \ \textrm{meters}$ and time is measured in years.
-This is appropriate for the scales we are looking at, since the Earth takes one year (surprise) to orbit the Sun once at an average distance of 1 AU.
+This is appropriate for the scales we are looking at, since the Earth takes one year to orbit the Sun once at an average distance of 1 AU.
 The gravitational constant can now be expressed as
@@
 \subsection{Our simplest model: Earth orbiting a stationary Sun}
+\label{subsec:simple}
 
@@
 \begin{equation*}
-    v_r = \frac{2\pi r}{T}.
+    v = \frac{2\pi r}{T}.
 \end{equation*}
@@
 
+If we do not want to introduce the period of an object in our calculations, we can instead calculate the velocity through Eq.\eqref{eq:accel}.
+In a perfectly circular orbit, the centripetal acceleration experienced by the orbiting body is $a = v^2/r$.
+This gives us
+%
+\begin{gather*}
+    \frac{v^2}{r} = \frac{GM_2}{r^2} \\
+    \Rightarrow \ v = \sqrt{\frac{G M_2}{r}},
+\end{gather*}
+%
+and in the case of $M_2 = M_\odot$ this gives
+%
+\begin{equation}\label{eq:circularorbit}
+    v = \sqrt{\frac{4\pi^2}{r}}.
+\end{equation}
+%
 In our algorithm we position the Sun in the origin of our coordinate system.
@@
 %
-The stability of our Euler and velocity Verlet algorithms are then tested using different time steps $\Delta t$.
+The stability of our Euler and velocity Verlet algorithms are then tested using different time steps $h$.
 By looking at the kinetic and potential energy of the system, we can also determine how well the two algorithms conserve energy.
-Total energy and angular momentum should be conserved in this system since gravity is the only force present, and because gravity is a conservative force we should not lose any energy.
-In theory, the Verlet algorithm should perform better (why?) than the Euler algorithm in this regard.
+Total energy and angular momentum should be conserved in this system since gravity is the only force present, and gravity is a conservative force.
+In theory, the Verlet algorithm should perform better than the Euler algorithm in this regard.
+This is because the Velocity Verlet algorithm is a second-order method, so the global error is already lower than the Euler method.
+In addition, the total energy in a system solved by the Verlet algorithm will oscillate around the true value of the energy \cite{bib:verlet}, while the Euler method can potentially add energy to the system indefinitely.
 
-After this, we will not be using Euler's method due to its lack of precision.
+For the next models, we will not be using the Euler method due to its lack of precision.
+
+\subsection{Escape velocity of the Earth}
+
+If the Earth were to have a high enough velocity, the planet would escape the gravitational influence of the Sun.
+The minimum velocity that achieves this is called the escape velocity, and we find the magnitude of this velocity by setting up the total energy before and after escaping.
+We find the potential energy of Earth through the expression for the gravitational force between Earth and the Sun:
+%
+\begin{equation*}
+    F_G = \frac{G M_\odot M_E}{r^2}.
+\end{equation*}
+%
+For conservative forces, the potential is defined as the integral of the force along a path. This gives us 
+%
+\begin{align*}
+    U & = \int F_G \ dr \\
+    & = \int \frac{G M_\odot M_E}{r^2} dr \\
+    & = -\frac{G M_\odot M_E}{r}.
+\end{align*}
+%
+The minimum velocity that the Earth needs to escape is found by setting the final velocity (after escaping) to zero.
+We can also set the potential energy to zero after the planet has escaped, since $r \rightarrow \infty$.
+The total energy before and after escaping is then
+%
+\begin{equation*}
+    \frac{1}{2}M_E v^2 -\frac{G M_\odot M_E}{r} = \frac{1}{2}M_E \cdot 0 - 0,
+\end{equation*}
+%
+so that 
+%
+\begin{equation}
+    v^2 =\frac{2G M_\odot}{r} \\
+\end{equation}
+%
+\begin{equation}\label{eq:esc_vel}
+     \Rightarrow \ v_{escape} = \sqrt{\frac{2G M_\odot}{r} } = \sqrt{\frac{4\pi^2}{r}}.
+\end{equation}
+%
+This analytical expression can be used to verify if our model is realistic, as we should see a planet escaping if we set its initial velocity to a value higher than $v_{escape}$.
+
+\subsection{Modified Newtonian gravity}
+\label{subsec:modified_gravity}
+
+Centripetal forces on the form $F \propto 1/r^2$ (such as Newtonian gravity) have the property that they allow for closed orbits \cite{bib:project3}.
+After one orbital period, an object with a closed orbit will end up at the same position it started.
+We modify the $r$-dependency in the model for Newtonian gravity (Eq.
+\eqref{eq:gforce}) so that
+%
+\begin{equation}
+    F_G = \frac{G M_1 M_2}{r^\beta}, \quad \beta \in \left[2,3\right].
+\end{equation}
+%
+We will study the effects of using this modified gravitational force, specifically in the model outlined in section \ref{subsec:simple}.
+Running a simulation of Earth orbiting the Sun with $\beta > 2$ should cause the Earth to no longer have a closed orbit.
+This effect should become more clear when $\beta \rightarrow 3$, as the gravitational force will weaken accordingly.
 
@@
 Introducing another planet means adding another term contributing to the total force on Earth.
-Jupiter will also experience this force, but the effect will be much less significant due the planet's high mass.
+Jupiter will also experience this force, but the effect will be much less significant due to the planet's high mass.
 The magnitude of this force is
@@
 \begin{equation*}
-    F_{\textrm{Earth-Jupiter}} = \frac{GM_{\textrm{Jupiter}} M_{\textrm{Earth}}}{r_{\textrm{Earth-Jupiter}}^2} = \frac{GM_E M_J}{r_{E-J}^2}
+    F_{\textrm{Earth-Jupiter}} = \frac{GM_{\textrm{Jupiter}} M_{\textrm{Earth}}}{r_{\textrm{Earth-Jupiter}}^2} = \frac{GM_E M_J}{r_{E-J}^2},
 \end{equation*}
@@
 This implies that the effects of Jupiter on Earth's orbit should be minimal.
-However, we will be scaling up the mass of Jupiter in order to examine the stability of our Verlet solver, as a higher Jupiter mass introduces stronger forces.
-This in turn causes less stability in the system.
+However, we will be scaling up the mass of Jupiter by a factor of 10 and 1000 in order to examine the stability of our Verlet solver.
+A higher Jupiter mass introduces stronger forces, which in turn makes the system less stable.
 
@@
 Until now we have not considered the gravitational effects from the planets on the Sun.
-In reality, the Sun will also gain a velocity of its own, though this velocity will be low compared to the planet velocities.
+In reality, the Sun will also gain a velocity of its own from gravitational interactions, though this velocity will be low compared to the planet velocities.
 For this reason it is best to center our reference system at the center of mass of the Solar System.
 We define the center of mass to be at the origin, then give the Sun an initial velocity that gives a total momentum of zero for the whole system.
-This ensures that the center of mass remains at the origin.
+This ensures that the center of mass remains at the origin, and the bodies in the Solar System will then have orbits centered around the origin.
 
-By using these conditions, we find the initial position of the Sun by defining the center of mass $R$ as:
+Calculating the initial values that put the center of mass at the origin is not necessary if we use NASA's tool \cite{bib:nasa} to find the appropriate initial conditions.
+However, we will perform the calculations for a simple Solar System with only the Sun, Earth and Jupiter present.
+
+We find the initial position of the Sun by defining the center of mass $R$ as:
 %
@@
 %
-\begin{equation}
+\begin{equation*}
     x_0 = -0.0049 \ \textrm{AU}, \quad y_0 = 0.
-\end{equation}
+\end{equation*}
 %
@@
 %
-Solving for $v_S$ this gives (again assuming that the Solar System lies along the x-axis):
+Solving for $v_S$ this gives (again assuming that the Solar System lies along the x-axis initially):
 %
@@
 
-\subsection{Escape velocity of the Earth}
+\subsection{Implementation of object-oriented code}
+The program for simulating the Solar System in its entirety consists of three classes and some external files to retrieve and output data. The classes are 
+SolarSystem, planet and solver.
 
-The escape velocity of an object in a gravitational field is found by setting up the total energy before and after escaping.
-If the Earth were to have a high enough velocity, the planet would escape the gravitational influence of the Sun.
-We consider a hypothetical scenario where the gravitational force between the Sun and Earth is defined by
+In the main file all the initial conditions position, velocity and mass are included. For each of these a planet object is made using the planet class.
 %
-\begin{equation*}
-    F_G = \frac{G M_\odot M_E}{r^\beta}, \quad \beta \in [2,3].
-\end{equation*}
-%
-For conservative forces, the potential is defined as the integral of the force along a path. This gives us 
-%
-\begin{align*}
-    U & = \int F_G \ dr \\
-    & = \int \frac{G M_\odot M_E}{r^\beta} dr \\
-    & = \frac{1}{1-\beta} \frac{G M_\odot M_E}{r^{\beta-1}}.
-\end{align*}
-%
-The minimum velocity that the Earth needs to escape is found by setting the final velocity (after escaping) to zero.
-We can also set the potential energy to zero after the planet has escaped, since $r \rightarrow \infty$.
-The total energy before and after escaping is then
-%
-\begin{equation*}
-    \frac{1}{2}M_E v^2 + \frac{1}{1-\beta} \frac{G M_\odot M_E}{r^{\beta-1}} = \frac{1}{2}M_E \cdot 0 + 0
-\end{equation*}
-%
-so that 
-%
-\begin{gather*}
-    v^2 = -\frac{2}{1-\beta} \frac{G M_\odot}{r^{\beta-1}} \\
-  \Rightarrow \ v_{escape} = \sqrt{ -\frac{1}{1 - \beta} \frac{8\pi^2}{r^{\beta - 1}} }.
-\end{gather*}
-%
-Since $\beta$ is always greater than 1 we avoid getting complex values.
-This lets us study the effects on the escape velocity by changing the strength of the gravitational force, as a higher value of $\beta$ corresponds to a weaker force.
+The SolarSystem class is initialized with empty (or predefined) vector of planet objects.
+To find the coordinates and velocities of the various planet objects over a given time we generate a object belonging to the solver class.
+This solver object has a method verlet that integrates every planet object in the SolarSystem object it takes with the Velocity Verlet.
+For each step of the Verlet method the acceleration is updated using the acceleration from the SolarSystem class.
 
@@
 
-The observed perihelion precession of Mercury is
+The observed perihelion precession of Mercury is an effect that cannot be explained by classical physics.
+It is instead explained as an effect of general relativity, a theory that postulates that any object with mass curves spacetime.
+This curvature is more significant close to massive objects, which is why general relativistic effects are more noticeable close to the Sun.
+This relativistic effect is implemented in our algorithm as a modification to the Newtonian gravitational force:
+%
+\begin{equation}\label{eq:precession}
+    F_G = \frac{G M_{S} M_M}{r^2}\left[1 + \frac{3l^2}{r^2 c^2}\right],
+\end{equation}
+%
+where $M_M$ is the mass of Mercury, $r$ is the distance between the Sun and Mercury and $l = \left| \Vec{r} \times \Vec{v}\right|$ is the magnitude of Mercury's orbital angular momentum.
+The speed of light in vacuum $c$ is written in terms of AU/yr, which gives
+%
+\begin{equation}
+    c = 63240.08034 \ \textrm{AU/yr}.
+\end{equation}
+%
+Over the course of a century, the perihelion of Mercury has been observed to precess 43'' due to general relativistic effects.
+Implementing Eq.\eqref{eq:precession} in our algorithm should demonstrate the same effect.
+Practically, this is done by running a simulation of Mercury orbiting the Sun for 100 years, which corresponds to the planet completing 415 orbits.
+For every orbit we find the lowest distance (perihelion) between Mercury and the Sun, and then calculate the perihelion angle $\theta_p$ using the coordinates of Mercury $(x_p, y_p)$ at that distance:
+%
+\begin{gather*}
+    \tan \theta_p = \frac{y_p}{x_p} \\
+    \Rightarrow \ \theta_p = \tan^{-1} \left(\frac{y_p}{x_p}\right).
+\end{gather*}
+%
+The value of $\theta_p$ after 100 simulated years can then be compared to the initial value of $\theta_p$. The initial position and velocity of Mercury in this simulation is 
+%
+\begin{equation*}
+    x = 0.3075 \ \textrm{AU}, \quad y = 0,
+\end{equation*}
+\begin{equation*}
+    v_x = 0, \quad v_y = 12.44 \ \textrm{AU/yr}.
+\end{equation*}
+%
+The reason that Eq.\eqref{eq:precession} introduces precession is that the force is no longer on the form $1/r^2$ (see section \ref{subsec:modified_gravity}).
+Forces that or on the this form allow for closed orbits, meaning that an object starting at a point in an orbit will end up at the same point after one orbital period.
+Adding the relativistic term will for this reason change the shape of Mercury's orbit.
 
@@
 
+\subsection{Our simplest model: Earth orbiting a stationary Sun}
+
+\begin{deluxetable}{lcc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results1}}
+\tablecomments{Table comparing CPU-time between the forward Euler and velocity Verlet algorithms, as a function of number of iterations, for the case of Earth revolving around the Sun, time complexity seemingly increases linearly for larger number of iterations.}
+\tablecolumns{3}
+\tablehead{Iterations  & Euler [s] & Verlet [s]}
+\startdata
+$10^{3}$ & 0.000572 & 0.000838 \\
+$10^{4}$ & 0.004041 & 0.006914 \\
+$10^{5}$ & 0.13305  & 0.239944 \\
+$10^{6}$ & 1.26058  & 2.39078  \\
+$10^{7}$ & 12.649   & 23.6886  \\
+$10^{8}$ & 125.378  & 236.845  
+\enddata
+\end{deluxetable}
+
+The following plots are made assuming that Earth starts 1AU away from the sun along the x-axis and has the initial velocity $v_{y} = 2\pi$ found from Eq.\eqref{eq:circularorbit} resulting in a circular orbit.
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3c-euler-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Orbit of Earth around the Sun for 2 years using Forward Euler method, this is not an energy conserving method, and our results show a significant drift in energy.}
+\label{fig:3c_euler}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3c-eulercromer-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Orbit of Earth around the Sun for 2 years using Euler-Cromer method, this is not an energy conserving method but it seemingly provides good results for higher resolutions (points per time).}
+\label{fig:3c_eulercromer}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3c-verlet-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Earth's orbit around the Sun for 2 years utilizing the Velocity Verlet algorithm, this method accounts for second order errors in the approximation.}
+\label{fig:3c_verlet}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=3c-energyEuler-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Variation in kinetic and potential energy over a two year period using the Forward Euler method with $1e3$ integration points.}
+\label{fig:energy_euler}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=3c-momentumEuler-n1e3-t2.pdf,width=\linewidth,clip=}}
+\caption{Variation in momentum over a two year period using the Forward Euler method with $1e3$ integration points.}
+\label{fig:mom_euler}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3c-energy-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Variation in kinetic and potential energy over a two year period using the Velocity Verlet method with $1e3$ integration points.}
+\label{fig:energy_verlet}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3c-momentum-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{Variation in momentum over a two year period using the Velocity Verlet method with $1e3$ integration points.}
+\label{fig:mom_verlet}
+\end{figure}
+
+\subsection{Escape velocity of the Earth}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3d-beta2'5-n1e4-t75.pdf,width=\linewidth,clip=}}
+
+\caption{A depiction of Earth orbiting the sun, where the $\frac{1}{r^{2}}$ dependency has been changed to incorporate $\frac{1}{r^{5/2}}$, a modification seemingly leading to a less stable orbit}
+\label{fig:beta2'5}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3d-beta3-n1e4-t75.pdf,width=\linewidth,clip=}}
+
+\caption{A depiction of Earth orbiting the sun, where the $\frac{1}{r^{2}}$ dependency has been changed to incorporate $\frac{1}{r^{3}}$, this modification seemingly results in Earth no longer being in a bound state in the Sun's gravity}
+\label{fig:beta3}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3d-escape-n1e3-t2.pdf,width=\linewidth,clip=}}
+
+\caption{The Earth with an initial velocity equal to the escape velocity of the Earth-Sun system.
+Using Eq.\eqref{fig:escape_vel} we found the escape velocity to be $v_{esc} = v_{0}\sqrt{2}$ where $v_{0} = 2\pi$ is the initial velocity required for circular orbit.}
+\label{fig:escape_vel}
+\end{figure}
+
+\subsection{Three body problem: Inlcuding Jupiter into our model}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3e-n1e6-t13.pdf,width=\linewidth,clip=}}
+
+\caption{A plot of the three body problem solved for stationary Sun.
+The figure represents the Earth and Jupiter orbiting the sun for 13 Earth years, with Jupiter's mass being unaltered the effect it has on Earth's orbit is minimal.}
+\label{fig:Jupiter1}
+\end{figure}
+
+\begin{figure}[t]
+
+\mbox{\epsfig{figure=3e-10m-n1e6-t13.pdf,width=\linewidth,clip=}}
+
+\caption{A plot of the three body problem solved for stationary Sun.
+The figure represents the Earth and Jupiter orbiting the sun for 13 Earth years, with Jupiter's mass being increased by a factor of 10 the effect it has on Earth's orbit is still not visible in the plot.}
+\label{fig:Jupiter10}
+\end{figure}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{3e-1000m-n1e3-t14.pdf}}
+\subfloat{(a) This plot is for for $n = 1000$ points, the stability of Verlet is very low and it is apparent that interaction between any of the bodies result in large numerical errors.}
+}\qquad
+{{\includegraphics[scale=0.53]{3e-1000m-n1e6-t13.pdf}}
+\subfloat{(b) This plot for $n = 10^{6}$ points, which is more stable.}
+}\qquad
+\caption{A plot of the three body problem solved for stationary Sun.
+The figure represents the Earth and Jupiter orbiting the sun for 13 Earth years, with Jupiter's mass being increased by a factor of 1000, we can see a drastic change in Earth's orbit.}
+\label{fig:Jupiter1000}
+\end{figure}
+
+\subsection{Final model: Including all planets and Pluto}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=3f-n1e6-t25_zoom.pdf,width=\linewidth,clip=}}
+\caption{Drift of COM of the Solar System when letting gravity affect the Sun. Avoiding a drift in the COM proved difficult, however the one shown in the figure was of order of magnitude $10^{-4}AU$.}
+\label{fig:3f_zoom}
+\end{figure}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{inner3planets.pdf}}
+\subfloat{(a) A 3d plot depicting the Sun and its three innermost planets in orbit for 1 Earth year.}
+}\qquad
+{{\includegraphics[scale=0.53]{solsys_full.pdf}}
+\subfloat{(b) A 3d plot depicting all planets in the Solar System in orbit for 250 Earth years ensuring that Pluto completes its orbit.}
+}\qquad
+\caption{Plots of the Solar System in two different spatial and timescales, depiction includes all planets finishing at least one orbit.}
+\label{fig:solsys}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=3g-arcseconds.pdf,width=\linewidth,clip=}}
+\caption{Perihelion precession of Murcery over 100 years.}
+\label{fig:perihelion}
+\end{figure}
+
 \section{Discussion}
 \label{sec:discussion}
+\subsection{Stability and CPU-time of Euler compared to Verlet}
+%
+As seen in table \ref{tab:results1}, time complexity of Velocity Verlet is seemingly close to double compared to time complexity of Euler.
+This however is consistent with the number of FLOPS used for each of the methods as our Verlet algorithm seemingly runs at close to double the amount of FLOPS compared to Euler.
+%
+The Forward Euler method is clearly spiraling outwards as seen in figure \ref{fig:3c_euler}.
+This is a consequence of poor energy conservation shown in figure \ref{fig:energy_euler}.
+Using Euler-Cromer the energy conservation is much better as seen in figure \ref{fig:3c_eulercromer}.
+With the same time step and and simulation time the orbit remains seemingly constant.
+The same conservation of energy is also apparent when using the Velocity Verlet method as seen in figure \ref{fig:energy_verlet} when computing for shorter periods of time.
+As seen in fig.\ref{fig:mom_euler} Verlet even conserves angular momentum for shorter time periods.
+%
+Closed elliptic orbits are special cases of centripetal forces proportional to $1/r^2$.
+In figure \ref{fig:beta2'5} there is a change from $F_G \propto 1/r^2$ to $F_G \propto 1/r^{5/2}$ giving unstable orbits.
+When the force is changed with a $1/r^3$ proportionality the planet spirals out.
+The planet is no longer in a bound state and will over a relatively short time escape the Sun in its entirety.
+
+\subsection{Stability of Verlet introducing multiple bodies}
+
+Adding another planetary body to our problem does not cause any visible destabilization in our Verlet method.
+However as the mass of this body increases beyond what is found among the planets of our Solar System the method destabilizes.
+
+Increasing the mass of Jupiter does not only destabilize the orbit of Earth, it also destabilizes our Verlet method as the probability of the Earth coming too close to the Sun increases.
+This leads to numerical artifacts as a result of our step size being too short to reasonably simulate the next step in the resulting interaction between the Sun and Earth.
+
+Pushing the three body problem further by letting gravity affect the Sun requires balancing of the total momentum done by moving the COM (centre of mass) to the center and setting the initial velocity so as to stop the COM from drifting.
+As seen in fig.\ref{fig:3f_zoom} the COM barely drifts when the Sun was in the initial condition explained above.
+However while the rest of the Planets will experience a similar drift the tiny order of magnitude prevents seeing any notable shifts in their orbits.
+
+Even with Velocity Verlet the energy is not conserved completely over long periods of time.
+In figure \ref{fig:solsys}(b) the full Solar System is simulated over a period of 250 years.
+In this time the orbits of the innermost planets in particular have an energy drift.
+To get near perfect orbits the time steps must be infinitesimally small.
+
+\subsection{Perihelion precession of Mercury}
+As a final test of the stability of our numerical method as well as a test of natural phenomena explained by general relativity, the sturdiness of our model is tested by simulating the orbit of Mercury.
+Newtons gravitational force ($F_G \propto \frac{1}{r^{2}}$) is unable to describe the perihelion precession of Mercury, so to expand the model the relativistic correction in to the gravitational force (Eq.
+\eqref{eq:precession}) is introduced.
+Analyzing our results for the precession of Mercury as seen in \ref{fig:perihelion} we get $26''$ as compared to the reference value of $43''$.
+This result is not impressive either when considering that the error can not be accounted for by numerical errors only.
 
@@
 \label{sec:conclusions}
+
+As demonstrated by the tests of numerical stability, the Velocity Verlet solver is clearly superior to the Euler solver when it comes to precision.
+When the Euler solver is used to compute the movement of a planet, the orbit quickly spirals out of its original shape.
+This shows that the Euler solver poorly conserves physical quantities such as the mechanical energy of the system.
+The Velocity Verlet method on the other hand conserves energy and momentum well with a slight trade off in the amount of FLOPS needed to preform the calculation.
+
+The Euler-Cromer method seemingly also conserves energy quite well, but becomes more unstable at fewer integration points.
+
+Lastly, even though our Verlet method failed the perihelion test it performed with overall reasonable stability when simulating the Solar System.
 
@@
 
-\begin{deluxetable}{lccc}
-%\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{Summary of main results.}
-\tablecolumns{4}
-\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
-\startdata
-Item 1 & Item 2 & Item 3 & Item 4
-\enddata
-\end{deluxetable}
-
 \begin{acknowledgements}
@@
 \begin{thebibliography}{}
-\bibitem{bib:lectures} Hjorth-Jensen, Morten, 2015, 
-\textit{Computational Physics - Lecture Notes Fall 2015}
+\bibitem{bib:project3} Hjorth-Jensen, Morten, 2018, \textit{Project 3} \url{http://compphysics.github.io/ComputationalPhysics/doc/
+Projects/2018/Project3/pdf/Project3.pdf} (24.10.18)
 
-\bibitem{bib:nasa} \textit{HORIZONS Web Interface}, retrieved from \url{https://ssd.jpl.nasa.gov/horizons.cgi#top} (22.10.18)
+\bibitem{bib:nasa} \textit{HORIZONS Web Interface} \url{https://ssd.jpl.nasa.gov/horizons.cgi} (22.10.18)
 
-\bibitem{1}
- Stig-Nicolai Foyn, Christer Dreierstad
-\textit{Project 1} \url{https://github.com/chdre/FYS3150/blob/master/Project1/report/FYS3150-Project1-stignicf-chrisdre.pdf} (01.10.18)
+\bibitem{bib:verlet} \textit{Verlet integration} \url{https://www.saylor.org/site/wp-content/uploads/2011/06/MA221-6.1.pdf} (24.10.18)
 \end{thebibliography}
 
-%\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
-  %Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
-  %A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89
-
 \end{document}

### 2018-11-15 Project4/main.cpp
Producing results

@@ -52,7 +52,7 @@ int main(int argc, char *argv[]){
         double TimeStart, TimeEnd;
         TimeStart = MPI_Wtime();
 
-        for(double T = 2.0; T <= 2.3; T+=0.02) {
+        for(double T = 2.0; T <= 2.3; T+=0.03) {
                 string temp = to_string(T);
                 string temp_filename = filename;
                 //temp_filename.append(temp); // unlock to make one file pr T

### 2018-11-15 Project4/plot.py

@@ -37,10 +37,7 @@ if sort == True:
 
     d = ({1,E_sorted[0]})
     j = 0
-    for i in range(0,len(E_sorted)):
-            if E_sorted[i+1] == E_sorted[j]:
-                d.update({E_sorted[j], E_sorted[j]})
-                j += 1
+    
 
 
 

### 2018-11-15 Project4/report/ms.tex
Adding tex file

--- 2018-10-22 LaTeX/template/ms.tex
@@
 \end{gather}
-In our case, $f$ is a function of position only.
+In our case, $a = f(x)$
 
@@
 
-\section{Data}
-\label{sec:data}
-
 \section{Results}
 \label{sec:results}
+
+\begin{deluxetable}{lcc}
+%\tablewidth{0pt}
+\tablecaption{\label{tab:results1}}
+\tablecomments{Table comparing CPU-time between the forward Euler and velocity Verlet algorithms, as a function of number of iterations, for the case of Earth revolving around the Sun, time complexity seemingly increases linearly for larger number of iterations.}
+\tablecolumns{3}
+\tablehead{Iterations  & Euler & Verlet}
+\startdata
+10^{3} & 0.000572 & 0.000838 \\
+10^{4} & 0.004041 & 0.006914 \\
+10^{5} & 0.13305 & 0.239944 \\
+10^{6} & 1.26058 & 2.39078 \\
+10^{7} & 12.649 & 23.6886 \\
+10^{8} & 125.378 & 236.845  \\
+\enddata
+\end{deluxetable}
+
+\section{Discussion}
+\label{sec:discussion}
+As seen in table \ref{tab:results1} time complexity of velocity Verlet is seemingly close to double compared to time complexity of Euler.
+This however is consistent with the number of FLOPS used for each of the methods as our Verlet algorithm seemingly runs at close to double the amount of FLOPS compared to Euler.
 

### 2018-11-15 Project4/report/ms.tex
Adding tex file

--- 2018-10-22 LaTeX/template/ms.tex
@@
 
-\author{Hans Erlend Bakken Glad, Torbjørn Lode Gjerberg, Christer Dreierstad, Stig-Nicolai Foyn}
+\author{}
 
@@
 
-Simulating the orbits of the planets in the Solar System is a suitable problem for making use of object-oriented code.
-While the two-body problem is straightforward to implement numerically, introducing multiple planets can quickly become tedious if the code is not written for this purpose.
-This is where object-orientation excels, as it is based on writing code once and then running it several times for different cases.
-In practice, this is done by writing classes in C++.
-
-Our procedure for creating an object-oriented code starts with solving the motion of Earth around the Sun with no object-orientation in mind.
-This gives us an idea of which parts of our code that is suited for object-orientation.
-The calculation of the orbits is performed with Euler's method and then the velocity Verlet method, which allows us to compare their stability and properties such as energy conservation.
-
-\section{Theory and method}
+\section{Method}
 \label{sec:method}
-a
-\subsection{Integration methods}
-The Euler method is a first-order method to approximate the solution to a differential equation.
-Given an analytical expression for the acceleration $a$ and an initial velocity $v(t)$ and position $x(t)$, the new values at a time $t + \Delta t$ can be found with
-
-\begin{gather}
-    v(t + \Delta t) = v(t) + a(t) \Delta t \\
-    x(t + \Delta t) = x(t) + v(t) \Delta t \\
-    a(t + \Delta t) = f( ... ).
-\end{gather}
-In our case, $f$ is a function of position only.
-
-\subsection{Object-orientation}
-a
-
-\subsection{Scaling the equations governing the Solar System}
-We will consider a model of our Solar System where gravity is the only force present.
-By Newton's law of gravitation, two celestial bodies with mass $M_1$ and $M_2$ separated by a distance $r$ will both experience a force with magnitude
-
-\begin{equation}\label{eq:gforce}
-    F_G = \frac{G M_1 M_2}{r^2},
-\end{equation}
-where $G = 6.67408 \times 10^{-11} \ \textrm{N} \ \textrm{kg}^{-2} \textrm{m}^2$. The acceleration experienced by body 1 is then
-\begin{equation}
-    a_1 = \frac{F_G}{M_1} = \frac{G M_2}{r^2},
-\end{equation}
-and similarly for body 2.
-
-If we consider a 'flat' Solar System with only two dimensions (naming the axes $x$ and $y$), the acceleration can be decomposed into
-
-\begin{equation}
-    a_x = \frac{G M_2}{r^2}\cdot\frac{\Delta x}{r}, \quad a_y = \frac{G M_2}{r^2}\cdot\frac{\Delta y}{r},
-\end{equation}
-where $\Delta_x$ and $\Delta_y$ are the differences in position coordinates $x$ and $y$ between the two bodies.
-This can easily be extended to include the third dimension (the 'height' of the Solar System) as well.
-
-In order to make our values easier to work with, we can scale the equations accordingly.
-We define the Sun to have a mass $M_\odot = 2\times 10^{30}$ kg and express the other masses (i.
-e.
-planets, moons) as a fraction of $M_\odot$.
-Distance is measured in terms of astronomical units AU, where $1 \ \textrm{AU} = 1.496 \times 10^{11} \ \textrm{meters}$ and time is measured in years.
-This is appropriate for the scales we are looking at, since the Earth takes one year (surprise) to orbit the Sun once at an average distance of 1 AU.
-The gravitational constant can now be expressed as
-
-\begin{equation}
-G = 4\pi^2 \ \frac{\textrm{AU}^3}{\textrm{yr}^{2}M} 
-\end{equation}
-where $M$ is the total mass in the system. From this we get the expression that will be used to calculate the acceleration $a$ induced by a body with mass $M$:
-
-\begin{equation}
-a = \frac{4\pi^2 M}{r^2}, \ \textrm{with} \ [a] = \frac{\textrm{AU}}{\textrm{yr}^2}.
-\end{equation}
-
-\subsection{Our simplest model: Earth orbiting a stationary Sun}
-
-To get an idea on how to produce a realistic model of the Solar System, we will first look at a simple model with only the Earth and the Sun present.
-Earth's orbit is set to be perfectly circular, while in reality it is slightly elliptical.
-The velocity of a circular orbit expressed in terms of the radius (or orbital distance) $r$ and period $T$ is
-
-\begin{equation}
-    v_r = \frac{2\pi r}{T}.
-\end{equation}
-For Earth, we have $r = 1 \ \textrm{AU}$ and $T = 1 \ \textrm{yr}$. Earth's orbital velocity is thus always $2\pi$ AU/yr.
-
-In our algorithm we position the Sun in the origin of our coordinate system.
-This further simplifies our calculations, as the distance between the Sun and Earth is simply defined by the coordinates $(x, y)$ of Earth.
-To test this simple case, we initialize the position of Earth at $x = 1 \ \textrm{AU}$ and $y = 0$.
-This means that the initial velocity in the $x$-direction has to be zero, otherwise we would not get a circular orbit.
-For a counter-clockwise orbit around the Sun, we then need to set the initial velocities to
-
-\begin{equation}
-    v_x = 0, \quad v_y = 2\pi.
-\end{equation}
-
-The stability of our Euler and velocity Verlet algorithms are then tested using different time steps $\Delta t$.
-By looking at the kinetic and potential energy of the system, we can also determine how well the two algorithms conserve energy.
-Total energy and angular momentum should be conserved in this system since gravity is the only force present, and because gravity is a conservative force we should not lose any energy.
-In theory, the Verlet algorithm should perform better (why?) than the Euler algorithm in this regard.
-
-After this, we will not be using Euler's method due to it's lack of precision.
-
-\subsection{Expanding our model: Adding Jupiter to the system}
-
-After object-orienting our code, adding more planets should only be a matter of introducing new objects that represent the physical parameters of the planets.
-Adding the planet Jupiter will let us study a three-body problem, though we will still keep the Sun motionless.
-Due to the significant mass of Jupiter ($\approx 300$ times more massive than Earth) we should see some changes in Earth's orbit.
-
-Introducing another planet means adding another term contributing to the total force on Earth.
-Jupiter will also experience this force, but the effect will be much less significant due the planet's high mass.
-The magnitude of this force is
-
-\begin{equation}
-    F_{\textrm{Earth-Jupiter}} = \frac{GM_{\textrm{Jupiter}} M_{\textrm{Earth}}}{r_{\textrm{Earth-Jupiter}}^2} = \frac{GM_E M_J}{r_{E-J}^2}
-\end{equation}
-where the masses of Earth and Jupiter are $M_{\textrm{E}} = 6 \times 10^{24}$ kg and $M_{\textrm{J}} = 1.9 \times 10^{27}$ kg.
-The distance between the two planets $r_{\textrm{E-J}}$ is calculated using the $x$ and $y$ coordinates of the planets,
-
-\begin{equation}
-    r_{\textrm{E-J}} = \sqrt{\left(x_E - x_j\right)^2 + \left(y_E - y_J\right)^2 } = \sqrt{(\Delta x_{E-J})^2 + (\Delta y_{E-J})^2}.
-\end{equation}
-Jupiter has an orbital radius of 5.2 AU with respect to the Sun, so that the distance between Earth and Jupiter can vary from 4.2 AU to 6.2 AU.
-Because $F \propto 1/r^2$, this means that the gravitational forces between the two planets vary by a factor
-
-\begin{equation}
-    \frac{1}{(4.2 \ \textrm{AU})^2} \bigg/  \frac{1}{(6.2 \ \textrm{AU})^2} \approx 2.2.
-\end{equation}
-At a distance of 4.2 AU away from Earth, the forces between the two are as high as possible. Comparing this to the force exerted by the Sun on Earth, we get
-
-\begin{equation}
-    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{G M_E M_{Sun}/r_{Earth-Sun}^2}{G M_E M_J/r_{Earth-Jupiter}^2} = \frac{M_{Sun} \ r_{Earth-Jupiter}^2}{M_J \ r_{Earth-Sun}^2}
-\end{equation}
-Using that $M_{J} = 9.5 \times 10^{-4} M_\odot$, we get
-
-\begin{equation}
-    \frac{F_{Earth-Sun}}{F_{Earth-Jupiter}} = \frac{M_\odot \cdot (4.2 \ \textrm{AU})^2}{9.5 \times 10^{-4} M_\odot \cdot (1 \ \textrm{AU})^2} \approx 18600.
-\end{equation}
-The gravitational force from Jupiter is around 18600 times weaker than the gravitational force from the Sun, even at the lowest possible Earth-Jupiter distance.
-This implies that the effects of Jupiter on Earth's orbit should be minimal.
-However, we will be scaling up the mass of Jupiter in order to examine the stability of our Verlet solver.
-
-\subsection{Final model: Include all planets and Pluto}
-
-Our final model of the Solar System will include all planets and the dwarf planet Pluto.
-Additionally, we will allow the gravitational attraction from the planets to also affect the Sun.
-We define the origin to be the center of mass of the system, then give the Sun an initial velocity that gives a total momentum of zero for the whole system.
-This ensures that the center of mass remains at the origin.
-
-\section{Data}
-\label{sec:data}
 
@@
 \label{sec:results}
+
+\section{Discussion}
+\label{sec:discussion}
 

### 2018-11-17 Project4/main3.cpp
Adding results

@@ -51,7 +51,7 @@ int main(int argc, char *argv[]){
         double TimeStart, TimeEnd;
         TimeStart = MPI_Wtime();
 
-        for(double T = 2.0; T <= 2.3; T+=0.01) {
+        for(double T = 2.0; T <= 2.3; T+=0.015) {
                 Metropolis(L, mcc, T, energies(T), numprocs, my_rank, myloop_end, myloop_begin, filename);
         }
 

### 2018-11-17 Project4/plot.py

@@ -86,39 +86,39 @@ if acceptOfT == True:
 
 
 if vary == True:
-    file = "data/test/L40-1e5.txt"
-    file2 = "data/test/L60-1e5.txt"
-    file3 = "data/test/L80-1e5.txt"
-    file4 = "data/test/L100-1e5.txt"
+    file = "data/test/L40-1e6.txt"
+    file2 = "data/test/L60-1e6.txt"
+    #file3 = "data/test/L80-1e5.txt"
+    #file4 = "data/test/L100-1e5.txt"
 
     E, Mabs, C_V, chi, T = np.loadtxt(file, usecols=(0,1,2,3,4), unpack=True)
     E2, Mabs2, C_V2, chi2 = np.loadtxt(file2, usecols=(0,1,2,3), unpack=True)
-    E3, Mabs3, C_V3, chi3 = np.loadtxt(file3, usecols=(0,1,2,3), unpack=True)
-    E4, Mabs4, C_V4, chi4 = np.loadtxt(file4, usecols=(0,1,2,3), unpack=True)
+    #E3, Mabs3, C_V3, chi3 = np.loadtxt(file3, usecols=(0,1,2,3), unpack=True)
+    #E4, Mabs4, C_V4, chi4 = np.loadtxt(file4, usecols=(0,1,2,3), unpack=True)
 
     plt.figure()
-    plt.plot(T, E, T, E2, T, E3, T, E4)
+    plt.plot(T, E, T, E2)#, T, E3, T, E4)
     plt.legend(["L=40", "L=60", "L=80", "L=100"], prop={'size':15})
     plt.title('Mean energy for various lattice sizes', size=15)
     plt.xlabel('T [kT/J]', size=15); plt.ylabel('$\\langle E \\rangle/L^2$', size=15)
     plt.show()
 
     plt.figure()
-    plt.plot(T, Mabs, T, Mabs2, T, Mabs3, T, Mabs4)
+    plt.plot(T, Mabs, T, Mabs2)#, T, Mabs3, T, Mabs4)
     plt.legend(["L=40", "L=60", "L=80", "L=100"], prop={'size':15})
     plt.title('Mean value of magnetic moment for various lattice sizes', size=15)
     plt.xlabel('T [kT/J]', size=15); plt.ylabel('$\\langle|M|\\rangle/L^2$', size=15)
     plt.show()
 
     plt.figure()
-    plt.plot(T, C_V, T, C_V2, T, C_V3, T, C_V4)
[…]

### 2018-11-17 Project4/report/ms.tex

@@
 
-\title{Project 3}
+\title{Project 4}
 
-\author{}
+\author{Christer Dreierstad}
 
-\email{ola.d.nordmann@astro.uio.no}
+\email{chrisdre@student.matnat.uio.no}
 
@@
 \end{abstract}
-\keywords{planetary movement --- cosmology: observations --- methods: velocity verlet, forward euler}
+\keywords{computational science: Ising model  --- methods: Metropolis, Monte Carlo}
 
@@
 \label{sec:introduction}
+Studying the phase trans
 

### 2018-11-17 Project4/plot.py
Adding new results, stashing old ones

@@ -86,39 +86,39 @@ if acceptOfT == True:
 
 
 if vary == True:
-    file = "data/test/L40-1e6.txt"
-    file2 = "data/test/L60-1e6.txt"
-    #file3 = "data/test/L80-1e5.txt"
-    #file4 = "data/test/L100-1e5.txt"
+    file = "data/test/L40-1e6-ext.txt"
+    file2 = "data/test/L60-1e6-ext.txt"
+    file3 = "data/test/L80-1e6-ext.txt"
+    file4 = "data/test/L100-1e6-ext.txt"
 
     E, Mabs, C_V, chi, T = np.loadtxt(file, usecols=(0,1,2,3,4), unpack=True)
     E2, Mabs2, C_V2, chi2 = np.loadtxt(file2, usecols=(0,1,2,3), unpack=True)
-    #E3, Mabs3, C_V3, chi3 = np.loadtxt(file3, usecols=(0,1,2,3), unpack=True)
-    #E4, Mabs4, C_V4, chi4 = np.loadtxt(file4, usecols=(0,1,2,3), unpack=True)
+    E3, Mabs3, C_V3, chi3 = np.loadtxt(file3, usecols=(0,1,2,3), unpack=True)
+    E4, Mabs4, C_V4, chi4 = np.loadtxt(file4, usecols=(0,1,2,3), unpack=True)
 
     plt.figure()
-    plt.plot(T, E, T, E2)#, T, E3, T, E4)
+    plt.plot(T, E, T, E2, T, E3, T, E4)
     plt.legend(["L=40", "L=60", "L=80", "L=100"], prop={'size':15})
     plt.title('Mean energy for various lattice sizes', size=15)
     plt.xlabel('T [kT/J]', size=15); plt.ylabel('$\\langle E \\rangle/L^2$', size=15)
     plt.show()
 
     plt.figure()
-    plt.plot(T, Mabs, T, Mabs2)#, T, Mabs3, T, Mabs4)
+    plt.plot(T, Mabs, T, Mabs2, T, Mabs3, T, Mabs4)
     plt.legend(["L=40", "L=60", "L=80", "L=100"], prop={'size':15})
     plt.title('Mean value of magnetic moment for various lattice sizes', size=15)
     plt.xlabel('T [kT/J]', size=15); plt.ylabel('$\\langle|M|\\rangle/L^2$', size=15)
     plt.show()
 
     plt.figure()
[…]

### 2018-11-17 Project4/report/ms.tex
Updates from ShareLaTeX

@@
 \label{sec:introduction}
-Studying the phase trans
+To study the phase transitions in a magnetic system we consider the Ising model for a system of finite lattice size. The energy of the system, by the Ising model, is 
+%
+\begin{gather*}
+    E = -J\sum_{< kl >}^N s_k s_l,
+\end{gather*}
+%
+when there is no external magnetic field applied to the system. For solving this system we consider the Metropolis algorithm, which is a Markov chain Monte Carlo method.
 
@@
 \label{sec:results}
+\begin{figure}[t]
+
+c)
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+\begin{figure}
+\mbox{\epsfig{figure=EofMCC-GS-T1-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:E(mcc)T1}
+\end{figure}
+
+MofMCC-GS-T1-L20-1e5.pdf
+
+EofMCC-GS-T2_4-L20-1e5.pdf
+MofMCC-GS-T2_4-L20-1e5.pdf
+
+EofMCC-T1-L20-1e6.pdf
+MofMCC-T1-L20-1e6.pdf
+
+EofMCC-T2_4-L20-1e6.pdf
+MofMCC-T2_4-L20-1e6.pdf
+
+accepts_of_T-T1T2_4-L20.pdf
+accepts-T1T2_4-L20-1e5.pdf
+CV-T2-23.pdf
+CV-T22-24.pdf
+E-T2-23.pdf
+E-T22-24.pdf
+FittedTc.pdf
+M-T2-23.pdf
+M-T22-24.pdf
+PE-T1-L20-1e6.pdf
+PE-T2_4-L20-1e6.pdf
+X-T2-23.pdf
+X-T22-24.pdf
+
+\begin{figure}
+\mbox{\epsfig{figure=,width=\linewidth,clip=}}
+\caption{text}
+\label{fig:figure_label}
+\end{figure}
 

### 2018-11-17 Project4/Tc.py
Adding results

@@ -3,10 +3,10 @@ import matplotlib.pyplot as plt
 import numpy.linalg
 
 
-file = "data/L40-1e6.txt"
-file2 = "data/L60-1e6.txt"
-file3 = "data/L80-1e6.txt"
-file4 = "data/L100-1e6.txt"
+file = "data/test/L40-1e6-vary.txt"
+file2 = "data/test/L60-1e6-vary.txt"
+file3 = "data/test/L80-1e6-vary.txt"
+file4 = "data/test/L100-1e6-vary.txt"
 
 
 chi, T = np.loadtxt(file, usecols=(3,4), unpack=True)

### 2018-11-17 Project4/plot.py

@@ -4,19 +4,19 @@ import os
 
 accepts = False
 acceptOfT = False
-EM = False
+EM = True
 PE = False
-vary = True
+vary = False
 variance = False
 
 if EM == True:
     file = "data/L20-GS-T1-1e6.txt"
     file2 = "data/L20-GS-T2_4-1e6.txt"
 
-    E = 20*20*np.loadtxt(file, usecols=0)
-    M = 20*20*np.loadtxt(file, usecols=1)
-    E2 = 20*20*np.loadtxt(file2, usecols=0)
-    M2 = 20*20*np.loadtxt(file2, usecols=1)
+    E = np.loadtxt(file, usecols=0)
+    M = np.loadtxt(file, usecols=1)
+    E2 = np.loadtxt(file2, usecols=0)
+    M2 = np.loadtxt(file2, usecols=1)
     mcc = np.linspace(0,len(M),len(M))
 
     plt.figure()
@@ -125,8 +125,8 @@ if vary == True:
     plt.show()
 
 if variance == True:
-    file = "data/varianceE-L20-T1-1e6.txt"
-    file2 = "data/varianceE-L20-T2_4-1e6.txt"
+    file = "data/L20-T1-1e6.txt"
+    file2 = "data/L20-T2_4-1e6.txt"
 
     var = np.loadtxt(file, usecols=4)
     var2 = np.loadtxt(file2, usecols=4)

### 2018-11-17 Project4/main.cpp
Results

@@ -109,8 +109,8 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
         for(int m = myloop_begin; m <= myloop_end; m++) {
                 for (int x = 0; x < L; x++) {
                         for (int y = 0; y < L; y++) {
-                                int xr = RNGPos(generator); // indices for random element int) (RNG(generator)*(double) n); //
-                                int yr = RNGPos(generator); // (int) (RNG(generator)*(double) n); // RNGSpin(generator);//;
+                                int xr = RNGPos(generator); // indices for random element
+                                int yr = RNGPos(generator);
 
                                 int dE = 2.0*S(xr,yr)*(S(xr,PB(yr,L,1)) + S(xr,PB(yr,L,-1))
                                                        + S(PB(xr,L,1),yr) + S(PB(xr,L,-1),yr));

### 2018-11-17 Project4/plot.py

@@ -4,9 +4,9 @@ import os
 
 accepts = False
 acceptOfT = False
-EM = True
+EM = False
 PE = False
-vary = False
+vary = True
 variance = False
 
 if EM == True:

### 2018-11-17 Project4/report/ms.tex
Updates from ShareLaTeX

--- 2018-11-17 Project4/report/ms.tex
@@
 \label{sec:introduction}
-Studying the phase trans
+To study the phase transitions of a magnetic system we consider the Ising model for a two-dimensional system with a finite lattice size.
+The particles (spins) that make up the crystal (lattice) can take two values representing spin up or down.
+The energy of the lattice can then, by the Ising model, be found by
+%
+\begin{gather*}
+    E = -J\sum_{< kl >}^N s_k s_l,
+\end{gather*}
+%
+when there is no external magnetic field applied to the system.
+The coupling constant J will be set to 1, which means we are studying a ferromagnetic interaction between the spins that make up the lattice.
+Since the coupling constant is dependent upon the material we study, setting this to 1 allows for a general model.
+For studying the phase transition we consider the  Metropolis algorithm, a Markov chain Monte Carlo method (MCMC).
 
@@
 \label{sec:method}
+h
+
+The two-dimensional binary Ising model has analytical expectation values.
+
+\begin{deluxetable}{cccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{....}
+\tablecolumns{4}
+\tablehead{Spins up  & Degeneracy & Energy & Magnetic moment}
+\startdata
+4 & 1 & -8J & 4 \\
+3 & 4 & 0 & 2 \\
+2 & 4 & 0 & 0 \\
+2 & 2 & 8J & 0 \\
+1 & 4 & 0 & -2 \\
+0 & 1 & -8J & -4
+\enddata
+\end{deluxetable}
 
@@
 \label{sec:results}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{EofMCC-GS-T1-L20-1e5.pdf}}
+\subfloat{(a) hva skjer}
+}\qquad
+{{\includegraphics[scale=0.53]{MofMCC-GS-T1-L20-1e5.pdf}}
+\subfloat{(b) }
+}\qquad
+\caption{text}
+\label{fig:GS}
+\end{figure}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{EofMCC-GS-T2_4-L20-1e5.pdf}}
+\subfloat{(a) }
+}\qquad
+{{\includegraphics[scale=0.53]{MofMCC-GS-T2_4-L20-1e5.pdf}}
+\subfloat{(b) }
+}\qquad
+\caption{text}
+\label{fig:GS}
+\end{figure}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{EofMCC-T1-L20-1e6.pdf}}
+\subfloat{(a) }
+}\qquad
+{{\includegraphics[scale=0.53]{MofMCC-T1-L20-1e6.pdf.pdf}}
+\subfloat{(b) }
+}\qquad
+\caption{text}
+\label{fig:GS}
+\end{figure}
+
+\begin{figure}[H]
+{{\includegraphics[scale=0.53]{EofMCC-T2_4-L20-1e6.pdf}}
+\subfloat{(b) }
+}\qquad
+{{\includegraphics[scale=0.53]{MofMCC-T2_4-L20-1e6.pdf}}
+\subfloat{(c) }
+}\qquad
+\caption{text}
+\label{fig:GS}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=CV-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=E-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=E-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=FittedTc.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=M-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=M-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=X-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
+
+\begin{figure}[t]
+\mbox{\epsfig{figure=X-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Description of figure -- explain all elements, but do not
+draw conclusions here.}
+\label{fig:figure_label}
+\end{figure}
 

### 2018-11-17 Project4/report/ms.tex
Updates from ShareLaTeX

@@
 \label{sec:method}
-h
+The two-dimensional binary Ising model has analytical expectation values. The partition function
+%
+\begin{equation*}
+    z = \sum_i \Omega(E_i) e^{\beta E_i},
+\end{equation*}
+%
+where $\Omega(E_i)$ is the degeneracy of energy $E_i$ and $\beta = 1/k_BT$.
+From Table \ref{tab:energies} we can see that there are 2 energies that are non-zero: $\pm 8J$, which both has a degeneracy of 2.
+The energy equal to zero has a degeneracy of 12.
+Expanding the sum of the partition function we get
+%
+\begin{align*}
+    z &= 2e^{8J\beta} + 12 + 2e^{-8J\beta} \\
+    &= 12 + 2\left(e^{8J\beta} + e^{-8J\beta}\right),
+\end{align*}
+%
+where we have inserted $e^0 = 1$. We know that $\cosh(x) = \frac{1}{2}\left( e^{-x} + e^x\right)$, inserting this gives
+%
+\begin{gather}\label{eq:z}
+    Z = 12 + 4\cosh(8J\beta).
+\end{gather}
+%
+The mean value of the energy is given by
+%
+\begin{align*}
+    \langle E \rangle = -\frac{\partial }{\partial \beta} \ln Z = -\frac{1}{Z}\frac{\partial Z}{\partial \beta},
+\end{align*}
+%
+which is found by inserting the partition function.
+%
+\begin{align*}
+    \langle E \rangle &= -\frac{1}{Z}\frac{\partial}{\partial \beta} \left(12 + 4 \cosh\left(8J\beta\right)\right)\\
+    &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right),
+\end{align*}
+pulling a factor 4 from the partition function and inserting we are left with a analytical solution to the mean energy of the lattice
+%
+\begin{gather*}
+    \langle E \rangle = -\frac{8J\sinh\left(8J\beta\right)}{3 + \cosh(8J\beta}.
+\end{gather*}
+%
+The heat capacity can be found by
+%
+\begin{align*}
+    C_V &= \frac{1}{kT^2}\frac{\partial^2}{\partial \beta^2} \ln Z = \frac{1}{kT^2}\frac{\partial}{\partial \beta}\left(\frac{\partial}{\partial \beta} \ln Z\right),
+\end{align*}
+%
+where we recognize the partial derivative in the parenthesis as the negative mean energy. 
+%
+\begin{align*}
+    C_V = \frac{1}{kT^2}&\left(\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right \\
+    &-\left \frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right),
+\end{align*}
+%
+by inserting for the mean energy and using the product rule. Further simplifications gives the mean heat capacity
+%
+\begin{gather*}
+    C_V = \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}.
+\end{gather*}
+The susceptibility is given as
+%
+\begin{equation*}
+    \chi = \frac{1}{kT}\left(\langle M^2 \rangle - \langle M \rangle^2\right).
+\end{equation*}
+%
+To find the susceptibility of the system we consider the mean magnetic moment
+%
+\begin{equation*}
+    \langle M \rangle = \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i},
+\end{equation*}
+where we find the magnetic moment for a given energy in Table \ref{tab:energies}. So the mean magnetic moment is
+%
+\begin{equation*}
+    \langle M \rangle = \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(-2e^0\right) + -4e^{8J\beta} \right) = 0.
+\end{equation*}
+%
+Further we calculate the mean of the  magnetic moment squared
+%
+\begin{equation*}
+    \langle M^2 \rangle = \frac{1}{Z}\sum_i M_i^2 e^{-\beta E_i},
+\end{equation*}
+and again referring to Table \ref{tab:energies} we get
+%
+\begin{align*}
+    \langle M^2 \rangle &= \frac{1}{Z}\left(16\left(2e^{8J\beta}\right) + 16\left(2e^0\right) \right) \\
+    &= \frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
+\end{align*}
+So the susceptibility of the system is given as
+%
+\begin{equation*}
+    \chi = \frac{1}{kT^2}\langle M^2 \rangle = \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}
+\end{equation*}
+%
+Further we find the mean of the absolute value of the magnetic moment by
 
-The two-dimensional binary Ising model has analytical expectation values.
+\begin{align*}
+    \langle |M| \rangle &= \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i} \\
+     &= \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(|-2|e^0\right) + |-4|e^{8J\beta} \right) \\
+     &= \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)},
+\end{align*}
+%
+after pulling out a factor 4 from the partition function. To summarize we have the following analytical expression for the system
+%
+\begin{align*}
+    \langle E \rangle &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right), \\
+    \langle |M| \rangle &= \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)}, \\
+    C_V &= \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}, \\
+    \chi &= \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
+\end{align*}
 
@@
 \tablewidth{0pt}
-\tablecaption{\label{tab:results}}
-\tablecomments{....}
+\tablecaption{\label{tab:energies}}
+\tablecomments{...}
 \tablecolumns{4}

### 2018-11-17 Project4/analytical.py
Adding results

@@ -4,7 +4,8 @@ n = 2
 T = 1.0
 
 E = -8.0*np.sinh(8/T)/(3.0 + np.cosh(8/T))
-CV = 1/T**2*(64.0*np.cosh(8/T)*(3 + np.cosh(8/T)) - 64*np.sinh(8/T)**2)/(3+np.cosh(8/T))**2
+#CV = 1/T**2*(64.0*np.cosh(8/T)*(3 + np.cosh(8/T)) - 64*np.sinh(8/T)**2)/(3+np.cosh(8/T))**2
+CV = (64*np.cosh(8.0/T))/(T**2*(3 + np.cosh(8.0/T)))
 Mabs = (2*np.exp(8/T) + 4.0)/(3.0 + np.cosh(8/T))
 M2 = 8.0*(np.exp(8/T) + 1.0)/(3.0 + np.cosh(8/T))
 chi = M2/T

### 2018-11-17 Project4/main.cpp

@@ -96,7 +96,6 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
         vec ExpectVals = zeros<vec>(5);
         vec TotalExpectVals = zeros<vec>(5);
         mat S = zeros<mat>(L,L);
-        mat EnergyMagSave = zeros<mat>(mcc,2);     // for storing energy and magnetic moment values
 
         // initializing lattice
         initialize(energy, magMoment, L, S, 0, generator);

### 2018-11-18 Project4/report/ms.tex

@@
 \usepackage{epsfig,floatflt}
+\usepackage{url}
+\usepackage{subfig}
+\setcitestyle{square}
 
@@
 %
-\begin{gather*}
+\begin{gather}\label{eq:energyIsing}
     E = -J\sum_{< kl >}^N s_k s_l,
-\end{gather*}
+\end{gather}
 %
@@
 Since the coupling constant is dependent upon the material we study, setting this to 1 allows for a general model.
-For studying the phase transition we consider the  Metropolis algorithm, a Markov chain Monte Carlo method (MCMC).
+For studying the phase transition we use a Markov chain Monte Carlo method (MCMC), and the probability of finding the system in a given state $i$ is
+%
+\begin{align*}
+    P(E_i) = \frac{e^{-\beta E_i}}{Z},
+\end{align*}
+%
+where Z is the partition function $Z = \sum_i e^{-\beta E_i}$, which is impossible to calculate unless we know all the states.
+To eliminate the partition function we will use the Metropolis algorithm, which eliminates the partition function.
+When applying the algorithm we will consider periodic boundary conditions of the lattice, which will allow us to simulate an infinitely large lattice, and we do not have to consider what happens on the boundary the lattice (the ends of the material).
+Increasing the size of the lattice will further improve the approximation of a large scale lattice.
+The goal of the MCMC simulation is to observe the energy and the magnetic moment as time (MC cycles) increase, we expect the system to stabilize in the most probable state.
 
 \section{Method}
+\subsection{Analytical expressions}
 \label{sec:method}
+\begin{deluxetable}{cccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:energies}}
+\tablecomments{States of the 2x2 lattice and corresponding values for energy and magnetic moment. Notice that the ground state (the lowest energy) has either all spins pointing up or down.}
+\tablecolumns{4}
+\tablehead{Spins up  & Degeneracy & Energy & Magnetic moment}
+\startdata
+4 & 1 & -8J & 4 \\
+3 & 4 & 0 & 2 \\
+2 & 4 & 0 & 0 \\
+2 & 2 & 8J & 0 \\
+1 & 4 & 0 & -2 \\
+0 & 1 & -8J & -4
+\enddata
+\end{deluxetable}
+
 The two-dimensional binary Ising model has analytical expectation values. The partition function
@@
 
-\begin{deluxetable}{cccc}
-\tablewidth{0pt}
-\tablecaption{\label{tab:energies}}
-\tablecomments{...}
-\tablecolumns{4}
-\tablehead{Spins up  & Degeneracy & Energy & Magnetic moment}
-\startdata
-4 & 1 & -8J & 4 \\
-3 & 4 & 0 & 2 \\
-2 & 4 & 0 & 0 \\
-2 & 2 & 8J & 0 \\
-1 & 4 & 0 & -2 \\
-0 & 1 & -8J & -4
-\enddata
-\end{deluxetable}
+\subsection{Ising model}
+When the energy of the Ising model we consider the equation given in the introduction, Eq.
+\eqref{eq:energyIsing}, where the subscript of the sum means we are taking the sum over the neighbouring spins, N is the total amount of spins and $s_k$, $s_l = \pm 1$, represents the spin of the particles.
+Consider a random spin at the lattice, in the two-dimensional case it has a total of four neighbours that it interacts with.
+An example for a spin surrounded by four spin up.
+%
+\begin{equation*}
+    \begin{matrix}
+    & \uparrow & \\
+    \uparrow & \uparrow & \uparrow \\
+    & \uparrow & \\
+    \end{matrix},
+   % \quad
+    %\longrightarrow
+    %\quad
+    %\begin{matrix}
+    %& \uparrow & \\
+    %\uparrow & \uparrow & \uparrow \\
+    %& \uparrow & \\
+    %\end{matrix}
+\end{equation*}
+%
+with energy $E = -4J$.
+Changing one of the spins in the example above results in a new configuration with a new energy and magnetic moment.
+Following are five examples that show the different energies $E = 0, \pm 2J, \pm 4J$, and the resulting energy when flipping the spin we are located at:
+
+\begin{align*}
+    E = -4J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \uparrow & \uparrow & \uparrow \\
+    & \uparrow & \\
+    \end{matrix}
+    \quad
+    \Longrightarrow
+    \quad
+    E = 4J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \uparrow & \downarrow & \uparrow \\
+    & \uparrow & \\
+    \end{matrix} \\
+\end{align*}
+where the change of energy $\Delta E = E_{after} - E_{before} = 8J$.
+\begin{align*}
+    E = -2J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \uparrow & \uparrow \\
+    & \uparrow & \\
+    \end{matrix}
+    \quad
+    \Longrightarrow
+    \quad
+    E = 2J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \downarrow & \uparrow \\
+    & \uparrow & \\
+    \end{matrix}
+\end{align*}
+with $\Delta E = 4J$.
+\begin{align*}
+    E = 0
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \uparrow & \uparrow \\
+    & \downarrow & \\
+    \end{matrix}
+    \quad
+    \Longrightarrow
+    \quad
+    E = 0
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \downarrow & \uparrow \\
+    & \downarrow & \\
+    \end{matrix}
+\end{align*}
+with $\Delta E = 0$.
+\begin{align*}
+    E = -2J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \uparrow & \downarrow \\
+    & \downarrow & \\
+    \end{matrix}
+    \quad
+    \Longrightarrow
+    \quad
+    E = 2J
+    \quad
+    \begin{matrix}
+    & \uparrow & \\
+    \downarrow & \downarrow & \downarrow \\
+    & \downarrow & \\
+    \end{matrix}
+\end{align*}
+with $\Delta E = -4J$.
+\begin{align*}
+    E = 4J
+    \quad
+    \begin{matrix}
+    & \downarrow & \\
+    \downarrow & \uparrow & \downarrow \\
+    & \downarrow & \\
+    \end{matrix}
+    \quad
+    \Longrightarrow
+    \quad
+    E = -4J
+    \quad
+    \begin{matrix}
+    & \downarrow & \\
+    \downarrow & \downarrow & \downarrow \\
+    & \downarrow & \\
+    \end{matrix}
+\end{align*}
+with $\Delta E = -8J$. We now see that for a random spin of the lattice we have known values for the change of energy $\Delta E$.
+
+\subsection{Algorithm}
+For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment.
+We then iterate over the lattice at random positions (i.e.
+we check $L^2$ random spins), for each of these positions we calculate the change in energy $\Delta E$.
+Consider the ratio of the probability of the new energy $P\left(E_{new}\right)$ and the previous energy $P\left(E_{prev}\right)$:
+%
+\begin{align*}
+    \frac{P\left(E_{new}\right)}{P\left(E_{prev}\right)} = \frac{e^{-\beta E_{new}}/Z}{e^{-\beta E_{prev}}/Z} = e^{-\beta\left(E_{new} - E_{prev}\right)} = e^{-\beta\Delta E},
+\end{align*}
+%
+where Z is the the partition function. We now perform the Metropolis test for a given $\Delta E$;
+%
+\begin{equation*}
+    r \leq e^{-\beta \Delta E}, \qquad r \in [0,1]
+\end{equation*}
+%
+i.e.
+if the ratio is larger than or equal to a random number $r \in [0,1]$ we accept a new configuration, and flip the spin.
+We are looking at the change in energy if we were to flip a random spin; if the change of energy is large, the likelihood of this spin being flipped is small, so we reject this solution, on the other hand, if the probability is large, we accept the configuration.
+Since we expect the system to move to more likely state, we are looking for states that does not increase the systems energy.
+A large change in energy means we are looking at a spin that has half or more of its surrounding spins in the same configuration, and the system seems to locally move towards a steady state, therefore the local system is not changed.
+For each such Metropolis test that succeeds we add the flip of spin as a contribution to the systems energy and magnetic moment.
+When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment.
+
+PBC
 
@@
 
-\begin{figure}[H]
-{{\includegraphics[scale=0.53]{EofMCC-GS-T1-L20-1e5.pdf}}
-\subfloat{(a) hva skjer}
-}\qquad
-{{\includegraphics[scale=0.53]{MofMCC-GS-T1-L20-1e5.pdf}}
-\subfloat{(b) }
-}\qquad
-\caption{text}
-\label{fig:GS}
+\begin{figure}[t]
+\subfloat[Energy pr. spin.]{
+  \includegraphics[width=9cm]{EofMCC-GS-T1-L20-1e5.pdf}
+  \label{fig:L20-GS-E-T1}}\\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{
+  \includegraphics[width=9cm]{MofMCC-GS-T1-L20-1e5.pdf}
+  \label{fig:evaluation:L20-GS-M-T1}}
+\caption{Time evolution (MC cycles) of energy and magnetic moment of an ordered initial state with temperature 1 kT/J. $10^5$ Monte Carlo cycles.}
 \end{figure}
 
-\begin{figure}[H]
-{{\includegraphics[scale=0.53]{EofMCC-GS-T2_4-L20-1e5.pdf}}
-\subfloat{(a) }
-}\qquad
-{{\includegraphics[scale=0.53]{MofMCC-GS-T2_4-L20-1e5.pdf}}
-\subfloat{(b) }
-}\qquad
-\caption{text}
-\label{fig:GS}
+\begin{figure}[t]
+\subfloat[Energy pr. spin.]{%
+  \includegraphics[width=8.65cm]{EofMCC-GS-T2_4-L20-1e5.pdf}
+  \label{fig:L20-GS-E-T2.4}}\\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{
+  \includegraphics[width=8.65cm]{MofMCC-GS-T2_4-L20-1e5.pdf}
+  \label{fig:L20-GS-M-T2.4}}
+\caption{Time evolution of energy and magnetic moment of an ordered initial state with temperature 2.4 kT/J. $10^5$ Monte Carlo cycles.}
 \end{figure}
 
-\begin{figure}[H]
-{{\includegraphics[scale=0.53]{EofMCC-T1-L20-1e6.pdf}}
-\subfloat{(a) }
-}\qquad
-{{\includegraphics[scale=0.53]{MofMCC-T1-L20-1e6.pdf.pdf}}
-\subfloat{(b) }
-}\qquad
-\caption{text}
-\label{fig:GS}
+\begin{figure}[t]
+\subfloat[Energy pr. spin.]{%
+  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6.pdf}%
+  \label{fig:L20-E-T1}%
+}\\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{%
+  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6.pdf}%
+  \label{fig:L20-M-T1}%
+}
+\caption{Time evolution of energy and magnetic moment of an random initial state with temperature 1 kT/J. $10^6$ Monte Carlo cycles.}
 \end{figure}
 
-\begin{figure}[H]
-{{\includegraphics[scale=0.53]{EofMCC-T2_4-L20-1e6.pdf}}
-\subfloat{(b) }
-}\qquad
-{{\includegraphics[scale=0.53]{MofMCC-T2_4-L20-1e6.pdf}}
-\subfloat{(c) }
-}\qquad
-\caption{text}
-\label{fig:GS}
+\begin{figure}[t]
+\subfloat[Energy pr. spin.]{
+  \includegraphics[width=8.65cm]{EofMCC-T2_4-L20-1e6.pdf}
+  \label{fig:evaluation:revenue}}\\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{
+  \includegraphics[width=8.65cm]{MofMCC-T2_4-L20-1e6.pdf}
+  \label{fig:evaluation:avgPrice}}
+\caption{Time evolution (MC cycles) of energy and magnetic moment of an random initial state with temperature 2.4 kT/J. $10^6$ Monte Carlo cycles.}
 \end{figure}
@@
 \mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
+\caption{Logarithm of the accepted states as a function of the temperature for different }
+\label{fig:logAccepts}
 \end{figure}

### 2018-11-18 Project4/main.cpp
Adding results

@@ -106,19 +106,17 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
         int cutoff = mcc*0.05/numprocs; // 5% cutoff on each thread
 
         for(int m = myloop_begin; m <= myloop_end; m++) {
-                for (int x = 0; x < L; x++) {
-                        for (int y = 0; y < L; y++) {
-                                int xr = RNGPos(generator); // indices for random element
-                                int yr = RNGPos(generator);
+                for (int x = 0; x < pow(L,2); x++) {
+                        int xr = RNGPos(generator);         // indices for random element
+                        int yr = RNGPos(generator);
 
-                                int dE = 2.0*S(xr,yr)*(S(xr,PB(yr,L,1)) + S(xr,PB(yr,L,-1))
-                                                       + S(PB(xr,L,1),yr) + S(PB(xr,L,-1),yr));
-                                if (RNG(generator) <= acceptE.find(dE)->second) {
-                                        accepts += 1;
-                                        S(xr,yr) *= -1.0; // flipping spin
-                                        magMoment += (double) 2*S(xr,yr);
-                                        energy += (double) dE;
-                                }
+                        int dE = 2.0*S(xr,yr)*(S(xr,PB(yr,L,1)) + S(xr,PB(yr,L,-1))
+                                               + S(PB(xr,L,1),yr) + S(PB(xr,L,-1),yr));
+                        if (RNG(generator) <= acceptE.find(dE)->second) {
+                                accepts += 1;
+                                S(xr,yr) *= -1.0;         // flipping spin
+                                magMoment += (double) 2*S(xr,yr);
+                                energy += (double) dE;
                         }
                 }
                 if(m >= myloop_begin+cutoff) AddExpectValsToVec(ExpectVals, energy, magMoment);

### 2018-11-18 Project4/plot.py

@@ -3,10 +3,10 @@ import numpy as np
 import os
 
 accepts = False
-acceptOfT = False
+acceptOfT = True
 EM = False
 PE = False
-vary = True
+vary = False
 variance = False
 
 if EM == True:
@@ -62,14 +62,14 @@ if accepts == True:
     plt.show()
 
 if acceptOfT == True:
-    acceptT1 = np.loadtxt("data/T1-L20-mcc1e3.txt", usecols=5)
-    acceptT2_4 = np.loadtxt("data/T2_4-L20-mcc1e3.txt", usecols=5)
-    acceptT12 = np.loadtxt("data/T1-L20-mcc1e4.txt", usecols=5)
-    acceptT2_42 = np.loadtxt("data/T2_4-L20-mcc1e4.txt", usecols=5)
-    acceptT13 = np.loadtxt("data/T1-L20-mcc1e5.txt", usecols=5)
-    acceptT2_43 = np.loadtxt("data/T2_4-L20-mcc1e5.txt", usecols=5)
-    acceptT14 = np.loadtxt("data/T1-L20-mcc1e6.txt", usecols=5)
-    acceptT2_44 = np.loadtxt("data/T2_4-L20-mcc1e6.txt", usecols=5)
+    acceptT1 = np.loadtxt("data/old/L20-T1-mcc1e3-accepts.txt", usecols=5)
+    acceptT2_4 = np.loadtxt("data/old/L20-T2_4-mcc1e3-accepts.txt", usecols=5)
+    acceptT12 = np.loadtxt("data/old/L20-T1-mcc1e4-accepts.txt", usecols=5)
+    acceptT2_42 = np.loadtxt("data/old/L20-T2_4-mcc1e4-accepts.txt", usecols=5)
+    acceptT13 = np.loadtxt("data/old/L20-T1-mcc1e5-accepts.txt", usecols=5)
+    acceptT2_43 = np.loadtxt("data/old/L20-T2_4-mcc1e5-accepts.txt", usecols=5)
+    acceptT14 = np.loadtxt("data/old/L20-T1-mcc1e6-accepts.txt", usecols=5)
+    acceptT2_44 = np.loadtxt("data/old/L20-T2_4-mcc1e6-accepts.txt", usecols=5)
 
     plt.loglog(1,acceptT1[-1], "ro")
     plt.loglog(1,acceptT12[-1], "bo")
@@ -79,7 +79,7 @@ if acceptOfT == True:
[…]

### 2018-11-18 Project4/plot.py
Results

@@ -2,8 +2,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 import os
 
-accepts = False
-acceptOfT = True
+accepts = True
+acceptOfT = False
 EM = False
 PE = False
 vary = False
@@ -50,13 +50,15 @@ if PE == True:
 
 
 if accepts == True:
+    file = "data/L20-T1-1e5.txt"
+    file2 = "data/L20-T2_4-1e5.txt"
     accepts = np.loadtxt(file, usecols=5)
     accepts2 = np.loadtxt(file2, usecols=5)
 
     mcc = np.linspace(0,len(accepts),len(accepts))
     plt.loglog(mcc,accepts)
     plt.loglog(mcc,accepts2)
-    plt.legend(["T=1", "T=2"], prop={'size': 15})
+    plt.legend(["T=1", "T=2.4"], prop={'size': 15})
     plt.title("Accepts in Monte Carlo cycle for 20x20 lattice", size=15)
     plt.xlabel("log(Monte Carlo cycles)", size=15); plt.ylabel("log(Accepts)", size=15)
     plt.show()

### 2018-11-18 Project4/report/ms.tex

@@
 \mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
-\caption{Logarithm of the accepted states as a function of the temperature for different }
-\label{fig:logAccepts}
+\caption{Logarithm of the accepted states as a function of the temperature for different number of Monte Carlo cycles.
+The left four dots corresponds to $T=1$ kT/J, and the right four dots corresponds to $T=2.4$ kT/J.
+The initial spins are randomly organized.}
+\label{fig:logAccepts(t)}
 \end{figure}
@@
 \mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
+\caption{Logarithmic plot of accepted states as a function of }
+\label{fig:logAccepts}
 \end{figure}

### 2018-11-19 Project4/analytical.py
Adding results

@@ -4,8 +4,7 @@ n = 2
 T = 1.0
 
 E = -8.0*np.sinh(8/T)/(3.0 + np.cosh(8/T))
-#CV = 1/T**2*(64.0*np.cosh(8/T)*(3 + np.cosh(8/T)) - 64*np.sinh(8/T)**2)/(3+np.cosh(8/T))**2
-CV = (64*np.cosh(8.0/T))/(T**2*(3 + np.cosh(8.0/T)))
+CV = 64/(T*T) * (np.cosh(8)*(3 + np.cosh(8)) - (np.sinh(8)**2))/(3 + np.cosh(8))**2
 Mabs = (2*np.exp(8/T) + 4.0)/(3.0 + np.cosh(8/T))
 M2 = 8.0*(np.exp(8/T) + 1.0)/(3.0 + np.cosh(8/T))
 chi = M2/T

### 2018-11-19 Project4/main.cpp

@@ -51,7 +51,7 @@ int main(int argc, char *argv[]){
         double TimeStart, TimeEnd;
         TimeStart = MPI_Wtime();
 
-        for(double T = 2.0; T <= 2.3; T+=0.03) {
+        for(double T = 1.0; T <= 1.0; T+=0.01) {
                 Metropolis(L, mcc, T, energies(T), numprocs, my_rank, myloop_end, myloop_begin, filename);
         }
 
@@ -103,7 +103,7 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
         if(my_rank == 0) outfile.open(filename, fstream::app);
 
         int accepts = 0;
-        int cutoff = mcc*0.05/numprocs; // 5% cutoff on each thread
+        int cutoff = 0;//mcc*0.05/numprocs; // 5% cutoff on each thread
 
         for(int m = myloop_begin; m <= myloop_end; m++) {
                 for (int x = 0; x < pow(L,2); x++) {

### 2018-11-19 Project4/main2.cpp

@@ -1,214 +1,0 @@
-#include "mpi.h"
-#include <cmath>
-#include <iostream>
-#include <fstream>
-#include <random>
-#include <armadillo>
-#include <random>
-#include <map>
-
-using namespace std;
-using namespace arma;
-
-ofstream outfile;
-
-inline int PB(int index, int spins, int correction) {
-        // returns the index of the element
-        return (index + spins + correction)%(spins);
-}
-
-void Energy(int L, mat &S, double &energy);
-void initialize(double &energy, double &magMoment, int L, mat S, int GS);
-void Metropolis(int L, int mcs, double T, map<double, double> acceptE, int numprocs, int my_rank, int myloop_end, int myloop_begin, string filename);
-map<double, double> energies(double T);
-void WriteToFile(double energy, double magMoment, int mcc, double T, int accepts, int L);
-void AddExpectValsToVec(vec &ExpectVals, double energy, double magMoment);
-void PrintExpectVals(vec TotalExpectVals, int mcc, double T);
-void WriteExpectValsToFile(vec TotalExpectVals, int mcc, double T, int L);
-
-
-int main(int argc, char *argv[]){
-        int L, mcc, numprocs, my_rank;
-        //double T;
-        string filename;
-
-        //T = atoi(argv[4]);    // temperature [kT/J]
-        L = atoi(argv[1]);    // number of spins
-        mcc = atoi(argv[2]);  // number of MC cycles
-        filename = argv[3];   // name of file
-
-        //  MPI initializations
-        MPI_Init (&argc, &argv);
-        MPI_Comm_size (MPI_COMM_WORLD, &numprocs);
-        MPI_Comm_rank (MPI_COMM_WORLD, &my_rank);
-
-        int no_intervals = mcc/numprocs;
-        int myloop_begin = my_rank*no_intervals + 1;
-        int myloop_end = (my_rank + 1)*no_intervals;
-
-        if((my_rank == numprocs - 1) && (myloop_end < mcc)) myloop_end = mcc;
-
-        double TimeStart, TimeEnd;
-        TimeStart = MPI_Wtime();
-
-        for(double T = 2.2; T <= 2.4; T+=0.01) {
-                Metropolis(L, mcc, T, energies(T), numprocs, my_rank, myloop_end, myloop_begin, filename);
-        }
-
-        TimeEnd = MPI_Wtime();
-        if(my_rank == 0) {
-                cout << "Time = " << TimeEnd - TimeStart << " on number of threads: " << numprocs << endl;
[…]

### 2018-11-19 Project4/main3.cpp

@@ -1,214 +1,0 @@
-#include "mpi.h"
-#include <cmath>
-#include <iostream>
-#include <fstream>
-#include <random>
-#include <armadillo>
-#include <random>
-#include <map>
-
-using namespace std;
-using namespace arma;
-
-ofstream outfile;
-
-inline int PB(int index, int spins, int correction) {
-        // returns the index of the element
-        return (index + spins + correction)%(spins);
-}
-
-void Energy(int L, mat &S, double &energy);
-void initialize(double &energy, double &magMoment, int L, mat S, int GS);
-void Metropolis(int L, int mcs, double T, map<double, double> acceptE, int numprocs, int my_rank, int myloop_end, int myloop_begin, string filename);
-map<double, double> energies(double T);
-void WriteToFile(double energy, double magMoment, int mcc, double T, int accepts, int L);
-void AddExpectValsToVec(vec &ExpectVals, double energy, double magMoment);
-void PrintExpectVals(vec TotalExpectVals, int mcc, double T);
-void WriteExpectValsToFile(vec TotalExpectVals, int mcc, double T, int L);
-
-
-int main(int argc, char *argv[]){
-        int L, mcc, numprocs, my_rank;
-        //double T;
-        string filename;
-
-        //T = atoi(argv[4]);    // temperature [kT/J]
-        L = atoi(argv[1]);    // number of spins
-        mcc = atoi(argv[2]);  // number of MC cycles
-        filename = argv[3];   // name of file
-
-        //  MPI initializations
-        MPI_Init (&argc, &argv);
-        MPI_Comm_size (MPI_COMM_WORLD, &numprocs);
-        MPI_Comm_rank (MPI_COMM_WORLD, &my_rank);
-
-        int no_intervals = mcc/numprocs;
-        int myloop_begin = my_rank*no_intervals + 1;
-        int myloop_end = (my_rank + 1)*no_intervals;
-
-        if((my_rank == numprocs - 1) && (myloop_end < mcc)) myloop_end = mcc;
-
-        double TimeStart, TimeEnd;
-        TimeStart = MPI_Wtime();
-
-        for(double T = 2.0; T <= 2.3; T+=0.015) {
-                Metropolis(L, mcc, T, energies(T), numprocs, my_rank, myloop_end, myloop_begin, filename);
-        }
-
-        TimeEnd = MPI_Wtime();
-        if(my_rank == 0) {
-                cout << "Time = " << TimeEnd - TimeStart << " on number of threads: " << numprocs << endl;
[…]

### 2018-11-19 Project4/plot.py

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import os
 
-accepts = True
+accepts = False
 acceptOfT = False
 EM = False
 PE = False

### 2018-11-19 Project4/report/ms.tex

@@
 \begin{gather*}
-    C_V = \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}.
+    C_V = \frac{64J^2}{k_B T^2} \left(\frac{\cosh(8\beta J)(3+\cosh(8 \beta J)) - \sinh^2(8\beta J)}{\left(3 + \cosh(8\beta J)\right)^2}\right)
 \end{gather*}
@@
 \begin{align*}
-    \langle E \rangle &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right), \\
-    \langle |M| \rangle &= \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)}, \\
-    C_V &= \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}, \\
-    \chi &= \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
+    &\langle E \rangle = -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right), \\
+    &\langle |M| \rangle = \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)}, \\
+    &C_V = \frac{64J^2}{k T^2} \left(\frac{\cosh(8\beta J)(3+\cosh(8 \beta J)) - \sinh^2(8\beta J)}{\left(3 + \cosh(8\beta J)\right)^2}\right), \\
+    &\chi = \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
 \end{align*}
@@
 
-\begin{figure}[t]
+\begin{deluxetable}{lccccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:analytical}}
+\tablecomments{Comparison of analytical and numerical results for various values.}
+\tablecolumns{4}
+\tablehead{ $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
+\startdata
+-7.983 & 3.994 & 0.128 & 15.973 
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lccccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:results}}
+\tablecomments{Comparison of analytical and numerical results for various MC cycles. Lattice size is 2x2, temperature 1 kT/J and initial spin randomly set.}
+\tablecolumns{5}
+\tablehead{MC cycles & $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
+\startdata
+1e3 & -7.976 & 3.992 & 0.191 & 0.023 \\
+5e3 & -7.984 & 3.994 & 0.127 & 0.027 \\
+1e4 & -7.968 & 3.989 & 0.254 & 15.928 \\
+1e5 & -7.984 & 3.994 & 0.122 & 15.403 \\
+1e6 & -7.984 & 3.994 & 0.127 & 15.973 \\
+Analytical & -7.983 & 3.994 & 0.128 & 15.973 
+\enddata
+\end{deluxetable}
+
+\begin{figure}[h]
 \subfloat[Energy pr. spin.]{
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 \subfloat[Energy pr. spin.]{%
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 \subfloat[Energy pr. spin.]{
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 \mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 \mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
-\caption{Logarithmic plot of accepted states as a function of }
+\caption{Logarithmic plot of accepted states as a function of Monte Carlo cycles. The initial spins are randomly organized.}
 \label{fig:logAccepts}
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
+\mbox{\epsfig{figure=E-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\label{fig:E-T2-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=M-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Mean absolute value of magnetic moment pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\label{fig:M-T2-23}
+\end{figure}
+
+\begin{figure}[h]
 \mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
+\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\label{fig:CV-T2-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=X-T2-23.pdf,width=\linewidth,clip=}}
+\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\label{fig:chi-T2-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=E-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\label{fig:E-T22-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=M-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Mean abslute value of the magnetic moment pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\label{fig:M-T22-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=CV-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\label{fig:CV-T22-24}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=X-T22-24.pdf,width=\linewidth,clip=}}
+\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\label{fig:chi-T22-23}
+\end{figure}
+
+\begin{figure}[h]
+\mbox{\epsfig{figure=FittedTc.pdf,width=\linewidth,clip=}}
+\caption{Linear regression used on the critical temperatures at points of 1/L for $L = 40$, $60$, $80$ and $100$ gives the function $y(x) = 3.45x + 2.22$.}
 \label{fig:figure_label}
@@
 
-\begin{figure}[t]
-\mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
+\begin{figure}[h]
+\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Histogram of the probability as a function of energy. Temperature is 1 kT/J. Initial spin states are random.}
 \label{fig:figure_label}
@@
 
-\begin{figure}[t]
-\mbox{\epsfig{figure=CV-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
+\begin{figure}[h]
+\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Histogram of the probability as a function of energy. Temperature is 2.4 kT/J. Initial spin states are random.}
 \label{fig:figure_label}
@@
 
-\begin{figure}[t]
-\mbox{\epsfig{figure=E-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
+\begin{deluxetable}{cccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:runtimes40}}
+\tablecomments{Run times for a 40x40 lattice for $10^5$ Monte Carlo cycles. Temperature set to 1 kT/J and the initial state for the spins are random.}
+\tablecolumns{4}
+\tablehead{Run & 1 core & 4 core & Difference}
+\startdata
+1 & 13.910 & 4.448 & 9.461 \\
+2 & 14.348 & 4.530 & 9.817 \\
+3 & 14.407 & 4.636 & 9.770 \\
+Average & 14.221 & 4.538 & 9.683
+\enddata
+\end{deluxetable}
 
-\begin{figure}[t]
-\mbox{\epsfig{figure=E-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=FittedTc.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=M-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=M-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=X-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[t]
-\mbox{\epsfig{figure=X-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Description of figure -- explain all elements, but do not
-draw conclusions here.}
-\label{fig:figure_label}
-\end{figure}
+\begin{deluxetable}{cccc}
+\tablewidth{0pt}
+\tablecaption{\label{tab:runtimes100}}
+\tablecomments{Run times for a 100x100 lattice for $10^5$ Monte Carlo cycles. Temperature set to 1 kT/J and the initial state for the spins are random.}
+\tablecolumns{5}
+\tablehead{Run & 1 core & 4 core & Difference}
+\startdata
+1 & 91.137 & 34.434 & 62.702 \\
+2 & 83.662 & 30.455 & 53.207 \\
+3 & 86.905 & 28.534 & 58.371 \\
+Average & 87.235 & 31.141 & 58.093
+\enddata
+\end{deluxetable}
 
@@
 
-%\begin{figure}[t]
+%\begin{figure}[h]
 %

### 2018-11-19 Project4/Tc.py
Results

@@ -17,7 +17,7 @@ chi4 = np.loadtxt(file4, usecols=3)
 files = [file,file2,file3,file4]
 
 L = np.array([1.0/100, 1.0/80, 1.0/60, 1.0/40])
-T = np.array([T[np.argmax(chi4)] - L[3], T[np.argmax(chi3)]- L[2], T[np.argmax(chi2)]- L[1], T[np.argmax(chi)]- L[0]])
+T = np.array([T[np.argmax(chi4)] - L[0], T[np.argmax(chi3)]- L[1], T[np.argmax(chi2)]- L[2], T[np.argmax(chi)]- L[3]])
 
 z = np.polyfit(L,T,1)
 zz = np.poly1d(z)

### 2018-11-19 Project4/plot.py

@@ -8,6 +8,7 @@ EM = False
 PE = False
 vary = False
 variance = False
+sigma = True
 
 if EM == True:
     file = "data/L20-GS-T1-1e6.txt"
@@ -36,7 +37,8 @@ if EM == True:
     plt.show()
 
 if PE == True:
-    E_sorted = np.sort(np.loadtxt(dfile, usecols=0))
+    file = "data/L20-T1-1e6.txt"
+    E_sorted = np.sort(np.loadtxt(file, usecols=0))
     mcc = np.linspace(0,len(E_sorted),len(E_sorted))
     bins = 0
     for i in range(0,len(E_sorted)-1):
@@ -46,6 +48,15 @@ if PE == True:
     plt.legend(["T=1"], prop={'size':15})
     plt.xlabel("Energy", size=15); plt.ylabel("Probability",size=15)
     plt.title("Probability distribution of energy for 20x20 lattice", size=15)
+    plt.show()
+if sigma == True:
+    file = "data/L20-T2_4-1e6.txt"
+
+    sigma2 = np.loadtxt(file, usecols=5)
+    plt.plot(sigma2)
+    plt.legend(["T=1"], prop={'size':15})
+    plt.xlabel("MC cycles", size=15); plt.ylabel("$\\sigma_E^2$",size=15)
+    plt.title("Variance of the energy for 20x20 lattice", size=15)
     plt.show()
 
 

### 2018-11-19 Project4/main.cpp
Results

@@ -51,7 +51,7 @@ int main(int argc, char *argv[]){
         double TimeStart, TimeEnd;
         TimeStart = MPI_Wtime();
 
-        for(double T = 1.0; T <= 1.0; T+=0.01) {
+        for(double T = 2.4; T <= 2.4; T+=0.01) {
                 Metropolis(L, mcc, T, energies(T), numprocs, my_rank, myloop_end, myloop_begin, filename);
         }
 
@@ -103,7 +103,7 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
         if(my_rank == 0) outfile.open(filename, fstream::app);
 
         int accepts = 0;
-        int cutoff = 0;//mcc*0.05/numprocs; // 5% cutoff on each thread
+        int cutoff = mcc*0.05/numprocs; // 5% cutoff on each thread
 
         for(int m = myloop_begin; m <= myloop_end; m++) {
                 for (int x = 0; x < pow(L,2); x++) {
@@ -132,7 +132,7 @@ void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numpr
 
         if(my_rank == 0) {
                 PrintExpectVals(TotalExpectVals, mcc - cutoff*numprocs, T);
-                WriteExpectValsToFile(TotalExpectVals, mcc - cutoff*numprocs, T, L);
+                //WriteExpectValsToFile(TotalExpectVals, mcc - cutoff*numprocs, T, L);
         }
         if(my_rank == 0) outfile.close();
 }
@@ -166,7 +166,7 @@ void WriteToFile(double energy, double magMoment, int mcc, double T, int accepts
         outfile << chi/pow(L,2) << " ";
         outfile << Mabs/pow(L,2) << " ";
         outfile << accepts << " ";
-        outfile << pow(E2 - pow(E,2),2) << endl;
+        outfile << E2/(double) mcc - pow(E/(double) mcc,2) << endl;
 }
 
 void AddExpectValsToVec(vec &ExpectVals, double energy, double magMoment){
@@ -182,7 +182,8 @@ void PrintExpectVals(vec TotalExpectVals, int mcc, double T) {
         cout << "E: " << TotalExpectVals(0) << " ";
         cout << "Mabs: " << TotalExpectVals(4) << " ";
         cout << "C_V: " << (TotalExpectVals(1) - pow(TotalExpectVals(0),2))/(pow(T,2)) << " ";
[…]

### 2018-11-19 Project4/plot.py

@@ -5,10 +5,10 @@ import os
 accepts = False
 acceptOfT = False
 EM = False
-PE = False
+PE = True
 vary = False
 variance = False
-sigma = True
+sigma = False
 
 if EM == True:
     file = "data/L20-GS-T1-1e6.txt"
@@ -37,22 +37,22 @@ if EM == True:
     plt.show()
 
 if PE == True:
-    file = "data/L20-T1-1e6.txt"
+    file = "data/test/L20-T2_4-1e6.txt"
     E_sorted = np.sort(np.loadtxt(file, usecols=0))
     mcc = np.linspace(0,len(E_sorted),len(E_sorted))
     bins = 0
     for i in range(0,len(E_sorted)-1):
             if E_sorted[i+1] != E_sorted[i]:
                 bins += 1
-    plt.hist(E_sorted,bins+10)
+    plt.hist(E_sorted,bins+7)
     plt.legend(["T=1"], prop={'size':15})
-    plt.xlabel("Energy", size=15); plt.ylabel("Probability",size=15)
-    plt.title("Probability distribution of energy for 20x20 lattice", size=15)
+    plt.xlabel("Energy", size=15); plt.ylabel("Occurences",size=15)
+    plt.title("Histogram of energy for 20x20 lattice", size=15)
     plt.show()
 if sigma == True:
-    file = "data/L20-T2_4-1e6.txt"
+    file = "data/test/L20-T2_4-1e5.txt"
 
-    sigma2 = np.loadtxt(file, usecols=5)
+    sigma2 = np.loadtxt(file, usecols=6)
     plt.plot(sigma2)
     plt.legend(["T=1"], prop={'size':15})
     plt.xlabel("MC cycles", size=15); plt.ylabel("$\\sigma_E^2$",size=15)

### 2018-11-20 Project4/report/Project4.tex
Finished

--- 2018-11-19 Project4/report/ms.tex
@@
 \usepackage{url}
-\usepackage{subfig}
+\usepackage[caption=false]{subfig}
+\usepackage{url}
 \setcitestyle{square}
@@
 \begin{abstract}
+In this study of phase transformations of a magnetic system we will estimate the critical temperature using the Ising model.
+To study the system of time we will use the Metropolis algorithm, which is a Markov chain-Monte Carlo method.
+Since we do not know all the possible states of the system beforehand, the Metropolis algorithm is beneficial to use, since it eliminates the partition function.
+The study will show that for a cold system of temperature 1 kT/J the system rarely has states of higher energies, when increasing the temperature to 2.4 kT/J the total energy of the system is increased, and we see that the probability for the system to be in a higher energy state is increased.
+When we look at how the system evolves as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature.
+When estimating the critical temperature we find this to be $T_C = 2.256$ kT/J, which is fairly close to the analytical value found by Lars Onsager \cite{bib:onsager}.
 
 \end{abstract}
-\keywords{computational science: Ising model  --- methods: Metropolis, Monte Carlo}
+\keywords{computational science: Ising model  --- methods: Metropolis algorithm, Markov chain-Monte Carlo}
 
@@
 %
-\begin{align*}
-    P(E_i) = \frac{e^{-\beta E_i}}{Z},
-\end{align*}
+\begin{align}\label{eq:P_i}
+    P_i = P(E_i) = \frac{e^{-\beta E_i}}{Z},
+\end{align}
 %
@@
 \begin{align*}
-    C_V = \frac{1}{kT^2}&\left(\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right \\
-    &-\left \frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right),
+    C_V = \frac{1}{kT^2}\bigg(&\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2} \\ &-\frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\bigg),
 \end{align*}
@@
 
-\subsection{Algorithm}
+\subsection{Implementation}
 For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment.
@@
 
-PBC
+\subsection{Critical temperature}
+As the temperature of the system moves towards the critical temperature, the susceptibility $\chi$ and the heat capacity $C_V$ moves towards an infinite value. We can see this by
+%
+\begin{gather*}
+    C_V \sim |T_C - T|^{-\alpha} \\
+    \chi \sim |T_C - T|^{-\nu},
+\end{gather*}
+%
+where $\alpha$ and $\nu$ are positive constants.
+For the results in this study we have clearer peaks for the susceptibility, so we will extract the critical temperature from these.
+We know that the critical tempeature scales as
+%
+\begin{gather*}
+    T_C\left(L\right) - T_C\left(L=\infty\right) = aL^{-1/\nu},
+\end{gather*}
+where we set $\nu = 1$, and a is a positive constant. Solving the above equation for the critical temperature as the lattice size is infinite we get
+%
+\begin{gather*}
+    T_C\left(L=\infty \right) \approx T_C\left(L\right) - \frac{1}{L}.
+\end{gather*}
 
@@
 \label{sec:results}
+Following are the results produced in this study. The results are found in plots and tables, and the content is presented in the following subsections.
 
-\begin{deluxetable}{lccccc}
+\subsection{Comparison of analytical and numerical values}
+Table \ref{tab:numAnal} shows a comparison of analytical and numerical values for a 2x2 lattice.
+
+\begin{deluxetable}{lcccc}
 \tablewidth{0pt}
-\tablecaption{\label{tab:analytical}}
-\tablecomments{Comparison of analytical and numerical results for various values.}
-\tablecolumns{4}
-\tablehead{ $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
-\startdata
--7.983 & 3.994 & 0.128 & 15.973 
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lccccc}
-\tablewidth{0pt}
-\tablecaption{\label{tab:results}}
+\tablecaption{\label{tab:numAnal}}
 \tablecomments{Comparison of analytical and numerical results for various MC cycles. Lattice size is 2x2, temperature 1 kT/J and initial spin randomly set.}
@@
 \startdata
-1e3 & -7.976 & 3.992 & 0.191 & 0.023 \\
-5e3 & -7.984 & 3.994 & 0.127 & 0.027 \\
-1e4 & -7.968 & 3.989 & 0.254 & 15.928 \\
-1e5 & -7.984 & 3.994 & 0.122 & 15.403 \\
+1e3 & -7.984 & 3.996 & 0.127 & 0.019 \\
+5e3 & -7.985 & 3.994 & 0.114 & 15.916 \\
+1e4 & -7.981 & 3.994 & 0.146 & 15.769 \\
+1e5 & -7.984 & 3.994 & 0.129 & 15.847 \\
 1e6 & -7.984 & 3.994 & 0.127 & 15.973 \\
@@
 
-\begin{figure}[h]
+\subsection{Mean energy and magnetisation 20x20 lattice}
+Figures \ref{fig:L20-GS-E-T1} and \ref{fig:L20-GS-M-T1} shows the energy and magnetisation of the ground state for temperature 1 kT/J, while figures \ref{fig:L20-GS-E-T2.4} and \ref{fig:L20-GS-M-T2.4} shows the same for temperature 2.4 kT/J.
+We see that the system stabilizes quickly.
+For the non-ground state plots see figures \ref{fig:L20-E-T1} and \ref{fig:L20-M-T1} for temperature 1 kT/J and figures \ref{fig:L20-E-T2.4} and \ref{fig:L20-M-T2.4} for temperature 2.4 kT/J.
+For the system that starts in a random spin state we see by figure \ref{fig:EM-T1_zoom} that the system stabilizes at about 250 MC cycles.
+%
+\begin{figure}
 \subfloat[Energy pr. spin.]{
-  \includegraphics[width=9cm]{EofMCC-GS-T1-L20-1e5.pdf}
-  \label{fig:L20-GS-E-T1}}\\
+  \includegraphics[width=8.65cm]{EofMCC-GS-T1-L20-1e5.pdf}
+  \label{fig:L20-GS-E-T1}} \\
 \subfloat[Absolute value of the magnetic moment pr. spin.]{
-  \includegraphics[width=9cm]{MofMCC-GS-T1-L20-1e5.pdf}
-  \label{fig:evaluation:L20-GS-M-T1}}
+  \includegraphics[width=8.65cm]{MofMCC-GS-T1-L20-1e5.pdf}
+  \label{fig:L20-GS-M-T1}}
 \caption{Time evolution (MC cycles) of energy and magnetic moment of an ordered initial state with temperature 1 kT/J. $10^5$ Monte Carlo cycles.}
+\label{fig:EM-GS-T1}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \subfloat[Energy pr. spin.]{%
   \includegraphics[width=8.65cm]{EofMCC-GS-T2_4-L20-1e5.pdf}
-  \label{fig:L20-GS-E-T2.4}}\\
+  \label{fig:L20-GS-E-T2.4}} \\
 \subfloat[Absolute value of the magnetic moment pr. spin.]{
@@
 \caption{Time evolution of energy and magnetic moment of an ordered initial state with temperature 2.4 kT/J. $10^5$ Monte Carlo cycles.}
+\label{fig:EM-GS-T2.4}
+\end{figure}
+%
+\begin{figure}
+\subfloat[Energy pr. spin.]{
+  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6.pdf}
+  \label{fig:L20-E-T1}} \\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{%
+  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6.pdf}
+  \label{fig:L20-M-T1}}
+\caption{Time evolution of energy and magnetic moment of an random initial state with temperature 1 kT/J. $10^6$ Monte Carlo cycles.}
+\label{fig:EM-T1}
+\end{figure}
+%
+\begin{figure}
+\subfloat[Zoom of figure \ref{fig:L20-E-T1}.]{
+  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6_zoom.pdf}
+  \label{fig:L20-E-T1_zoom}} \\
+\subfloat[Zoom of figure \ref{fig:L20-M-T1}.]{%
+  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6_zoom.pdf}
+  \label{fig:L20-M-T1_zoom}}
+\caption{Zoom of previous plot to show where the energy and magnetisation stabilized.}
+\label{fig:EM-T1_zoom}
+\end{figure}
+%
+\begin{figure}
+\subfloat[Energy pr. spin.]{
+  \includegraphics[width=8.65cm]{EofMCC-T2_4-L20-1e6.pdf}
+  \label{fig:L20-E-T2.4}} \\
+\subfloat[Absolute value of the magnetic moment pr. spin.]{ 
+  \includegraphics[width=8.65cm]{MofMCC-T2_4-L20-1e6.pdf}
+  \label{fig:L20-M-T2.4}}
+\caption{Time evolution (MC cycles) of energy and magnetic moment of an random initial state with temperature 2.4 kT/J. $10^6$ Monte Carlo cycles.}
+\label{fig:EM-T2.4}
 \end{figure}
 
-\begin{figure}[h]
-\subfloat[Energy pr. spin.]{%
-  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6.pdf}%
-  \label{fig:L20-E-T1}%
-}\\
-\subfloat[Absolute value of the magnetic moment pr. spin.]{%
-  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6.pdf}%
-  \label{fig:L20-M-T1}%
-}
-\caption{Time evolution of energy and magnetic moment of an random initial state with temperature 1 kT/J. $10^6$ Monte Carlo cycles.}
-\end{figure}
-
-\begin{figure}[h]
-\subfloat[Energy pr. spin.]{
-  \includegraphics[width=8.65cm]{EofMCC-T2_4-L20-1e6.pdf}
-  \label{fig:evaluation:revenue}}\\
-\subfloat[Absolute value of the magnetic moment pr. spin.]{
-  \includegraphics[width=8.65cm]{MofMCC-T2_4-L20-1e6.pdf}
-  \label{fig:evaluation:avgPrice}}
-\caption{Time evolution (MC cycles) of energy and magnetic moment of an random initial state with temperature 2.4 kT/J. $10^6$ Monte Carlo cycles.}
-\end{figure}
-
-\begin{figure}[h]
+\subsection{Accepted states in the Metropolis algorithm}
+The accepted states as a function of temperature and MC cycles are found in figures \ref{fig:logAccepts(t)} and \ref{fig:logAccepts} respectively, both are logarithmic plots.
+\begin{figure}
 \mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
@@
 \end{figure}
-
-\begin{figure}[h]
+\begin{figure}
 \mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
@@
 
-\begin{figure}[h]
+\subsection{Probability of energy}
+In figures \ref{fig:histT1} and \ref{fig:histT2.4} we can see the probability distribution for a given energy for temperature 1 kT/J and 2.4 kT/J respectively.
+The variance of the energy $\sigma_E^2 = 9.398$ for temperature 1 kT/J and $\sigma_E^2 = 3264.7$ for temperature 2.4 kT/J.
+%
+\begin{figure}
+\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Histogram of the probability as a function of energy. Temperature is 1 kT/J. Initial spin states are random. 1e6 MC cycles.}
+\label{fig:histT1}
+\end{figure}
+%
+\begin{figure}
+\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
+\caption{Histogram of the probability as a function of energy. Temperature is 2.4 kT/J. Initial spin states are random. 1e6 MC cycles.}
+\label{fig:histT2.4}
+\end{figure}
+
+\subsection{Varying temperature}
+\begin{figure}
 \mbox{\epsfig{figure=E-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\caption{Mean energy pr. spin for various lattice sizes. $10^6$ MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
 \label{fig:E-T2-23}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=M-T2-23.pdf,width=\linewidth,clip=}}
@@
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\caption{Heat capacity pr. spin for various lattice sizes. $10^6$ MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
 \label{fig:CV-T2-23}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=X-T2-23.pdf,width=\linewidth,clip=}}
-\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
+\caption{Susceptibility pr. spin for various lattice sizes. $10^6$ MC cycles.. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
 \label{fig:chi-T2-23}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=E-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\caption{Mean energy pr. spin for various lattice sizes. $10^6$ MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
 \label{fig:E-T22-23}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=M-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Mean abslute value of the magnetic moment pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\caption{Mean abslute value of the magnetic moment pr.
+spin for various lattice sizes.
+$10^6$ MC cycles..
+Initial spin state is random.
+Temperature varies with a step size of 0.01, making it twenty steps.}
 \label{fig:M-T22-23}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=CV-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\caption{Heat capacity pr. spin for various lattice sizes. $10^6$ MC cycles.. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
 \label{fig:CV-T22-24}
 \end{figure}
-
-\begin{figure}[h]
+%
+\begin{figure}
 \mbox{\epsfig{figure=X-T22-24.pdf,width=\linewidth,clip=}}
-\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
+\caption{Susceptibility pr. spin for various lattice sizes. $10^6$ MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
 \label{fig:chi-T22-23}
 \end{figure}
+%
+In figures \ref{fig:E-T2-23}-\ref{fig:chi-T2-23} we can see plots of mean energy, mean magnetisation, heat capacity and susceptibility as a function of temperature.
+The temperature varies from 2-2.3 kT/J.
+The number of MC cycles are $10^6$, with a random initial state for the spins.
+In figures \ref{fig:E-T22-23}-\ref{fig:chi-T22-23} the temperature interval and temperature step are changed to $2.2-2.4$ kT/J and $0.01$ respectively, this is done to focus on the temperature where we expect to find the critical temperature.
 
-\begin{figure}[h]
+\subsection{Critical temperature}
+\begin{figure}
 \mbox{\epsfig{figure=FittedTc.pdf,width=\linewidth,clip=}}
-\caption{Linear regression used on the critical temperatures at points of 1/L for $L = 40$, $60$, $80$ and $100$ gives the function $y(x) = 3.45x + 2.22$.}
-\label{fig:figure_label}
+\caption{Linear regression used on the critical temperatures at points of 1/L for $L = 40$, $60$, $80$ and $100$ gives the function $y(x) = 1.586x + 2.256$.}
+\label{fig:FittedTc}
 \end{figure}
+%
+Figure \ref{fig:FittedTc} the critical temperature as a function of 1/L is plottet.
+Critical temperature was extracted from the susceptibility plotted in figure \ref{fig:chi-T22-23}, that is; we have extracted the corresponding temperature of the highest value of the susceptibility.
+Linear regression gives the function $y(x) = 1.586x + 2.256$.
+When $L \rightarrow \infty$, this corresponds to $1/L \rightarrow 0$.
+We can see that $y(0) = 2.256$, which is the numerical value for the critical temperature when the lattice size $L \rightarrow \infty$.
 
-\begin{figure}[h]
-\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
-\caption{Histogram of the probability as a function of energy. Temperature is 1 kT/J. Initial spin states are random.}
-\label{fig:figure_label}
-\end{figure}
-
-\begin{figure}[h]
-\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
-\caption{Histogram of the probability as a function of energy. Temperature is 2.4 kT/J. Initial spin states are random.}
-\label{fig:figure_label}
-\end{figure}
-
+\subsection{Run times}
+In table \ref{tab:runtimes40} and \ref{tab:runtimes100} the runes times for a 40x40 and a 100x100 lattice, respectively. 
+%
 \begin{deluxetable}{cccc}
@@
 \tablecaption{\label{tab:runtimes40}}
-\tablecomments{Run times for a 40x40 lattice for $10^5$ Monte Carlo cycles. Temperature set to 1 kT/J and the initial state for the spins are random.}
+\tablecomments{Run times for a 40x40 lattice for $10^5$ Monte Carlo cycles.
+Temperature set to 1 kT/J and the initial state for the spins are random.
+The machine that has been used for the calculations has two cores, where each core has two threads, making up a total of four threads.}
 \tablecolumns{4}
-\tablehead{Run & 1 core & 4 core & Difference}
+\tablehead{Run & 1 thread & 4 threads & Difference}
 \startdata
@@
 \end{deluxetable}
-
+%
 \begin{deluxetable}{cccc}
@@
 \label{sec:discussion}
+ When comparing the analytical and the numerical values for a 2x2 lattice we can see in Table \ref{tab:numAnal} that the numerical values for the mean energy and mean absolute value of the magnetic moment are fairly stable, and does not deviate much from the analytical values, even for few MC cycles.
+At MC cycles less than 10000 we see that the results fluctuate, and after they have stabilized.
+
+ In the figures \ref{fig:EM-GS-T1} and \ref{fig:EM-GS-T2.4} showing the energy and magnetic moment for the ground state (i.e.
+ordered initial state) for temperature 1 and 2.4 kT/J respectively, shows that the system stabilizes quickly.
+We expect the system to stabilize quickly due to the fact that the initial state in the most likely state.
+For the low temperature situation there are few states where the energy of the system increases.
+
+ If we start the system in a random spin state the system takes some time to stabilize, as shown in figure \ref{fig:EM-T1_zoom}, a zoom of figure \ref{fig:EM-T1}, for the low temperature.
+For the increased temperature we see in figure \ref{fig:EM-T2.4} that the stable state of the system have much higher amplitudes of the energy and the magnetisation than that of the lower temperature, since there is more energy present in the system.
+ %
+ From the probability distribution, Eq. \eqref{eq:P_i}, we see that 
+ %
+ \begin{align*}
+ P_i \propto e^{-\beta E_i} = e^{-E_i/kT} \propto e^{-1/T}
+ \end{align*}
+ %
+ for a state with energy $E_i$.
+So the probability of a state with the energy $E_i$ increases with the temperature.
+In the case of the higher temperature, the probability of reaching a state with a higher energy is increased.
+In figure \ref{fig:EM-T2.4} we see that there are more states with higher energies, which is a direct result from increasing the temperature, and therefore the energy of the system.
+Looking at the accepted states as a function of MC cycles shown in figure \ref{fig:logAccepts} and \ref{fig:logAccepts(t)} we can see that there are more accepted states for the case where the system has the highest temperature.
+Further we see in the histograms in figure \ref{fig:histT1} and \ref{fig:histT2.4} that the system has more states with with varying energy when the temperature is higher.
+
+Studying the Ising model as the temperature of the system increases we see in figure \ref{fig:E-T2-23}-\ref{fig:chi-T2-23} that the system is stable, and there is not much happening, so the model of 40x40 lattice is sufficient.
+When the temperature reaches $\sim 2.20-2.23$ the temperature has reached the analytical value of the critical temperature, which suggests a phase transition.
+In figures \ref{fig:E-T22-23} and \ref{fig:chi-T22-23} the temperature interval is from $2.2-2.4$, and the resolution of the temperature is increased.
+The figures shows that there is a phase transition, which depends on the lattice size.
+In figure \ref{fig:chi-T22-23} we can see that the peak of the susceptibility and the heat capacity spikes as we increase the lattice size, which is expected for the Ising model.
+Further increasing the lattice size we would expect sharper and taller peak, and as $L\rightarrow \infty$ we expect $\chi$, $C_V \rightarrow \infty$.
+Extracting the values of the peak from the susceptibility we use linear regression to find the value of the critical temperature as $L\rightarrow \infty \Longleftrightarrow 1/L \rightarrow 0$, the function is shown in figure \ref{fig:FittedTc}.
+When increasing the lattice size, we can see from figures \ref{fig:E-T2-23}-\ref{fig:chi-T22-23} that this corresponds to increasing the resolution of our calculation, which is what one would expect, considering that we are getting a better approximation of a real magnetic system when increasing the lattice size.
+When parallelising our software we gain a significant speed up, as shown in Table \ref{tab:runtimes40} and \ref{tab:runtimes100}.
+The speed up does not decrease by a factor four if running four processes, due to each run having to prepare and do a clean up for each run.
 
@@
 \label{sec:conclusions}
-
-%\begin{figure}[h]
-%
-%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
-%
-%\caption{Description of figure -- explain all elements, but do not
-%draw conclusions here.}
-%\label{fig:figure_label}
-%\end{figure}
-
-%\begin{deluxetable}{lccc}
-%\tablewidth{0pt}
-%\tablecaption{\label{tab:results}}
-%\tablecomments{Summary of main results.}
-%\tablecolumns{4}
-%\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
-%\startdata
-%Item 1 & Item 2 & Item 3 & Item 4
-%\enddata
-%\end{deluxetable}
-
-\begin{acknowledgements}
-
-\end{acknowledgements}
+The Ising model for a finite lattice proves to give a good estimate of the analytical value of the critical temperature, which is the temperature of which a magnetic system undergoes a phase transformation.
+Increasing the lattice size gives a better approximations and a more realistic model, increasing it further than what has been done in this study will allow for a better approximation of the critical temperature.
+The figures provided for the mean energy and magnetisation, the heat capacity and the susceptibility shows a consistent behavior of a phase transition for a temperature $\sim T_C$.
 
 \begin{thebibliography}{}
-
-\bibitem[G{\'o}rski et al.(1994)]{gorski:1994} G{\'o}rski, K. M.,
-  Hinshaw, G., Banday, A. J., Bennett, C. L., Wright, E. L., Kogut,
-  A., Smoot, G. F., and Lubin, P.\ 1994, ApJL, 430, 89
-
+\bibitem{bib:onsager} Onsager, Lars,
+ \textit{Crystal Statistics. I. A Two-Dimensional Model with an Order-Disorder Transition}
+\url{https://link.aps.org/doi/10.1103/PhysRev.65.117}
 \end{thebibliography}

### 2018-11-20 Project4/Tc.py

@@ -2,11 +2,14 @@ import numpy as np
 import matplotlib.pyplot as plt
 import numpy.linalg
 
+"""
+To calculate the fitted critical temperature.
+"""
 
-file = "data/test/L40-1e6-vary.txt"
-file2 = "data/test/L60-1e6-vary.txt"
-file3 = "data/test/L80-1e6-vary.txt"
-file4 = "data/test/L100-1e6-vary.txt"
+file = "data/L40-1e6-vary.txt"
+file2 = "data/L60-1e6-vary.txt"
+file3 = "data/L80-1e6-vary.txt"
+file4 = "data/L100-1e6-vary.txt"
 
 
 chi, T = np.loadtxt(file, usecols=(3,4), unpack=True)

### 2018-11-20 Project4/analytical.py

@@ -1,5 +1,9 @@
 import numpy as np
 
+"""
+Returns the analytical values of the Ising model.
+"""
+
 n = 2
 T = 1.0
 

### 2018-11-20 Project4/main.cpp

@@ -13,11 +13,10 @@ using namespace arma;
 ofstream outfile;
 
 inline int PB(int index, int spins, int correction) {
-        // returns the index of the element
+        // Periodic boundary conidition, returns the index "periodic index"
         return (index + spins + correction)%(spins);
 }
 
-void Energy(int L, mat &S, double &energy);
 void initialize(double &energy, double &magMoment, int L, mat S, int GS);
 void Metropolis(int L, int mcs, double T, map<double, double> acceptE, int numprocs, int my_rank, int myloop_end, int myloop_begin, string filename);
 map<double, double> energies(double T);
@@ -63,6 +62,9 @@ int main(int argc, char *argv[]){
 }
 
 void initialize(double &energy, double &magMoment, int L, mat &S, int GS, mt19937_64 generator){
+        /* Initializes the spin matrix. If GS=1 we are initiating ground state
+           with all spin up. If GS=0 we set a random initial condition. Also
+           calculates energy and magnetic moment of initial state. */
         if (GS==1) {
                 S.fill(1.0); // ground state
                 magMoment += pow(double(L),2); // initial all spin up
@@ -75,7 +77,6 @@ void initialize(double &energy, double &magMoment, int L, mat &S, int GS, mt1993
                         }
                 }
         }
-
         for(int x = 0; x < L; x++) {
                 for(int y = 0; y < L; y++) {
                         energy -= (double) S(x,y)*(S(x,PB(y,L,-1)) + S(PB(x,L,-1),y));
@@ -85,13 +86,14 @@ void initialize(double &energy, double &magMoment, int L, mat &S, int GS, mt1993
 }
 
 void Metropolis(int L, int mcc, double T, map<double, double> acceptE, int numprocs, int my_rank, int myloop_end, int myloop_begin, string filename){
+        /* Metropolis algorithm. */
         random_device rd;
         mt19937_64 generator(rd());
-        uniform_real_distribution<float> RNG(0.0,1.0);
-        uniform_int_distribution<int> RNGPos(0,L-1);
+        uniform_real_distribution<float> RNG(0.0,1.0);  // Random number
+        uniform_int_distribution<int> RNGPos(0,L-1);    // random spin position of lattice
 
-        double energy = 0.0;  // energy
-        double magMoment = 0.0; // magnetic moment (initial value, of ground state = n²)
+        double energy = 0.0;
+        double magMoment = 0.0;
 
[…]

### 2018-11-20 Project4/plot.py

@@ -2,17 +2,21 @@ import matplotlib.pyplot as plt
 import numpy as np
 import os
 
+"""
+For plotting various data.
+"""
+
 accepts = False
-acceptOfT = False
+acceptOfT = True
 EM = False
-PE = True
+PE = False
 vary = False
 variance = False
 sigma = False
 
 if EM == True:
-    file = "data/L20-GS-T1-1e6.txt"
-    file2 = "data/L20-GS-T2_4-1e6.txt"
+    file = "data/L20-T1-1e6.txt"
+    file2 = "data/L20-T2_4-1e6.txt"
 
     E = np.loadtxt(file, usecols=0)
     M = np.loadtxt(file, usecols=1)
@@ -21,23 +25,23 @@ if EM == True:
     mcc = np.linspace(0,len(M),len(M))
 
     plt.figure()
-    #plt.plot(mcc,E)
-    plt.plot(mcc,E2)
+    plt.plot(mcc,E)
+    #plt.plot(mcc,E2)
     plt.legend(["T=1"], prop={'size': 15})
-    plt.title("Energy for a 20x20 lattice (ground state)", size=15)
+    plt.title("Energy of a 20x20 lattice", size=15)
     plt.xlabel("Monte Carlo cycles", size=15); plt.ylabel("$ E /L^2$", size=15)
     plt.show()
 
     plt.figure()
-    #plt.plot(mcc,M)
-    plt.plot(mcc,M2)
+    plt.plot(mcc,M)
+    #plt.plot(mcc,M2)
     plt.legend(["T=1"], prop={'size': 15})
-    plt.title("Magnetisation for a 20x20 lattice (ground state)", size=15)
+    plt.title("Magnetisation of a 20x20 lattice", size=15)
     plt.xlabel("Monte Carlo cycles", size=15); plt.ylabel("$|M|/L^2$", size=15)
     plt.show()
 
 if PE == True:
-    file = "data/test/L20-T2_4-1e6.txt"
+    file = "data/L20-T2_4-1e6.txt"
     E_sorted = np.sort(np.loadtxt(file, usecols=0))
     mcc = np.linspace(0,len(E_sorted),len(E_sorted))
     bins = 0
@@ -50,7 +54,7 @@ if PE == True:
[…]

### 2018-11-21 Project4/report/Project4.tex
Finished

@@
 The study will show that for a cold system of temperature 1 kT/J the system rarely has states of higher energies, when increasing the temperature to 2.4 kT/J the total energy of the system is increased, and we see that the probability for the system to be in a higher energy state is increased.
-When we look at how the system evolves as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature.
+When we look at how the system evolve as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature.
 When estimating the critical temperature we find this to be $T_C = 2.256$ kT/J, which is fairly close to the analytical value found by Lars Onsager \cite{bib:onsager}.
@@
 When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment.
+When we are calculating the final mean values of the energy and magnetic moment, we ignore the first 5\% of the data, so we are not including the values before the system is in equilibrium.
 
@@
 \begin{thebibliography}{}
-\bibitem{bib:onsager} Onsager, Lars,
+\bibitem{bib:onsager} 
+Onsager, Lars,
  \textit{Crystal Statistics. I. A Two-Dimensional Model with an Order-Disorder Transition}
 \url{https://link.aps.org/doi/10.1103/PhysRev.65.117}
+
 \end{thebibliography}
